### 2024-10-10 paper/paper.qmd
Add in Starter Files

---
title: "MarriageLicense"
subtitle: "My subtitle if needed"
author: 
  - Mariko Lee
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)

data <- read_csv(here::here("data/raw_data/raw_data.csv"))

```

# Introduction

You can and should cross-reference sections and sub-sections. We use @talia, @opendatatorotno, and @rohan.

The remainder of this paper is structured as follows. @sec-mydatasection

# Data {#sec-data}

Some of our data is of marriage license (@fig-marriage), from @TorontoOpenData

```{r}
#| label: fig-marriage
#| fig-cap: Marriage License
#| echo: false

data |> 
  ggplot(aes(x = date, y = marriage_licenses)) +
  geom_point()

```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

\newpage

# References

### 2024-10-10 scripts/01-download_data.R
Download president poll data

@@ -1,33 +1,19 @@
 #### Preamble ####
-# Purpose: Downloads and saves the data from Open Data Toronto 
-# Author: Mariko Lee
-# Date: 19 September 2024 
-# Contact: mariko.lee@mail.utoronto.ca
+# Purpose: Downloads and saves the data
+# Authors: Cristina Su Lam, Karen Riani, Mariko Lee
+# Date: 10 October 2024 
 # License: MIT
 # Pre-requisites: None
 # Any other information needed? None
 
-#### Workspace setup ####
-library(opendatatoronto)
-library(tidyverse)
+# Load necessary libraries
+library(readr)
 
-
-#### Download data ####
-# get package
-package <- show_package("e28bc818-43d5-43f7-b5d9-bdfb4eda5feb")
-
-# get all resources for this package
-resources <- list_package_resources("e28bc818-43d5-43f7-b5d9-bdfb4eda5feb")
-
-# identify datastore resources; by default, Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
-datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))
-
-# load the first datastore resource as a sample
-data <- filter(datastore_resources, row_number()==1) %>% 
-  get_resource()
-
-
-#### Save data ####
-write_csv(data, "data/raw_data/raw_data.csv") 
-
+# Define the path to the data
+data_path <- "data/raw_data/raw_president_polls.csv"
          
+# Read the CSV file into a data frame
+polls_data <- read_csv(data_path)
+
+# Display the first few rows of the data
+head(polls_data)

### 2024-10-10 paper/paper.qmd

@@
 ---
-title: "MarriageLicense"
+title: "temp"
 subtitle: "My subtitle if needed"
@@
   - Mariko Lee
-thanks: "Code and data are available at: LINK."
+  - Karen Riani
+  - Cristina Su Lam
+thanks: "Code and data are available at: https://projects.fivethirtyeight.com/polls/president-general/2024/national/"
 date: today

### 2024-10-19 paper/paper.qmd
Add appendix A

--- 2024-10-10 paper/paper.qmd
@@
 ---
-title: "MarriageLicense"
-subtitle: "My subtitle if needed"
+title: "2024 US Presidential Election Forecasting"
+subtitle: "subtitle"
 author: 
   - Mariko Lee
-thanks: "Code and data are available at: LINK."
+  - Karen Riani
+  - Cristina Su Lam
+thanks: "Code and data are available at: https://projects.fivethirtyeight.com/polls/president-general/2024/national/"
 date: today
@@
 
-# Additional data details
+# Methodology Analysis of The Washington Post Polling
+With an evaluation of sampling methodology, recruitment, handling non-response, and questionnaire design, this appendix offers an analysis of the polling methodology used by The Washington Post in collaboration with ABC News.
+The objective is to analyze these approaches’ strengths and weaknesses and determine how they affect polling accuracy.
+
+## Population, Frame, and Sample
+The Washington Post, in partnership with ABC News, employs a combination of text-to-web polls and random digit dialing (RDD) for landlines and mobile phones to reach a large and representative sample of American adults and registered voters [@WashPost2024_ABCMethodology].
+
+The Washington Post's polling averages use only national and state-level polls that comply with strict quality and transparency criteria.
+These surveys were chosen because they employ suitable stratification and weighting strategies in addition to random sample approaches [@WashPost2024_PollingAverages].
+To represent critical demographics such as age, race, gender, and education, the samples are meticulously weighted [@WashPost2024_ABCMethodology] [@WashPost2024_PollingAverages].
+
+## Sample Recruitment
+Live phone interviews and text-to-web surveys collect samples for The Washington Post polls, focusing on ensuring comprehensive demographic coverage.
+In a typical 2024 poll, text-to-web invites reached 21% of respondents, landlines reached 15%, and mobile phones reached 64% of respondents [@WashPost2024_ABCMethodology].
+Younger and minority voters, who might not be well represented in conventional landline-based surveys, can be efficiently reached by pollsters using this technique.
+
+By using address-based sampling from the Delivery Sequence File of the US Postal Service, ABC News also leverages probability-based recruiting through the Ipos KnowledgePanel.
+Since internet connections and equipment are offered at no cost, this guarantees that even households without internet connections or digital devices are involved [@ABCNews2024_Methodology].
+
+## Sampling Approaches and Trade-offs
+Using stratified random sampling, The Washington Post ensures that important demographic groups are represented proportionately to their voter base.
+By using stratified sampling, the polls are more likely to represent the diversity of the voting population accurately.
+To account for over- or under-representation of particular groups, samples are further weighted [@WashPost2024_ABCMethodology] [@WashPost2024_PollingAverages].
+
+Particularly in situations where state-level polling data is scarce, The Washington Post’s polling averages consider the state’s voting record in the last two presidential elections [@WashPost2024_PollingAverages].
+This adjustment offers a more accurate representation of voters' preferences in states with fewer high-quality polls.
+However, there may be a trade-off since, depending solely on historical data, we may miss recent shifts in voter sentiment [@WashPost2024_ABCMethodology].
+
+## Non-response Handling
+The Washington Post uses response weighting, which modifies the results according to demographic variables such as age, race, and education, in order to address non-response bias.
+In spite of variations in response rates among demographic groupings, this ensures that the final sample more accurately represents the population [@WashPost2023_Standard].
+
+ABC News also addressed non-response bias by applying post-stratification adjustments and sending email reminders to non-respondents.
+In addition, The Washington Post and ABC News both ensure that their samples are weighted to account for any anomalies in non-response [@ABCNews2024_Methodology] [@WashPost2023_Standards].
+
+Despite these initiatives, non-response bias is still a concern, especially for populations that are less inclined to take part in surveys, including younger or less politically active people [@WashPost2023_Standards].
+
+## Questionnaire Design
+To prevent respondents from being guided toward predetermined responses, The Washington Post creates its surveys with neutrality and clarity in mind.
+The questions are randomized, and respondents are given multiple choices, including “No Opinion,” to avoid pressuring answers [@WashPost2024_ABCMethodology].
+Question order bias can affect how respondents understand and respond to follow-up questions.
+Therefore, rotation helps mitigate this effect [@WashPost2023_Standards].
+
+Similar ideas are utilized by ABC News, which offers surveys in both Spanish and English to reach a more representative sample of the general public.
+Leading questions are purposefully omitted from the questionnaires to ensure that the information gathered accurately reflects public opinion [@ABCNews2024_Methodology].
+
+## Strengths and Weaknesses of the Methodology
+**Strength:**
+**Comprehensive Sampling Method:** The Washington Post can reach a broad demographic, including younger and more difficult-to-reach voters, by combining RDD, text-to-web polls, and live phone interviews [@WashPost2024_ABCMethodology] [@WashPost2023_Standards].
+**Post-stratification Weighting:** To account for demographic imbalances and increase the accuracy of their polls, The Washington Post and ABC News both use strong post-stratification weighting [@WashPost2024_ABCMethodology] [@ABCNews2024_Methodology].
+**Transparent Approach:** The Washington Post’s polling data is more credible since they only employ high-quality polls in their averages and is transparent about their methodology [@WashPost2023_Standards] [@WashPost2024_PollingAverages].
+**Weaknesses:**
+**Non-response Bias:** Even if both organizations use weighting adjustments, non-response bias still remains a challenge, particularly when it comes to groups that are less likely to respond to surveys [@WashPost2023_Standards] [@WashPost2024_ABCMethodology].
+**Dependency on Historical Data:** In states with fewer polls, The Washington Post relies on historical data (the last two presidential elections), which raises the possibility that the polling averages might not accurately reflect current changes in voters preferences [@WashPost2024_PollingAverages].
+
+## Conclusion
+The polling methodologies used by The Washington Post and ABC News offer a strict framework for gauging popular sentiment in the 2024 US presidential election.
+Their surveys often represent the electorate since they employ various sampling strategies, stratification, and weighting methodologies.
+However, obstacles such as non-response bias and the use of historical data in some states must be addressed appropriately to protect the accuracy and reliability of their polling averages.
 

### 2024-10-19 scripts/00-simulate_data.R
add title

@@ -1,8 +1,7 @@
 #### Preamble ####
 # Purpose: Simulates Data
-# Author: Mariko Lee
-# Date: 19 September 2024
-# Contact: mariko.lee@mail.utoronto.ca =
+# Author: Cristina Su Lam, Karen Riani, Mariko Lee
+# Date: 10 October 2024 
 # License: MIT
 # Pre-requisites: None
 # Any other information needed? None

### 2024-10-19 scripts/01-download_data.R

@@ -7,6 +7,7 @@
 # Any other information needed? None
 
 # Load necessary libraries
+library(tidyverse)
 library(readr)
 
 # Define the path to the data
@@ -17,3 +18,4 @@ polls_data <- read_csv(data_path)
 
 # Display the first few rows of the data
 head(polls_data)
+

### 2024-10-19 scripts/02-data_cleaning.R

@@ -1,8 +1,7 @@
 #### Preamble ####
-# Purpose: Cleans the raw marriage data into an analysis dataset
-# Author: Mariko Lee
-# Date: 19 September 2024
-# Contact: mariko.lee@mail.utoronto.ca 
+# Purpose: Cleans the presidential poll data into an analysis data set
+# Author: Cristina Su Lam, Karen Riani, Mariko Lee
+# Date: 10 October 2024 
 # License: MIT
 # Pre-requisites: Need to have downloaded the data
 # Any other information needed? None
@@ -11,16 +10,14 @@
 library(tidyverse)
 
 #### Clean data ####
-raw_data <- read_csv("data/raw_data/raw_data.csv")
+poll_data <- read_csv("data/raw_data/raw_president_polls.csv")
 
-cleaned_data <-
-  raw_data |>
-  janitor::clean_names() |> 
-  separate(col = time_period,
-           into = c("year", "month"),
-           sep = "-") |> 
-  mutate(date = lubridate::ymd(paste(year, month, "01", sep = "-"))
-  )
+polls_data <- polls_data %>%
+  janitor::clean_names()
+
+# Handle missing values (example: remove incomplete rows)
+poll_data <- poll_data %>%
+  drop_na()
 
 #### Save data ####
-write_csv(cleaned_data, "data/analysis_data/analysis_data.csv")
+write_csv(analysis_data, "data/analysis_data/analysis_data.csv")

### 2024-10-19 scripts/03-test_data.R

@@ -1,8 +1,7 @@
 #### Preamble ####
 # Purpose: Sanity check of the data
-# Author: Mariko Lee
-# Date: 19 September 2024
-# Contact: mariko.lee@mail.utoronto.ca
+# Author: Cristina Su Lam, Karen Riani, Mariko Lee
+# Date: 10 October 2024 
 # License: MIT
 # Pre-requisites: Need to have simulated data
 # Any other information needed? None

### 2024-10-21 scripts/00-simulate_data.R
Added inital simulation- training and test sets + logistic regression
issue with my RStudio, had to manually add

@@ -8,14 +8,49 @@
 
 #### Workspace setup ####
 library(tidyverse)
+library(tidymodels)
+library(arrow)
 
+#### Read Data ####
+
+upef2024_data <- 
+  read_parquet(file = here::here("data/analysis_data/polls_data.parquet"))
+# here::here not necessary, just more robust 
 
 #### Simulate data ####
-set.seed(304)
+set.seed(333)
 
-# Define the start and end date
-start_date <- as.Date("2018-01-01")
-end_date <- as.Date("2023-12-31")
+## Split data into test and training sets
+upef2024_split <- initial_split(upef2024_data, prop = 0.80)
+upef2024_train <- training(upef2024_split)
+upef2024_test <- testing(upef2024_split)
+
+## Fit logistic model
+upef2024_tidymodels <-
+  logistic_reg(mode = "classification") |>
+  set_engine("glm") |>
+  fit(
+    pct ~ pollster + is_state,
+    data = upef2024_train
+  )
+
+upef2024_tidymodels
+
+## Evaluate on test set.. this is not
+upef2024_tidymodels |>
+  predict(new_data = upef2024_test) |>
+  cbind(upef2024_test) |>
+  conf_mat(truth = pct, estimate = .pred_class)
+
+## K-Fold Cross Validation
+
+
+
+### Simulation Tests
+
+# Confirm the start and end date
+start_date <- as.Date("2024-01-01")
+end_date <- as.Date("2024-00-00")
 
 # Set the number of random dates you want to generate
 number_of_dates <- 100

### 2024-10-21 scripts/01-download_data.R
Added initial filters and mutations
categorized pct, binary for national/state, filtered for Harris and quality

@@ -1,21 +1,40 @@
 #### Preamble ####
-# Purpose: Downloads and saves the data
-# Authors: Cristina Su Lam, Karen Riani, Mariko Lee
+# Purpose: Cleans the presidential poll data into an analysis data set
+# Author: Cristina Su Lam, Karen Riani, Mariko Lee
 # Date: 10 October 2024 
 # License: MIT
-# Pre-requisites: None
+# Pre-requisites: Need to have downloaded the data
 # Any other information needed? None
 
-# Load necessary libraries
+#### Workspace setup ####
 library(tidyverse)
-library(readr)
+library(arrow)
 
-# Define the path to the data
-data_path <- "data/raw_data/raw_president_polls.csv"
-         
-# Read the CSV file into a data frame
-polls_data <- read_csv(data_path)
+#### Clean data ####
+polls_data <- read_csv("data/raw_data/raw_president_polls.csv")
 
-# Display the first few rows of the data
-head(polls_data)
+#polls_data <- polls_data %>%
+ # janitor::clean_names()
+
+# Handle missing values (example: remove incomplete rows)
+#polls_data <- polls_data %>%
+ # drop_na()
+
+# Binary for state or national polls
+polls_data <- polls_data %>%
+  mutate(is_state = as.numeric(!is.na(state)))
+
+# Categorize pct for Harris into High and Low Support
+polls_data <- polls_data %>%
+  mutate(pct = as.factor(ifelse(pct > 50, "High", "Low")))
+
+# Filter for high quality pollsters and state-specific Harris polls
+polls_data %>%
+  filter(numeric_grade >= 3.0, 
+         candidate_name == "Kamala Harris")
+
+#### Save data ####
+
+write_parquet(polls_data, "data/analysis_data/polls_data.parquet")
+# delete analysis data csv 
 

### 2024-10-21 scripts/02-data_cleaning.R
Update cleaning data

@@ -8,16 +8,34 @@
 
 #### Workspace setup ####
 library(tidyverse)
+library(arrow)
 
 #### Clean data ####
-poll_data <- read_csv("data/raw_data/raw_president_polls.csv")
+high_quality_harris <- read_csv("data/raw_data/raw_president_polls.csv") |>
+  clean_names()
 
-polls_data <- polls_data %>%
-  janitor::clean_names()
+# Binary for state or national polls
+high_quality_harris <- high_quality_harris %>%
+  mutate(is_state = as.numeric(!is.na(state)),
+  end_date = mdy(end_date))
 
-# Handle missing values (example: remove incomplete rows)
-poll_data <- poll_data %>%
-  drop_na()
+# Categorize pct for Harris into High and Low Support
+high_quality_harris <- high_quality_harris %>%
+  mutate(pct = as.factor(ifelse(pct > 50, "High", "Low")))
+
+# Filter for high quality pollsters and state-specific Harris polls
+high_quality_harris %>%
+  filter(
+    numeric_grade >= 3.0, 
+    candidate_name == "Kamala Harris"
+  ) |>
+  filter(end_date >= as.Date("2024-07-21")) |> # When Harris declared
+  mutate(
+    num_harris = round((pct / 100) * sample_size, 0) # Need number not percent for some models
+  )
+
 
 #### Save data ####
-write_csv(analysis_data, "data/analysis_data/analysis_data.csv")
+write_parquet(high_quality_harris, "data/analysis_data/high_quality_harris.parquet")
+
+# delete analysis data csv 

### 2024-10-22 scripts/02-data_cleaning.R
corrected mutations

@@ -1,41 +1,0 @@
-#### Preamble ####
-# Purpose: Cleans the presidential poll data into an analysis data set
-# Author: Cristina Su Lam, Karen Riani, Mariko Lee
-# Date: 10 October 2024 
-# License: MIT
-# Pre-requisites: Need to have downloaded the data
-# Any other information needed? None
-
-#### Workspace setup ####
-library(tidyverse)
-library(arrow)
-
-#### Clean data ####
-high_quality_harris <- read_csv("data/raw_data/raw_president_polls.csv") |>
-  clean_names()
-
-# Binary for state or national polls
-high_quality_harris <- high_quality_harris %>%
-  mutate(is_state = as.numeric(!is.na(state)),
-  end_date = mdy(end_date))
-
-# Categorize pct for Harris into High and Low Support
-high_quality_harris <- high_quality_harris %>%
-  mutate(pct = as.factor(ifelse(pct > 50, "High", "Low")))
-
-# Filter for high quality pollsters and state-specific Harris polls
-high_quality_harris %>%
-  filter(
-    numeric_grade >= 3.0, 
-    candidate_name == "Kamala Harris"
-  ) |>
-  filter(end_date >= as.Date("2024-07-21")) |> # When Harris declared
-  mutate(
-    num_harris = round((pct / 100) * sample_size, 0) # Need number not percent for some models
-  )
-
-
-#### Save data ####
-write_parquet(high_quality_harris, "data/analysis_data/high_quality_harris.parquet")
-
-# delete analysis data csv 

### 2024-10-22 scripts/02_analysis_and_modelling/Bayesian_models.R
Added class examples

@@ -1,0 +1,148 @@
+#### Preamble ####
+# Purpose: !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
+# Author: Cristina Su Lam, Karen Riani, Mariko Lee
+# Date: 10 October 2024 
+# License: MIT
+# Pre-requisites: Need to have downloaded the data
+# Any other information needed? None
+
+#### Workspace setup ####
+library(tidyverse)
+library(janitor)
+library(lubridate)
+library(broom)
+library(modelsummary)
+library(rstanarm)
+library(splines)
+
+#### Read Data ####
+upef2024_data <- 
+  read_parquet(file = here::here("data/analysis_data/high_quality_harris.parquet"))
+
+#### Bayesian models ####
+# Change 'pollster' and 'state' to factor variables
+upef2024_data <- upef2024_data |>
+  mutate(
+    pollster = factor(pollster),
+    state = factor(state)
+  )
+
+### Multi-level Regression ###
+# Model 1
+model_formula_1 <- cbind(num_harris, sample_size - num_harris) ~ (1 | pollster)
+
+# Model 2
+model_formula_2 <- cbind(num_harris, sample_size - num_harris) ~ (1 | pollster) + (1 | state)
+
+# Specify priors
+priors <- normal(0, 2.5, autoscale = TRUE)
+
+# Fit the models
+bayesian_model_1 <- stan_glmer(
+  formula = model_formula_1,
+  data = upef2024_data,
+  family = binomial(link = "logit"),
+  prior = priors,
+  prior_intercept = priors,
+  seed = 123,
+  cores = 4,
+  adapt_delta = 0.95
+)
+
+bayesian_model_2 <- stan_glmer(
+  formula = model_formula_2,
+  data = upef2024_data,
+  family = binomial(link = "logit"),
+  prior = priors,
+  prior_intercept = priors,
+  seed = 123,
+  cores = 4,
+  adapt_delta = 0.95
+)
+# Note the warnings here and follow the instructions for dealing with it - come back and fix.
+
+# Posterior predictive checks
+pp_check(bayesian_model_1)
+pp_check(bayesian_model_2)
+
+# Summarize the model
+summary(bayesian_model_1)
+summary(bayesian_model_2)
+
+# Plot random effects
+plot(bayesian_model_1, pars = "(Intercept)", prob = 0.95)
+plot(bayesian_model_2, pars = "(Intercept)", prob = 0.95)
+
+# Model summary works the same as above for Bayesian models.
+
[…]

### 2024-10-22 scripts/02_analysis_and_modelling/initial_dataset_models.R

@@ -1,0 +1,51 @@
+#### Preamble ####
+# Purpose: !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
+# Author: Cristina Su Lam, Karen Riani, Mariko Lee
+# Date: 10 October 2024 
+# License: MIT
+# Pre-requisites: Need to have downloaded the data
+# Any other information needed? None
+
+#### Workspace setup ####
+library(tidyverse)
+library(janitor)
+library(lubridate)
+library(broom)
+library(modelsummary)
+
+#### Read Data ####
+high_quality_harris <- 
+  read_parquet(file = here::here("data/analysis_data/high_quality_harris.parquet"))
+
+#### Plot data ####
+base_plot <- ggplot(high_quality_harris, aes(x = end_date, y = pct)) +
+  theme_classic() +
+  labs(y = "Harris percent", x = "Date")
+
+# Plots poll estimates and overall smoothing
+base_plot +
+  geom_point() +
+  geom_smooth()
+
+# Color by pollster
+# This gets messy - need to add a filter - see line 21
+base_plot +
+  geom_point(aes(color = pollster)) +
+  geom_smooth() +
+  theme(legend.position = "bottom")
+
+# Facet by pollster
+# Make the line 21 issue obvious
+# Also - is there duplication???? Need to go back and check
+base_plot +
+  geom_point() +
+  geom_smooth() +
+  facet_wrap(vars(pollster))
+
+print(unique(high_quality_harris$pollster))
+
+# Color by pollscore
+base_plot +
+  geom_point(aes(color = factor(pollscore))) +
+  geom_smooth() +
+  theme(legend.position = "bottom")

### 2024-10-22 scripts/02_analysis_and_modelling/linear_models.R

@@ -1,0 +1,203 @@
+#### Preamble ####
+# Purpose: !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
+# Author: Cristina Su Lam, Karen Riani, Mariko Lee
+# Date: 10 October 2024 
+# License: MIT
+# Pre-requisites: Need to have downloaded the data
+# Any other information needed? None
+
+#### Workspace setup ####
+
+#### Plot data ####
+base_plot <- ggplot(high_quality_harris, aes(x = end_date, y = pct)) +
+  theme_classic() +
+  labs(y = "Harris percent", x = "Date")
+
+# Plots poll estimates and overall smoothing
+base_plot +
+  geom_point() +
+  geom_smooth()
+
+# Color by pollster
+# This gets messy - need to add a filter - see line 21
+base_plot +
+  geom_point(aes(color = pollster)) +
+  geom_smooth() +
+  theme(legend.position = "bottom")
+
+# Facet by pollster
+# Make the line 21 issue obvious
+# Also - is there duplication???? Need to go back and check
+base_plot +
+  geom_point() +
+  geom_smooth() +
+  facet_wrap(vars(pollster))
+
+# Color by pollscore
+base_plot +
+  geom_point(aes(color = factor(pollscore))) +
+  geom_smooth() +
+  theme(legend.position = "bottom")
+
+
+#### Starter models ####
+# Model 1: pct as a function of end_date
+model_date <- lm(pct ~ end_date, data = high_quality_harris)
+
+# Model 2: pct as a function of end_date and pollster
+model_date_pollster <- lm(pct ~ end_date + pollster, data = high_quality_harris)
+
+# Augment data with model predictions
+high_quality_harris <- high_quality_harris |>
+  mutate(
+    fitted_date = predict(model_date),
+    fitted_date_pollster = predict(model_date_pollster)
+  )
+
+# Plot model predictions
+# Model 1
+ggplot(high_quality_harris, aes(x = end_date)) +
+  geom_point(aes(y = pct), color = "black") +
+  geom_line(aes(y = fitted_date), color = "blue", linetype = "dotted") +
+  theme_classic() +
+  labs(y = "Harris percent", x = "Date", title = "Linear Model: pct ~ end_date")
+
+# Model 2
+ggplot(high_quality_harris, aes(x = end_date)) +
+  geom_point(aes(y = pct), color = "black") +
[…]

### 2024-10-22 scripts/01_data_cleaning_and_preparation/02-data_cleaning.R
moved cleaning script

@@ -1,0 +1,46 @@
+#### Preamble ####
+# Purpose: Cleans the presidential poll data into an analysis data set
+# Author: Cristina Su Lam, Karen Riani, Mariko Lee
+# Date: 10 October 2024 
+# License: MIT
+# Pre-requisites: Need to have downloaded the data
+# Any other information needed? None
+
+#### Workspace setup ####
+library(tidyverse)
+library(arrow)
+library(janitor)
+library(lubridate)
+library(broom)
+
+#### Clean data ####
+high_quality_harris <- read_csv("data/raw_data/raw_president_polls.csv") |>
+  clean_names()
+
+# Binary for state or national polls
+high_quality_harris <- high_quality_harris %>%
+  mutate(is_state = as.numeric(!is.na(state)),
+  end_date = mdy(end_date))
+
+# Categorize pct for Harris into High and Low Support in new col
+high_quality_harris$pct_category <- ifelse(high_quality_harris$pct > 50, "High", "Low")
+high_quality_harris$pct_category <- as.factor(high_quality_harris$pct_category)
+
+# Filter for high quality pollsters and state-specific Harris polls
+high_quality_harris %>%
+  filter(
+    numeric_grade >= 3.0, 
+    candidate_name == "Kamala Harris"
+  ) 
+
+high_quality_harris <- high_quality_harris %>%
+  filter(end_date >= as.Date("2024-07-21"))
+high_quality_harris <- high_quality_harris %>%
+  mutate(
+    num_harris = round((pct / 100) * sample_size, 0) # Convert pct to actual count of Harris supporters
+  )
+
+#### Save data ####
+write_parquet(high_quality_harris, "data/analysis_data/high_quality_harris.parquet")
+
+# delete analysis data csv 

### 2024-10-22 scripts/01_data_cleaning_and_preparation/01-download_data.R
Cleaned Raw Dataset

@@ -1,40 +1,17 @@
 #### Preamble ####
-# Purpose: Cleans the presidential poll data into an analysis data set
+# Purpose: Download the presidential poll data 
 # Author: Cristina Su Lam, Karen Riani, Mariko Lee
 # Date: 10 October 2024 
 # License: MIT
-# Pre-requisites: Need to have downloaded the data
+# Pre-requisites: None
 # Any other information needed? None
 
 #### Workspace setup ####
-library(tidyverse)
-library(arrow)
+library(readr)
 
-#### Clean data ####
-polls_data <- read_csv("data/raw_data/raw_president_polls.csv")
+#### Download/Read Dataset ####
+raw_president_polls <- read_csv("data/raw_data/raw_president_polls.csv")
 
-#polls_data <- polls_data %>%
- # janitor::clean_names()
 
-# Handle missing values (example: remove incomplete rows)
-#polls_data <- polls_data %>%
- # drop_na()
 
-# Binary for state or national polls
-polls_data <- polls_data %>%
-  mutate(is_state = as.numeric(!is.na(state)))
-
-# Categorize pct for Harris into High and Low Support
-polls_data <- polls_data %>%
-  mutate(pct = as.factor(ifelse(pct > 50, "High", "Low")))
-
-# Filter for high quality pollsters and state-specific Harris polls
-polls_data %>%
-  filter(numeric_grade >= 3.0, 
-         candidate_name == "Kamala Harris")
-
-#### Save data ####
-
-write_parquet(polls_data, "data/analysis_data/polls_data.parquet")
-# delete analysis data csv 
 

### 2024-10-22 scripts/01_data_cleaning_and_preparation/02-data_cleaning.R
Cleaned Raw Dataset

@@ -3,10 +3,11 @@
 # Author: Cristina Su Lam, Karen Riani, Mariko Lee
 # Date: 10 October 2024 
 # License: MIT
-# Pre-requisites: Need to have downloaded the data
+# Pre-requisites: Run 01-download_data.R
 # Any other information needed? None
 
 #### Workspace setup ####
+library(readr)
 library(tidyverse)
 library(arrow)
 library(janitor)
@@ -14,33 +15,52 @@ library(lubridate)
 library(broom)
 
 #### Clean data ####
-high_quality_harris <- read_csv("data/raw_data/raw_president_polls.csv") |>
+clean_president_polls <- read_csv("data/raw_data/raw_president_polls.csv", 
+                                col_types = cols(poll_id = col_skip(), 
+                                                 pollster_id = col_skip(), sponsor_ids = col_skip(), 
+                                                 sponsors = col_skip(), display_name = col_skip(), 
+                                                 pollster_rating_id = col_skip(), 
+                                                 pollster_rating_name = col_skip(), 
+                                                 pollscore = col_skip(), methodology = col_skip(), 
+                                                 transparency_score = col_skip(), 
+                                                 sponsor_candidate_id = col_skip(), 
+                                                 sponsor_candidate = col_skip(), sponsor_candidate_party = col_skip(), 
+                                                 endorsed_candidate_id = col_skip(), 
+                                                 endorsed_candidate_name = col_skip(), 
+                                                 endorsed_candidate_party = col_skip(), 
+                                                 question_id = col_skip(), population = col_skip(), 
+                                                 subpopulation = col_skip(), population_full = col_skip(), 
+                                                 tracking = col_skip(), created_at = col_skip(), 
+                                                 notes = col_skip(), url = col_skip(), 
+                                                 url_article = col_skip(), url_topline = col_skip(), 
+                                                 url_crosstab = col_skip(), source = col_skip(), 
+                                                 internal = col_skip(), partisan = col_skip(), 
+                                                 race_id = col_skip(), cycle = col_skip(), 
+                                                 office_type = col_skip(), seat_number = col_skip(), 
+                                                 seat_name = col_skip(), election_date = col_skip(), 
[…]

### 2024-10-22 scripts/01_data_cleaning_and_preparation/02-data_cleaning.R
Filtering only by candidate_name

@@ -48,17 +48,20 @@ clean_president_polls <- read_csv("data/raw_data/raw_president_polls.csv",
 # Create binary variable for state or national polls
 clean_president_polls <- clean_president_polls %>%
   mutate(
+    state = str_replace(state, "Maine CD-[12]", "Maine"),
+    state = str_replace(state, "Nebraska CD-2", "Nebraska"),
     is_state = as.numeric(!is.na(state)),
     end_date = mdy(end_date),
     start_date = mdy(start_date),
     is_harris = ifelse(candidate_name == "Kamala Harris", 1, 0))
 
+
 # Filter for high quality pollsters and state-specific Harris and Trump polls
 clean_president_polls <- clean_president_polls %>%
-  filter(numeric_grade >= 3.0,
-         candidate_name %in% c("Kamala Harris", "Donald Trump"),
-         start_date >= as.Date("2024-07-21")
+  filter(
+         candidate_name %in% c("Kamala Harris", "Donald Trump")
          ) 
+
 
 #### Save data ####
 write_parquet(clean_president_polls, "data/analysis_data/clean_president_polls.parquet")

### 2024-10-22 scripts/01_data_cleaning_and_preparation/01-download_data.R
Read Raw Dataset

@@ -10,7 +10,7 @@
 library(readr)
 
 #### Download/Read Dataset ####
-raw_president_polls <- read_csv("data/raw_data/raw_president_polls.csv")
+raw_president_polls <- read_csv("data/01-raw_data/raw_president_polls.csv")
 
 
 

### 2024-10-22 scripts/01_data_cleaning_and_preparation/02-data_cleaning.R
Cleaned Raw Dataset

@@ -15,7 +15,7 @@ library(lubridate)
 library(broom)
 
 #### Clean data ####
-clean_president_polls <- read_csv("data/raw_data/raw_president_polls.csv", 
+clean_president_polls <- read_csv("data/01-raw_data/raw_president_polls.csv", 
                                 col_types = cols(poll_id = col_skip(), 
                                                  pollster_id = col_skip(), sponsor_ids = col_skip(), 
                                                  sponsors = col_skip(), display_name = col_skip(), 
@@ -62,8 +62,7 @@ clean_president_polls <- clean_president_polls %>%
          candidate_name %in% c("Kamala Harris", "Donald Trump")
          ) 
 
-
 #### Save data ####
-write_parquet(clean_president_polls, "data/analysis_data/clean_president_polls.parquet")
+write_parquet(clean_president_polls, "data/02-analysis_data/clean_president_polls.parquet")
 
 

### 2024-10-22 scripts/00_simulations/00-simulate_data.R
Delete scripts/00_simulations directory

@@ -1,75 +1,0 @@
-#### Preamble ####
-# Purpose: Simulates Data
-# Author: Cristina Su Lam, Karen Riani, Mariko Lee
-# Date: 10 October 2024 
-# License: MIT
-# Pre-requisites: None
-# Any other information needed? None
-
-#### Workspace setup ####
-library(tidyverse)
-library(tidymodels)
-library(arrow)
-
-#### Read Data ####
-
-upef2024_data <- 
-  read_parquet(file = here::here("data/analysis_data/polls_data.parquet"))
-# here::here not necessary, just more robust 
-
-#### Simulate data ####
-set.seed(333)
-
-## Split data into test and training sets
-upef2024_split <- initial_split(upef2024_data, prop = 0.80)
-upef2024_train <- training(upef2024_split)
-upef2024_test <- testing(upef2024_split)
-
-## Fit logistic model
-upef2024_tidymodels <-
-  logistic_reg(mode = "classification") |>
-  set_engine("glm") |>
-  fit(
-    pct ~ pollster + is_state,
-    data = upef2024_train
-  )
-
-upef2024_tidymodels
-
-## Evaluate on test set.. this is not
-upef2024_tidymodels |>
-  predict(new_data = upef2024_test) |>
-  cbind(upef2024_test) |>
-  conf_mat(truth = pct, estimate = .pred_class)
-
-## K-Fold Cross Validation
-
-
-
-### Simulation Tests
-
-# Confirm the start and end date
-start_date <- as.Date("2024-01-01")
-end_date <- as.Date("2024-00-00")
-
-# Set the number of random dates you want to generate
-number_of_dates <- 100
-
-data <-
-  tibble(
-    dates = as.Date(
-      runif(
-        n = number_of_dates,
-        min = as.numeric(start_date),
-        max = as.numeric(end_date)
-      ),
-      origin = "1970-01-01"
-    ),
-    number_of_marriage = rpois(n = number_of_dates, lambda = 15)
-  )
-
-
-#### Write_csv
-write_csv(data, file = "data/raw_data/simulated.csv")
-
-

### 2024-10-22 scripts/00-simulate_data.R
Simulated Data and Tested

@@ -1,52 +1,65 @@
 #### Preamble ####
-# Purpose: Simulates a dataset of Australian electoral divisions, including the 
-  #state and party that won each division.
-# Author: Rohan Alexander
-# Date: 26 September 2024
-# Contact: rohan.alexander@utoronto.ca
+# Purpose: Simulates Data
+# Author: Cristina Su Lam, Karen Riani, Mariko Lee
+# Date: 10 October 2024 
 # License: MIT
-# Pre-requisites: The `tidyverse` package must be installed
-# Any other information needed? Make sure you are in the `starter_folder` rproj
-
+# Pre-requisites: None
+# Any other information needed? None
 
 #### Workspace setup ####
 library(tidyverse)
-set.seed(853)
-
+set.seed(123)
 
 #### Simulate data ####
-# State names
-states <- c(
-  "New South Wales",
-  "Victoria",
-  "Queensland",
-  "South Australia",
-  "Western Australia",
-  "Tasmania",
-  "Northern Territory",
-  "Australian Capital Territory"
+# Defining the states and candidates
+states <- state.name  # U.S. state names
+candidates <- c("Trump", "Harris")  # Presidential candidates
+
+# Define a list of pollsters
+pollsters <- c("YouGov", "Siena/NYT", "The Washington Post", "Ipsos", "Quinnipiac")
+
+# Number of simulated polls per state
+num_polls <- 10  # Simulating 10 polls per state
+
+# Simulate sample sizes for each poll (between 500 and 2000 voters)
+sample_sizes <- sample(500:2000, num_polls * length(states), replace = TRUE)
+
+# Simulate percentage of support (pct) for each candidate (between 40% and 60%)
+pct_support <- runif(num_polls * length(states), min = 40, max = 60)
+
+# Create the simulated data frame
+simulated_data <- tibble(
+  state = rep(states, each = num_polls),  # Repeat each state for the number of polls
+  pollster = sample(pollsters, num_polls * length(states), replace = TRUE),  # Random pollsters for each poll
+  candidate_preference = sample(candidates, num_polls * length(states), replace = TRUE, prob = c(0.5, 0.5)),  # Equal chance for each candidate
+  pct = pct_support,  # Simulated percentage support
+  sample_size = sample_sizes  # Simulated sample sizes
 )
 
-# Political parties
[…]

### 2024-10-22 scripts/01-test_simulated_data.R

@@ -1,89 +1,49 @@
 #### Preamble ####
-# Purpose: Tests the structure and validity of the simulated Australian 
-  #electoral divisions dataset.
-# Author: Rohan Alexander
-# Date: 26 September 2024
-# Contact: rohan.alexander@utoronto.ca
+# Purpose: Tests the structure and validity of the simulated American elections dataset.
+# Author: Cristina Su Lam, Karen Riani, Mariko Lee
+# Date: 10 October 2024 
 # License: MIT
 # Pre-requisites: 
   # - The `tidyverse` package must be installed and loaded
   # - 00-simulate_data.R must have been run
-# Any other information needed? Make sure you are in the `starter_folder` rproj
+# Any other information needed? None
 
 
 #### Workspace setup ####
 library(tidyverse)
 
-analysis_data <- read_csv("data/00-simulated_data/simulated_data.csv")
+simulated_data <- read_csv("data/00-simulated_data/simulated_data.csv")
 
 # Test if the data was successfully loaded
-if (exists("analysis_data")) {
+if (exists("simulated_data")) {
   message("Test Passed: The dataset was successfully loaded.")
 } else {
   stop("Test Failed: The dataset could not be loaded.")
 }
 
+#### Test Simulated Data ####
 
-#### Test data ####
+expected_candidates <- c("Trump", "Harris")
+expected_states <- state.name
+expected_pollsters <- c("YouGov", "Siena/NYT", "The Washington Post", "Ipsos", "Quinnipiac")
 
-# Check if the dataset has 151 rows
-if (nrow(analysis_data) == 151) {
-  message("Test Passed: The dataset has 151 rows.")
-} else {
-  stop("Test Failed: The dataset does not have 151 rows.")
-}
+# Test 1: Ensure there are no missing values
+stopifnot(!any(is.na(simulate_data)))
 
-# Check if the dataset has 3 columns
-if (ncol(analysis_data) == 3) {
-  message("Test Passed: The dataset has 3 columns.")
-} else {
-  stop("Test Failed: The dataset does not have 3 columns.")
-}
+# Test 2: Ensure `pct` values are within a valid range (0% to 100%)
+stopifnot(all(simulate_data$pct >= 0 & simulate_data$pct <= 100))
 
-# Check if all values in the 'division' column are unique
-if (n_distinct(analysis_data$division) == nrow(analysis_data)) {
-  message("Test Passed: All values in 'division' are unique.")
-} else {
-  stop("Test Failed: The 'division' column contains duplicate values.")
-}
[…]

### 2024-10-22 scripts/01_data_cleaning_and_preparation/01-download_data.R
Delete scripts/01_data_cleaning_and_preparation directory

@@ -1,17 +1,0 @@
-#### Preamble ####
-# Purpose: Download the presidential poll data 
-# Author: Cristina Su Lam, Karen Riani, Mariko Lee
-# Date: 10 October 2024 
-# License: MIT
-# Pre-requisites: None
-# Any other information needed? None
-
-#### Workspace setup ####
-library(readr)
-
-#### Download/Read Dataset ####
-raw_president_polls <- read_csv("data/01-raw_data/raw_president_polls.csv")
-
-
-
-

### 2024-10-22 scripts/01_data_cleaning_and_preparation/02-data_cleaning.R

@@ -1,68 +1,0 @@
-#### Preamble ####
-# Purpose: Cleans the presidential poll data into an analysis data set
-# Author: Cristina Su Lam, Karen Riani, Mariko Lee
-# Date: 10 October 2024 
-# License: MIT
-# Pre-requisites: Run 01-download_data.R
-# Any other information needed? None
-
-#### Workspace setup ####
-library(readr)
-library(tidyverse)
-library(arrow)
-library(janitor)
-library(lubridate)
-library(broom)
-
-#### Clean data ####
-clean_president_polls <- read_csv("data/01-raw_data/raw_president_polls.csv", 
-                                col_types = cols(poll_id = col_skip(), 
-                                                 pollster_id = col_skip(), sponsor_ids = col_skip(), 
-                                                 sponsors = col_skip(), display_name = col_skip(), 
-                                                 pollster_rating_id = col_skip(), 
-                                                 pollster_rating_name = col_skip(), 
-                                                 pollscore = col_skip(), methodology = col_skip(), 
-                                                 transparency_score = col_skip(), 
-                                                 sponsor_candidate_id = col_skip(), 
-                                                 sponsor_candidate = col_skip(), sponsor_candidate_party = col_skip(), 
-                                                 endorsed_candidate_id = col_skip(), 
-                                                 endorsed_candidate_name = col_skip(), 
-                                                 endorsed_candidate_party = col_skip(), 
-                                                 question_id = col_skip(), population = col_skip(), 
-                                                 subpopulation = col_skip(), population_full = col_skip(), 
-                                                 tracking = col_skip(), created_at = col_skip(), 
-                                                 notes = col_skip(), url = col_skip(), 
-                                                 url_article = col_skip(), url_topline = col_skip(), 
-                                                 url_crosstab = col_skip(), source = col_skip(), 
-                                                 internal = col_skip(), partisan = col_skip(), 
-                                                 race_id = col_skip(), cycle = col_skip(), 
-                                                 office_type = col_skip(), seat_number = col_skip(), 
-                                                 seat_name = col_skip(), election_date = col_skip(), 
-                                                 stage = col_skip(), nationwide_batch = col_skip(), 
-                                                 ranked_choice_reallocated = col_skip(), 
[…]

### 2024-10-22 scripts/02-download_data.R
Read and Cleaned Raw Data

@@ -1,26 +1,16 @@
 #### Preamble ####
-# Purpose: Downloads and saves the data from [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Download the presidential poll data 
+# Author: Cristina Su Lam, Karen Riani, Mariko Lee
+# Date: 10 October 2024 
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
+# Pre-requisites: None
+# Any other information needed? None
 
 #### Workspace setup ####
-library(opendatatoronto)
-library(tidyverse)
-# [...UPDATE THIS...]
+library(readr)
 
-#### Download data ####
-# [...ADD CODE HERE TO DOWNLOAD...]
+#### Download/Read Dataset ####
+raw_president_polls <- read_csv("data/01-raw_data/raw_president_polls.csv")
 
-
-
-#### Save data ####
-# [...UPDATE THIS...]
-# change the_raw_data to whatever name you assigned when you downloaded it.
-write_csv(the_raw_data, "inputs/data/raw_data.csv") 
 
          

### 2024-10-22 scripts/03-clean_data.R

@@ -1,44 +1,66 @@
 #### Preamble ####
-# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 6 April 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Cleans the presidential poll data into an analysis data set
+# Author: Cristina Su Lam, Karen Riani, Mariko Lee
+# Date: 10 October 2024 
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: Run 01-download_data.R
+# Any other information needed? None
 
 #### Workspace setup ####
+library(readr)
 library(tidyverse)
+library(arrow)
+library(janitor)
+library(lubridate)
+library(broom)
 
 #### Clean data ####
-raw_data <- read_csv("inputs/data/plane_data.csv")
+clean_president_polls <- read_csv("data/01-raw_data/raw_president_polls.csv", 
+                                  col_types = cols(poll_id = col_skip(), 
+                                                   pollster_id = col_skip(), sponsor_ids = col_skip(), 
+                                                   sponsors = col_skip(), display_name = col_skip(), 
+                                                   pollster_rating_id = col_skip(), 
+                                                   pollster_rating_name = col_skip(), 
+                                                   pollscore = col_skip(), methodology = col_skip(), 
+                                                   transparency_score = col_skip(), 
+                                                   sponsor_candidate_id = col_skip(), 
+                                                   sponsor_candidate = col_skip(), sponsor_candidate_party = col_skip(), 
+                                                   endorsed_candidate_id = col_skip(), 
+                                                   endorsed_candidate_name = col_skip(), 
+                                                   endorsed_candidate_party = col_skip(), 
+                                                   question_id = col_skip(), population = col_skip(), 
+                                                   subpopulation = col_skip(), population_full = col_skip(), 
+                                                   tracking = col_skip(), created_at = col_skip(), 
+                                                   notes = col_skip(), url = col_skip(), 
+                                                   url_article = col_skip(), url_topline = col_skip(), 
+                                                   url_crosstab = col_skip(), source = col_skip(), 
[…]

### 2024-10-22 scripts/03-clean_data.R
Drop na.values(numerical_grade)

@@ -53,8 +53,9 @@ clean_president_polls <- clean_president_polls %>%
     is_state = as.numeric(!is.na(state)),
     end_date = mdy(end_date),
     start_date = mdy(start_date),
-    is_harris = ifelse(candidate_name == "Kamala Harris", 1, 0))
-
+    is_harris = ifelse(candidate_name == "Kamala Harris", 1, 0)) %>%
+  drop_na(sample_size) %>% 
+  drop_na(numeric_grade)
 
 # Filter for high quality pollsters and state-specific Harris and Trump polls
 clean_president_polls <- clean_president_polls %>%

### 2024-10-22 scripts/04-test_analysis_data.R
Tested Cleaned Dataset

@@ -1,69 +1,93 @@
 #### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 26 September 2024 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Some tests for Cleaned Dataset
+# Author: Cristina Su Lam, Karen Riani, Mariko Lee
+# Date: 10 October 2024 
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: Run 03-clean_data.R script
+# Any other information needed? None
 
 
 #### Workspace setup ####
 library(tidyverse)
 library(testthat)
+library(arrow)
 
-data <- read_csv("data/02-analysis_data/analysis_data.csv")
-
+data <- read_parquet("data/02-analysis_data/clean_president_polls.parquet")
 
 #### Test data ####
-# Test that the dataset has 151 rows - there are 151 divisions in Australia
-test_that("dataset has 151 rows", {
-  expect_equal(nrow(analysis_data), 151)
-})
 
-# Test that the dataset has 3 columns
-test_that("dataset has 3 columns", {
-  expect_equal(ncol(analysis_data), 3)
-})
+# Test 1: Ensure there are no missing values in important columns
+# Since state might be NA for national polls, we don’t include it in this test.
+stopifnot(!any(is.na(clean_president_polls$pollster)))
+stopifnot(!any(is.na(clean_president_polls$numeric_grade)))
+stopifnot(!any(is.na(clean_president_polls$candidate_name)))
+stopifnot(!any(is.na(clean_president_polls$pct)))
+stopifnot(!any(is.na(clean_president_polls$sample_size)))
 
-# Test that the 'division' column is character type
-test_that("'division' is character", {
-  expect_type(analysis_data$division, "character")
-})
+#### Test 2: Ensure numeric_grade values are within the expected range
+stopifnot(all(clean_president_polls$numeric_grade >= 0.5 & clean_president_polls$numeric_grade <= 3.0))
 
-# Test that the 'party' column is character type
-test_that("'party' is character", {
-  expect_type(analysis_data$party, "character")
-})
+#### Test 3: Ensure state values are either valid U.S. state names or NA (for national polls)
+expected_states <- c(state.name, NA)
[…]

### 2024-10-22 scripts/04_tests_and_checks/03-test_data.R
Delete scripts/04_tests_and_checks directory

@@ -1,21 +1,0 @@
-#### Preamble ####
-# Purpose: Sanity check of the data
-# Author: Cristina Su Lam, Karen Riani, Mariko Lee
-# Date: 10 October 2024 
-# License: MIT
-# Pre-requisites: Need to have simulated data
-# Any other information needed? None
-
-
-#### Workspace setup ####
-library(tidyverse)
-
-
-#### Test data ####
-data <- read_csv("data/raw_data/simulated.csv")
-
-# Test for negative numbers
-data$number_of_marriage |> min() <= 0
-
-# Test for NAs
-all(is.na(data$number_of_marriage))

### 2024-10-22 scripts/05-exploratory_data_analysis.R
initial model

@@ -1,8 +1,8 @@
 #### Preamble ####
 # Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Author: Cristina Su Lam, Karen Riani, Mariko Lee
+# Date: 
+# Contact: 
 # License: MIT
 # Pre-requisites: [...UPDATE THIS...]
 # Any other information needed? [...UPDATE THIS...]
@@ -11,27 +11,40 @@
 #### Workspace setup ####
 library(tidyverse)
 library(rstanarm)
+library(arrow)
+library(dplyr)
 
-#### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+#### Read upcoming presidential election forecast data ####
+upef2024_data <- 
+  read_parquet(file = here::here("data/02-analysis_data/clean_president_polls.parquet"))
 
-### Model data ####
-first_model <-
-  stan_glm(
-    formula = flying_time ~ length + width,
-    data = analysis_data,
-    family = gaussian(),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
+# Change 'pollster' and 'state' to factor variables
+upef2024_data <- upef2024_data |>
+  mutate(
+    pollster = factor(pollster),
+    state = factor(state)
   )
 
+### Multi-level Regression ###
+# Model 1
+model_formula_1 <- cbind(round((pct / 100) * sample_size, 0), sample_size - round((pct / 100) * sample_size, 0)) ~ (1 | pollster) + (1 | state)
+
+# Specify priors
+priors <- normal(0, 2.5, autoscale = TRUE)
+
+first_model <- stan_glmer(
+  formula = model_formula_1,
+  data = upef2024_data,
+  family = binomial(link = "logit"),
+  prior = priors,
+  prior_intercept = priors,
+  seed = 123,
+  cores = 4,
+  adapt_delta = 0.95
+)
 
 #### Save model ####
 saveRDS(
   first_model,
   file = "models/first_model.rds"
 )
-
-

### 2024-10-22 scripts/02_analysis_and_modelling/Bayesian_models.R
Delete old scripts/02_analysis_and_modelling directory

@@ -1,148 +1,0 @@
-#### Preamble ####
-# Purpose: !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
-# Author: Cristina Su Lam, Karen Riani, Mariko Lee
-# Date: 10 October 2024 
-# License: MIT
-# Pre-requisites: Need to have downloaded the data
-# Any other information needed? None
-
-#### Workspace setup ####
-library(tidyverse)
-library(janitor)
-library(lubridate)
-library(broom)
-library(modelsummary)
-library(rstanarm)
-library(splines)
-
-#### Read Data ####
-upef2024_data <- 
-  read_parquet(file = here::here("data/analysis_data/high_quality_harris.parquet"))
-
-#### Bayesian models ####
-# Change 'pollster' and 'state' to factor variables
-upef2024_data <- upef2024_data |>
-  mutate(
-    pollster = factor(pollster),
-    state = factor(state)
-  )
-
-### Multi-level Regression ###
-# Model 1
-model_formula_1 <- cbind(num_harris, sample_size - num_harris) ~ (1 | pollster)
-
-# Model 2
-model_formula_2 <- cbind(num_harris, sample_size - num_harris) ~ (1 | pollster) + (1 | state)
-
-# Specify priors
-priors <- normal(0, 2.5, autoscale = TRUE)
-
-# Fit the models
-bayesian_model_1 <- stan_glmer(
-  formula = model_formula_1,
-  data = upef2024_data,
-  family = binomial(link = "logit"),
-  prior = priors,
-  prior_intercept = priors,
-  seed = 123,
-  cores = 4,
-  adapt_delta = 0.95
-)
-
-bayesian_model_2 <- stan_glmer(
-  formula = model_formula_2,
-  data = upef2024_data,
-  family = binomial(link = "logit"),
-  prior = priors,
-  prior_intercept = priors,
-  seed = 123,
-  cores = 4,
-  adapt_delta = 0.95
-)
-# Note the warnings here and follow the instructions for dealing with it - come back and fix.
-
-# Posterior predictive checks
-pp_check(bayesian_model_1)
-pp_check(bayesian_model_2)
-
-# Summarize the model
-summary(bayesian_model_1)
-summary(bayesian_model_2)
-
-# Plot random effects
-plot(bayesian_model_1, pars = "(Intercept)", prob = 0.95)
-plot(bayesian_model_2, pars = "(Intercept)", prob = 0.95)
-
-# Model summary works the same as above for Bayesian models.
-
[…]

### 2024-10-22 scripts/02_analysis_and_modelling/initial_dataset_models.R

@@ -1,51 +1,0 @@
-#### Preamble ####
-# Purpose: !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
-# Author: Cristina Su Lam, Karen Riani, Mariko Lee
-# Date: 10 October 2024 
-# License: MIT
-# Pre-requisites: Need to have downloaded the data
-# Any other information needed? None
-
-#### Workspace setup ####
-library(tidyverse)
-library(janitor)
-library(lubridate)
-library(broom)
-library(modelsummary)
-
-#### Read Data ####
-high_quality_harris <- 
-  read_parquet(file = here::here("data/analysis_data/high_quality_harris.parquet"))
-
-#### Plot data ####
-base_plot <- ggplot(high_quality_harris, aes(x = end_date, y = pct)) +
-  theme_classic() +
-  labs(y = "Harris percent", x = "Date")
-
-# Plots poll estimates and overall smoothing
-base_plot +
-  geom_point() +
-  geom_smooth()
-
-# Color by pollster
-# This gets messy - need to add a filter - see line 21
-base_plot +
-  geom_point(aes(color = pollster)) +
-  geom_smooth() +
-  theme(legend.position = "bottom")
-
-# Facet by pollster
-# Make the line 21 issue obvious
-# Also - is there duplication???? Need to go back and check
-base_plot +
-  geom_point() +
-  geom_smooth() +
-  facet_wrap(vars(pollster))
-
-print(unique(high_quality_harris$pollster))
-
-# Color by pollscore
-base_plot +
-  geom_point(aes(color = factor(pollscore))) +
-  geom_smooth() +
-  theme(legend.position = "bottom")

### 2024-10-22 scripts/02_analysis_and_modelling/linear_models.R

@@ -1,203 +1,0 @@
-#### Preamble ####
-# Purpose: !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
-# Author: Cristina Su Lam, Karen Riani, Mariko Lee
-# Date: 10 October 2024 
-# License: MIT
-# Pre-requisites: Need to have downloaded the data
-# Any other information needed? None
-
-#### Workspace setup ####
-
-#### Plot data ####
-base_plot <- ggplot(high_quality_harris, aes(x = end_date, y = pct)) +
-  theme_classic() +
-  labs(y = "Harris percent", x = "Date")
-
-# Plots poll estimates and overall smoothing
-base_plot +
-  geom_point() +
-  geom_smooth()
-
-# Color by pollster
-# This gets messy - need to add a filter - see line 21
-base_plot +
-  geom_point(aes(color = pollster)) +
-  geom_smooth() +
-  theme(legend.position = "bottom")
-
-# Facet by pollster
-# Make the line 21 issue obvious
-# Also - is there duplication???? Need to go back and check
-base_plot +
-  geom_point() +
-  geom_smooth() +
-  facet_wrap(vars(pollster))
-
-# Color by pollscore
-base_plot +
-  geom_point(aes(color = factor(pollscore))) +
-  geom_smooth() +
-  theme(legend.position = "bottom")
-
-
-#### Starter models ####
-# Model 1: pct as a function of end_date
-model_date <- lm(pct ~ end_date, data = high_quality_harris)
-
-# Model 2: pct as a function of end_date and pollster
-model_date_pollster <- lm(pct ~ end_date + pollster, data = high_quality_harris)
-
-# Augment data with model predictions
-high_quality_harris <- high_quality_harris |>
-  mutate(
-    fitted_date = predict(model_date),
-    fitted_date_pollster = predict(model_date_pollster)
-  )
-
-# Plot model predictions
-# Model 1
-ggplot(high_quality_harris, aes(x = end_date)) +
-  geom_point(aes(y = pct), color = "black") +
-  geom_line(aes(y = fitted_date), color = "blue", linetype = "dotted") +
-  theme_classic() +
-  labs(y = "Harris percent", x = "Date", title = "Linear Model: pct ~ end_date")
-
-# Model 2
-ggplot(high_quality_harris, aes(x = end_date)) +
-  geom_point(aes(y = pct), color = "black") +
[…]

### 2024-10-22 paper/paper.qmd
Add Appendix A
Add Appendix B

@@
 
-data <- read_csv(here::here("data/raw_data/raw_data.csv"))
+data <- read_csv(here::here("data/01-raw_data/raw_president_polls.csv"))
 
@@
 
-data |> 
-  ggplot(aes(x = date, y = marriage_licenses)) +
-  geom_point()
+# data |> 
+#   ggplot(aes(x = date, y = marriage_licenses)) +
+#   geom_point()
 
@@
 ## Strengths and Weaknesses of the Methodology
-**Strength:**
-**Comprehensive Sampling Method:** The Washington Post can reach a broad demographic, including younger and more difficult-to-reach voters, by combining RDD, text-to-web polls, and live phone interviews [@WashPost2024_ABCMethodology] [@WashPost2023_Standards].
-**Post-stratification Weighting:** To account for demographic imbalances and increase the accuracy of their polls, The Washington Post and ABC News both use strong post-stratification weighting [@WashPost2024_ABCMethodology] [@ABCNews2024_Methodology].
+**Strength:** \
+**Comprehensive Sampling Method:** The Washington Post can reach a broad demographic, including younger and more difficult-to-reach voters, by combining RDD, text-to-web polls, and live phone interviews [@WashPost2024_ABCMethodology] [@WashPost2023_Standards].\
+**Post-stratification Weighting:** To account for demographic imbalances and increase the accuracy of their polls, The Washington Post and ABC News both use strong post-stratification weighting [@WashPost2024_ABCMethodology] [@ABCNews2024_Methodology].\
 **Transparent Approach:** The Washington Post’s polling data is more credible since they only employ high-quality polls in their averages and is transparent about their methodology [@WashPost2023_Standards] [@WashPost2024_PollingAverages].
-**Weaknesses:**
+\
+
+**Weaknesses:** \
 **Non-response Bias:** Even if both organizations use weighting adjustments, non-response bias still remains a challenge, particularly when it comes to groups that are less likely to respond to surveys [@WashPost2023_Standards] [@WashPost2024_ABCMethodology].
+\
 **Dependency on Historical Data:** In states with fewer polls, The Washington Post relies on historical data (the last two presidential elections), which raises the possibility that the polling averages might not accurately reflect current changes in voters preferences [@WashPost2024_PollingAverages].
+\
 
@@
 
+# Idealized Survey & Methodology - $100K Budget
+
+## Overview
+Using a $100K budget, this appendix outlines a carefully designed survey methodology for predicting the 2024 US Presidential Election.
+The objective is to collect representative, high-quality data using recruiting, poll aggregation, and selective sample methods.
+Through rigorous validation, this approach ensures data accuracy and reduces common survey research errors.
+
+## Sampling Approach
+We will implement stratified random sampling to ensure that key demographic and geographic subgroups are fairly represented.
+This approach reduces bias and offers more reliable insights into voter preference.
+
+**Stratification Criteria:**  
+- Age Groups  
+- Gender  
+- Education Levels  
+- Geographic Representation  
+- Political Affiliation
+
+**Sample Size Goal:** 10,000 respondents across states and demographics to achieve **high statistical power** with a margin of error below ±1%
+
+**Trade-offs:**  
+- Although stratified sampling increases representativeness, it necessitates accurate demographic information and may raise operating expenses.  
+- **Missing Data:** It’s possible that some demographics (e.g. men) may be less likely to respond. Post-stratification weighting and data imputation will be used to address this issue.
+
+## Recruitment Strategy
+Outline outreach and telephone surveys will be combined in our recruitment strategy to ensure widespread participation from various demographic groups.
+
+**Online Recruitment:**  
+- Target ads on Google, Facebook, and Twitter to engage younger voters and urban populations.  
+- Budget Allocation: $25,000
+
+**Random-Digit Dialing (RDD):**  
+- Phone outreach to reach older, rural voters with poor internet connection. 
+- Budget Allocation: $30,000
+
+**Incentives:**  
+- Participants are offered $5 gift cards to increase response rates. 
+- Budget Allocation: $20,000
+
+**Non-Response Handling:**  
+- Increase recruitment incentives for underrepresented groups and use numerous follow-up reminders.
+
+## Data Validation
+To ensure the accuracy and reliability of responses, we will implement several data validation techniques:
+
+**Survey Logic Check:**  
+- Recognize and flag responses contradicting one another (e.g., reporting under 18 but registered to vote).  
+**Attention Check:**  
+- Utilize questions to confirm respondents are actively engaged (e.g., “Select ‘Confirm’ to start questionnaire”)  
+**Post-Stratification Weighting: **  
+- Adjusting for over- and under-enumeration and weighting the sample to reflect the demographic of the US population.
+
+**Mode and Measurement Errors:**   
+- We mitigate the impact of using mixed modes (online and telephones) by training enumerators and reducing enumerator bias. Misreporting will be reduced through straightforward questions.
+
+## Poll Aggregation Methodology
+We will employ a poll-of-polls aggregation method to reduce bias and smooth fluctuation across individual polls.
+
+**Weighting Criteria:**  
+**Sample Size:** larger samples receive more weight to mirror greater reliability.
+
+**Recency:** More recent polls are given higher weight to capture modern voter sentiment.
+
+**Pollster Rating:** Polls from highly rated pollsters receive higher weights to reduce the impact of bias.
+
+## Survey Implementation 
+Google Forms was used to create and implement the survey, allowing for efficient data collection and safe storage. The main section and sample questions are listed below.
+
+Access the survey: [Google Form](https://forms.gle/F1v3gswDjra8YLP79)
+
+**Survey Overview** \
+**Title:** 2024 US Presidential Election Poll \
+**Purpose:** To gather public sentiment and predict election outcomes. \
+**Estimated Time:** Less than 5 minutes \
+**Confidentiality:** All responses are anonymous and used only for research purposes.
+
+1. What is your age?
+- Under 18
+- 18-24
+- 25-34
+- 35-44
+- 45-54
+- 55-64
+- 65+
+
+2. What is your gender?
+- Male
+- Female
+- Non-binary / Prefer not to say
+
+3. What is the highest level of education you have completed?
+- Less than high school
+- High school diploma or GED
+- Some college
+- Bachelor’s degree
+- Master’s degree or higher
+
+4. Are you registered to vote?
+- Yes
+- No
+
+5. Who do you intend to vote for in the upcoming presidential election?
+- Donald Trump (Republican)
+- Kamala Harris (Democrat)
+- Other
+- Undecided
+
+6. How likely are you to vote in the upcoming election?
+- Very Likely
+- Somewhat Likely
+- Not Likely
+
+## Budget Breakdown
+- **Online Recruitment** : $25,000
+- **RDD Recruitment**: $30,000
+- **Incentives for Participants**: $20,000
+- **Data Processing & Validation**: $15,000
+- **Miscellaneous Expenses**: $10,000
+
+**Total**: $100,000
+
+## Conclusion
+This survey methodology uses stratified sampling, multi-channel recruitment, and rigorous data validation procedures to ensure accurate forecasting of the 2024 US Presidential Election.
+We provide a more stable and reliable prediction through poll-polls aggregation, smoothing out fluctuations across polls.
+This design balances accuracy, inclusivity, and efficiency with a carefully considered $100K budget, ensuring the poll gathers meaningful insights into voter sentiment and behavior.
+
 \newpage

### 2024-10-22 scripts/02-download_data.R
change download from local to web

@@ -10,7 +10,7 @@
 library(readr)
 
 #### Download/Read Dataset ####
-raw_president_polls <- read_csv("data/01-raw_data/raw_president_polls.csv")
+raw_president_polls <- read_csv("https://projects.fivethirtyeight.com/polls/data/president_polls.csv")
 
-
-         
+#### Save Data ####
+write_csv(raw_president_polls, "data/01-raw_data/raw_president_polls.csv") 

### 2024-10-22 scripts/02-download_data.R
Added Some Tests

@@ -10,7 +10,7 @@
 library(readr)
 
 #### Download/Read Dataset ####
-raw_president_polls <- read_csv("data/01-raw_data/raw_president_polls.csv")
+raw_president_polls <- read_csv("data/01-raw_data/raw_president_polls_oct22.csv")
 
 
          

### 2024-10-22 scripts/03-clean_data.R

@@ -15,7 +15,7 @@ library(lubridate)
 library(broom)
 
 #### Clean data ####
-clean_president_polls <- read_csv("data/01-raw_data/raw_president_polls.csv", 
+clean_president_polls <- read_csv("data/01-raw_data/raw_president_polls_oct22.csv", 
                                   col_types = cols(poll_id = col_skip(), 
                                                    pollster_id = col_skip(), sponsor_ids = col_skip(), 
                                                    sponsors = col_skip(), display_name = col_skip(), 
@@ -55,12 +55,14 @@ clean_president_polls <- clean_president_polls %>%
     start_date = mdy(start_date),
     is_harris = ifelse(candidate_name == "Kamala Harris", 1, 0)) %>%
   drop_na(sample_size) %>% 
-  drop_na(numeric_grade)
+  drop_na(numeric_grade) %>%
+  drop_na(state)
 
 # Filter for high quality pollsters and state-specific Harris and Trump polls
 clean_president_polls <- clean_president_polls %>%
   filter(
-    candidate_name %in% c("Kamala Harris", "Donald Trump")
+    candidate_name %in% c("Kamala Harris", "Donald Trump"),
+    start_date >= ymd("2024-07-21")
   ) 
 
 #### Save data ####

### 2024-10-22 scripts/04-test_analysis_data.R

@@ -12,23 +12,23 @@ library(tidyverse)
 library(testthat)
 library(arrow)
 
-data <- read_parquet("data/02-analysis_data/clean_president_polls.parquet")
+clean_president_polls <- read_parquet("data/02-analysis_data/clean_president_polls.parquet")
 
 #### Test data ####
 
 # Test 1: Ensure there are no missing values in important columns
-# Since state might be NA for national polls, we don’t include it in this test.
 stopifnot(!any(is.na(clean_president_polls$pollster)))
 stopifnot(!any(is.na(clean_president_polls$numeric_grade)))
 stopifnot(!any(is.na(clean_president_polls$candidate_name)))
 stopifnot(!any(is.na(clean_president_polls$pct)))
 stopifnot(!any(is.na(clean_president_polls$sample_size)))
+stopifnot(!any(is.na(clean_president_polls$state)))
 
 #### Test 2: Ensure numeric_grade values are within the expected range
 stopifnot(all(clean_president_polls$numeric_grade >= 0.5 & clean_president_polls$numeric_grade <= 3.0))
 
 #### Test 3: Ensure state values are either valid U.S. state names or NA (for national polls)
-expected_states <- c(state.name, NA)
+expected_states <- c(state.name)
 stopifnot(all(clean_president_polls$state %in% expected_states))
 
 #### Test 4: Ensure candidate_name contains only "Kamala Harris" and "Donald Trump"

### 2024-10-23 scripts/05-exploratory_data_analysis.R
Added a map but need adjustments

@@ -1,50 +1,112 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
+# Purpose: Exploratory Data Analysis for Cleaned Dataset
 # Author: Cristina Su Lam, Karen Riani, Mariko Lee
-# Date: 
-# Contact: 
+# Date: 10 October 2024 
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
+# Pre-requisites: Run 03-clean_data.R script
+# Any other information needed? None
 
 #### Workspace setup ####
 library(tidyverse)
 library(rstanarm)
 library(arrow)
 library(dplyr)
+library(maps)
 
 #### Read upcoming presidential election forecast data ####
-upef2024_data <- 
-  read_parquet(file = here::here("data/02-analysis_data/clean_president_polls.parquet"))
+clean_president_polls <- read_parquet("data/02-analysis_data/clean_president_polls.parquet")
 
-# Change 'pollster' and 'state' to factor variables
-upef2024_data <- upef2024_data |>
-  mutate(
-    pollster = factor(pollster),
-    state = factor(state)
+# Summary statistics for pct and sample_size
+summary_stats <- clean_president_polls %>%
+  summarize(
+    mean_pct = mean(pct, na.rm = TRUE),
+    sd_pct = sd(pct, na.rm = TRUE),
+    min_pct = min(pct, na.rm = TRUE),
+    max_pct = max(pct, na.rm = TRUE),
+    mean_sample_size = mean(sample_size, na.rm = TRUE),
+    sd_sample_size = sd(sample_size, na.rm = TRUE),
+    min_sample_size = min(sample_size, na.rm = TRUE),
+    max_sample_size = max(sample_size, na.rm = TRUE)
   )
 
-### Multi-level Regression ###
-# Model 1
-model_formula_1 <- cbind(round((pct / 100) * sample_size, 0), sample_size - round((pct / 100) * sample_size, 0)) ~ (1 | pollster) + (1 | state)
-
-# Specify priors
-priors <- normal(0, 2.5, autoscale = TRUE)
-
-first_model <- stan_glmer(
-  formula = model_formula_1,
-  data = upef2024_data,
-  family = binomial(link = "logit"),
-  prior = priors,
-  prior_intercept = priors,
-  seed = 123,
-  cores = 4,
-  adapt_delta = 0.95
+# Create a summary table in the desired format
+summary_table <- data.frame(
[…]

### 2024-10-24 scripts/02-download_data.R
Added some models

@@ -8,9 +8,11 @@
 
 #### Workspace setup ####
 library(readr)
+library(kagglehub)
 
 #### Download/Read Dataset ####
-raw_president_polls <- read_csv("data/01-raw_data/raw_president_polls_oct22.csv")
+raw_president_polls <- read_csv("https://projects.fivethirtyeight.com/polls/data/president_polls.csv")
 
+#### Save Data ####
+write_csv(raw_president_polls, "data/01-raw_data/raw_president_polls.csv") 
 
-         

### 2024-10-24 scripts/06-model_data.R

@@ -1,37 +1,65 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Models
+# Author: Cristina Su Lam, Karen Riani, Mariko Lee
+# Date: 23 October 2024
 # License: MIT
 # Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Any other information needed? None
 
 
 #### Workspace setup ####
 library(tidyverse)
+library(arrow)
+library(dplyr)
 library(rstanarm)
 
-#### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+#### Read upcoming presidential election forecast data ####
+clean_president_polls <- read_parquet("data/02-analysis_data/clean_president_polls.parquet")
 
-### Model data ####
-first_model <-
-  stan_glm(
-    formula = flying_time ~ length + width,
-    data = analysis_data,
-    family = gaussian(),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
-  )
+### Logistic Regression ###
 
+# Normalize weights to a reasonable scale
+clean_president_polls <- clean_president_polls %>%
+  mutate(weight = sample_size / mean(sample_size)) %>%
+  mutate(weight = weight * numeric_grade)
+
+model_logistic <- glm(
+  is_harris ~ pollster + state + sample_size + pct,
+  data = clean_president_polls,
+  family = binomial(link = "logit"),
+  weights = weight
+)
+model_logistic
+
+# Predict the probabilities of Harris winning
+clean_president_polls <- clean_president_polls %>%
+  mutate(predicted_prob_harris = predict(model_logistic, type = "response"))
+clean_president_polls
+
+
+### Bayesian Model ###
+
+# Specify priors 
+priors <- normal(0, 2.5, autoscale = TRUE)
+
+# Model 
+model_formula <- is_harris ~ sample_size + pct + (1 | pollster) + (1 | state)
+
+# Fit the model
[…]

### 2024-10-24 scripts/06-model_data.R
added model summary, pp checks

@@ -58,6 +58,15 @@ bayesian_model <- stan_glmer(
   adapt_delta = 0.95
 )
 
+# Posterior predictive checks
+pp_check(bayesian_model)
+
+# Summarize the model
+summary(bayesian_model)
+
+# Plot random effects
+plot(bayesian_model, pars = "(Intercept)", prob = 0.95)
+
 #### Save model ####
 saveRDS(
   first_model,

### 2024-10-25 scripts/06-model_data.R
Added Model Validation for 1st Model

@@ -12,13 +12,18 @@ library(tidyverse)
 library(arrow)
 library(dplyr)
 library(rstanarm)
+library(pROC)
 
 #### Read upcoming presidential election forecast data ####
 clean_president_polls <- read_parquet("data/02-analysis_data/clean_president_polls.parquet")
 
+# Convert categorical variables to factors
+clean_president_polls$pollster <- as.factor(clean_president_polls$pollster)
+clean_president_polls$state <- as.factor(clean_president_polls$state)
+
 ### Logistic Regression ###
 
-# Normalize weights to a reasonable scale
+# Weights
 clean_president_polls <- clean_president_polls %>%
   mutate(weight = sample_size / mean(sample_size)) %>%
   mutate(weight = weight * numeric_grade)
@@ -34,8 +39,45 @@ model_logistic
 # Predict the probabilities of Harris winning
 clean_president_polls <- clean_president_polls %>%
   mutate(predicted_prob_harris = predict(model_logistic, type = "response"))
-clean_president_polls
 
+
+# Model Validation & Checking
+set.seed(123)
+train_indices <- sample(seq_len(nrow(clean_president_polls)), size = 0.7 * nrow(clean_president_polls))
+
+# Create training and test datasets
+train_data <- clean_president_polls[train_indices, ]
+test_data <- clean_president_polls[-train_indices, ]
+
+model_logistic_train <- glm(
+  is_harris ~ pollster + state + sample_size + pct + pollster:pct,
+  data = train_data,
+  family = binomial(link = "logit"),
+  weights = weight
+)
+
+
+# Remove unused levels from the model object
+model_logistic_clean <- model_logistic
+model_logistic_clean$xlevels$pollster <- intersect(levels(test_data$pollster), levels(train_data$pollster))
+model_logistic_clean$xlevels$state <- intersect(levels(test_data$state), levels(train_data$state))
+
+# Run the prediction
+test_data <- test_data %>%
+  mutate(predicted_prob_harris = predict(model_logistic_clean, newdata = test_data, type = "response"))
+
+# Create a binary prediction (Harris wins if probability > 0.5)
+test_data <- test_data %>%
+  mutate(prediction = ifelse(predicted_prob_harris > 0.5, 1, 0))
+
+# Calculate accuracy
+accuracy <- mean(test_data$prediction == test_data$is_harris)
+print(accuracy)
+
+# Calculate AUC
[…]

### 2024-10-28 scripts/06-model_data.R
Added Model Validation + Checking

@@ -14,6 +14,7 @@ library(dplyr)
 library(rstanarm)
 library(pROC)
 
+
 #### Read upcoming presidential election forecast data ####
 clean_president_polls <- read_parquet("data/02-analysis_data/clean_president_polls.parquet")
 
@@ -21,34 +22,37 @@ clean_president_polls <- read_parquet("data/02-analysis_data/clean_president_pol
 clean_president_polls$pollster <- as.factor(clean_president_polls$pollster)
 clean_president_polls$state <- as.factor(clean_president_polls$state)
 
-### Logistic Regression ###
+# Set seed for reproducibility
+set.seed(123)
 
-# Weights
+### Step 1: Prepare Data and Define Initial Model ###
+
+# Add weights to the dataset
 clean_president_polls <- clean_president_polls %>%
-  mutate(weight = sample_size / mean(sample_size)) %>%
-  mutate(weight = weight * numeric_grade)
+  mutate(weight = (sample_size / mean(sample_size)) * numeric_grade)
 
-model_logistic <- glm(
+# Fit the initial logistic regression model
+logistic_model <- glm(
   is_harris ~ pollster + state + sample_size + pct,
   data = clean_president_polls,
   family = binomial(link = "logit"),
   weights = weight
 )
-model_logistic
 
-# Predict the probabilities of Harris winning
+# Predict the probabilities of Harris winning for the entire dataset
 clean_president_polls <- clean_president_polls %>%
-  mutate(predicted_prob_harris = predict(model_logistic, type = "response"))
+  mutate(predicted_prob_harris = predict(logistic_model, type = "response"))
 
+### Step 2: Train/Test Split for Model Validation ###
 
-# Model Validation & Checking
-set.seed(123)
+# Create training and test datasets (70% train, 30% test)
 train_indices <- sample(seq_len(nrow(clean_president_polls)), size = 0.7 * nrow(clean_president_polls))
-
-# Create training and test datasets
 train_data <- clean_president_polls[train_indices, ]
 test_data <- clean_president_polls[-train_indices, ]
 
+### Step 3: Fit Logistic Regression Model on Training Set ###
+
+# Fit a logistic regression model with interaction term on the training data
 model_logistic_train <- glm(
   is_harris ~ pollster + state + sample_size + pct + pollster:pct,
   data = train_data,
[…]

### 2024-10-28 scripts/06-model_data.R
Got overall result prediction

@@ -43,6 +43,35 @@ logistic_model <- glm(
 clean_president_polls <- clean_president_polls %>%
   mutate(predicted_prob_harris = predict(logistic_model, type = "response"))
 
+# Calculate the unweighted average predicted probability for Kamala Harris
+overall_predicted_prob_harris <- mean(clean_president_polls$predicted_prob_harris)
+
+# Convert to percentage
+overall_percentage_harris <- overall_predicted_prob_harris * 100
+overall_percentage_trump <- (1 - overall_predicted_prob_harris) * 100
+
+# Print the results
+cat("Overall Percentage for Kamala Harris:", overall_percentage_harris, "%\n")
+cat("Overall Percentage for Donald Trump:", overall_percentage_trump, "%\n")
+
+
+# Calculate the average predicted probability by state
+state_predictions <- clean_president_polls %>%
+  group_by(state) %>%
+  summarize(avg_predicted_prob_harris = mean(predicted_prob_harris))
+
+# Determine the winner for each state based on the average probability
+state_predictions <- state_predictions %>%
+  mutate(state_winner = ifelse(avg_predicted_prob_harris > 0.5, "Harris", "Trump"))
+
+# Count the number of states won by each candidate
+overall_winner_summary <- state_predictions %>%
+  count(state_winner)
+
+# View results
+state_predictions    
+overall_winner_summary 
+
 ### Step 2: Train/Test Split for Model Validation ###
 
 # Create training and test datasets (70% train, 30% test)

### 2024-10-28 scripts/05-exploratory_data_analysis.R
fixed state map

@@ -10,7 +10,6 @@
 library(tidyverse)
 library(rstanarm)
 library(arrow)
-library(dplyr)
 library(maps)
 
 #### Read upcoming presidential election forecast data ####
@@ -96,17 +95,18 @@ state_winners$state <- tolower(state_winners$state)
 # Load USA state map data
 usa_states <- map_data("state")
 
-# Merge the winners with the USA map data
-merged_data <- merge(usa_states, state_winners, by.x = "region", by.y = "state", all.x = TRUE)
+# Merge and sort the winners with the USA map data
+merged_data <- merge(usa_states, state_winners, by.x = "region", by.y = "state", sort = FALSE, all.x = TRUE)
+merged_data <- merged_data[order(merged_data$group, merged_data$order), ]
 
 # Plot the election map
 ggplot(data = merged_data, aes(x = long, y = lat, group = group, fill = winner)) +
-  geom_polygon(color = "black", size = 0.3) +  # Add state borders
+  geom_polygon(color = "black", size = 0.3) + 
+  labs(fill = "Winner") +
   scale_fill_manual(values = c("Kamala Harris" = "blue", "Donald Trump" = "red")) +  # Blue for Harris, Red for Trump
-  coord_fixed(1.3) +  # Fix the aspect ratio for a proper USA map
-  labs(title = "2024 Election Results by State", x = "Longitude", y = "Latitude") +
-  theme_void() +  # Remove gridlines and axes for a cleaner map
-  theme(legend.position = "none")  # Remove the legend if you don't want it
+  coord_fixed(1.3) + 
+  labs(title = "Predicted 2024 Election Results by State", x = "Longitude", y = "Latitude") +
+  theme_void() 
 
 
 

### 2024-10-28 scripts/05-exploratory_data_analysis.R
fixed Harris vs non-Harris plot

@@ -77,9 +77,9 @@ ggplot(clean_president_polls, aes(x = sample_size)) +
   scale_x_continuous(labels = scales::comma)
 
 # 5. Plot for is_harris variable
-ggplot(clean_president_polls_filtered, aes(x = factor(is_harris))) +
+ggplot(clean_president_polls, aes(x = factor(is_harris))) +
   geom_bar(fill = "purple", alpha = 0.7) +
-  labs(title = "Distribution of Polls by Harris vs Non-Harris", x = "Is Harris", y = "Number of Polls") +
+  labs(title = "Distribution of Polls by Harris vs Trump", x = "Is Harris", y = "Number of Polls") +
   scale_x_discrete(labels = c("0" = "Trump", "1" = "Harris"))
 
 # 6. Map

### 2024-10-29 scripts/06-model_data.R
preamble updates

@@ -1,16 +1,15 @@
 #### Preamble ####
-# Purpose: Models
+# Purpose: Fitting Model and Making Predictions
 # Author: Cristina Su Lam, Karen Riani, Mariko Lee
 # Date: 23 October 2024
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
+# Pre-requisites: Run 03-clean_data.R script
 # Any other information needed? None
 
 
 #### Workspace setup ####
 library(tidyverse)
 library(arrow)
-library(dplyr)
 library(rstanarm)
 library(pROC)
 

### 2024-10-29 paper/paper.qmd
initial abstract

@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "We forecast the 2024 U.S.
+presidential winner using state-level poll averages for Kamala Harris and Donald Trump, weighted by poll quality and sample size.
+Applying a logistic regression model, we find a slight 0.8% lead for Trump over Harris, using polls taken since the reelection bid on July 21, 2024.
+Weighting the polls increases the influence of the most reliable data in shaping the forecast—limiting skewed results and showing a clearer voter sentiment amidst a particularly competitive election.
+These results show who may be leading the elections via polls, giving campaigns the insights they need and reinforcing public trust in the election process."
 format: pdf
@@
 
-You can and should cross-reference sections and sub-sections. We use @talia, @opendatatorotno, and @rohan.
+The following structure of this paper is as follows:  @sec-mydatasection presents the data sources and methodology, then  @sec-myresults followed by  @sec-mydisc, which discusses the results and their implications.
+@sec-mylimits concludes with a discussion of the limitations of the study and suggestions for future research.
 
-The remainder of this paper is structured as follows. @sec-mydatasection
-
-# Data {#sec-data}
+# Data {#sec-mydatasection}
 
@@
 
-# data |> 
-#   ggplot(aes(x = date, y = marriage_licenses)) +
-#   geom_point()
-
 ```
 
-Talk more about it.
+# Results {#sec-myresults}
 
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
-
-# Discussion
+# Discussion {#sec-mydisc}
 

### 2024-10-31 scripts/05-exploratory_data_analysis.R
a

@@ -78,7 +78,7 @@ ggplot(clean_president_polls, aes(x = sample_size)) +
   scale_x_continuous(labels = scales::comma)
 
 # 5. Plot for is_harris variable
-ggplot(clean_president_polls_filtered, aes(x = factor(is_harris))) +
+ggplot(clean_president_polls, aes(x = factor(is_harris))) +
   geom_bar(fill = "purple", alpha = 0.7) +
   labs(title = "Distribution of Polls by Harris vs Non-Harris", x = "Is Harris", y = "Number of Polls") +
   scale_x_discrete(labels = c("0" = "Trump", "1" = "Harris"))

### 2024-10-31 scripts/02-download_data.R
updated dataset

@@ -8,7 +8,6 @@
 
 #### Workspace setup ####
 library(readr)
-library(kagglehub)
 
 #### Download/Read Dataset ####
 raw_president_polls <- read_csv("https://projects.fivethirtyeight.com/polls/data/president_polls.csv")

### 2024-10-31 scripts/03-clean_data.R

@@ -15,7 +15,7 @@ library(lubridate)
 library(broom)
 
 #### Clean data ####
-clean_president_polls <- read_csv("data/01-raw_data/raw_president_polls_oct22.csv", 
+clean_president_polls <- read_csv("data/01-raw_data/raw_president_polls.csv", 
                                   col_types = cols(poll_id = col_skip(), 
                                                    pollster_id = col_skip(), sponsor_ids = col_skip(), 
                                                    sponsors = col_skip(), display_name = col_skip(), 
@@ -49,7 +49,9 @@ clean_president_polls <- read_csv("data/01-raw_data/raw_president_polls_oct22.cs
 clean_president_polls <- clean_president_polls %>%
   mutate(
     state = str_replace(state, "Maine CD-[12]", "Maine"),
+    state = str_replace(state, "Nebraska CD-1", "Nebraska"),
     state = str_replace(state, "Nebraska CD-2", "Nebraska"),
+    state = str_replace(state, "Nebraska CD-3", "Nebraska"),
     end_date = mdy(end_date),
     start_date = mdy(start_date),
     is_harris = ifelse(candidate_name == "Kamala Harris", 1, 0)) %>%

### 2024-10-31 paper/paper.qmd
intro ref + paper

@@
 title: "2024 US Presidential Election Forecasting"
-subtitle: "subtitle"
+subtitle: "Trump forcasted to win popular vote by state over Harris by 0.8%"
 author: 
@@
   - Cristina Su Lam
-thanks: "Code and data are available at: https://projects.fivethirtyeight.com/polls/president-general/2024/national/"
+thanks: "Code and data are available at: https://github.com/karenrni/Forecasting-the-2024-US-Presidential-Elections"
 date: today
@@
 library(tidyverse)
+library(arrow)
 
-data <- read_csv(here::here("data/01-raw_data/raw_president_polls.csv"))
+#data <- read_csv(here::here("data/01-raw_data/raw_president_polls.csv"))
+# cleaned_data <- read_parquet("data/02-analysis_data/clean_president_polls.parquet")
 
@@
 # Introduction
+
+Amid the chaotic, high-stakes 2024 U.S.
+presidential election, forecasting models offer crucial insights into voter sentiment as Americans face a tight race between top candidates Kamala Harris and Donald Trump.
+With the attempted assassination of Trump, Joe Biden’s endorsement of Harris, and recent crises such as hurricanes disrupting voter access in swing states, public opinion remains volatile [@CBSNews2024_CampaignPromises; @ABCNews2024_HurricaneImpact].
+Key events like Trump’s policy proposals on social security and public safety, Harris’s advocacy for reproductive rights, and other rising debates on gun control have further polarized the electorate, making forecasting models essential in offering clarity amid this competitive landscape [@HRW2022_RoeVWade].
+
+This study joins the growing body of forecast models—including media outlets’ own polls-of-polls approaches and various Bayesian methods—by refining the aggregation of poll data with logistic regression, weighted by sample size and pollster rating [@NYTimes2024_ElectionPolls; @AP2024_ElectionForecasts].
+In an environment where neither candidate shows a decisive lead, this model aims to add precision by filtering out lower-quality polls, improving forecast reliability for strategic decision-making and public confidence in election projections [@Bloomberg2024_KeyStates].
+
+We estimate a slim 0.8% lead for Trump over Harris in the overall popular vote by aggregating state-level polling data, a close margin that highlights the importance of updated, representative data amidst competing forecasts and fluctuating polls.
+By using polls weighted by reliability and focusing on data collected after Harris' campaign declaration, this model maintains transparency, offering a robust perspective on public sentiment and helping to clarify perceptions in a highly competitive election.
 
@@
 
-## Weaknesses and next steps
+## Weaknesses and next steps {#sec-mylimits}
 
@@
 The Washington Post uses response weighting, which modifies the results according to demographic variables such as age, race, and education, in order to address non-response bias.
-In spite of variations in response rates among demographic groupings, this ensures that the final sample more accurately represents the population [@WashPost2023_Standard].
+In spite of variations in response rates among demographic groupings, this ensures that the final sample more accurately represents the population [@WashPost2023_Standards].
 

### 2024-11-01 scripts/06-model_data.R
grammar fix

@@ -90,7 +90,7 @@ model_logistic_train <- glm(
 )
 
 # Remove unused levels from the model object
-model_logistic_clean <- model_logistic
+model_logistic_clean <- model_logistic_train
 model_logistic_clean$xlevels$pollster <- intersect(levels(test_data$pollster), levels(train_data$pollster))
 model_logistic_clean$xlevels$state <- intersect(levels(test_data$state), levels(train_data$state))
 
@@ -217,7 +217,7 @@ auc_value
 
 #### Save model ####
 saveRDS(
-  model_logistic,
+  model_logistic_clean,
   file = "models/first_model.rds"
 )
 

### 2024-11-03 scripts/06-model_data.R
!

@@ -138,8 +138,6 @@ bayesian_model <- stan_glmer(
   adapt_delta = 0.95
 )
 
-
-
 # Predict probabilities for Harris winning across all posterior draws
 predicted_probs_matrix <- posterior_predict(bayesian_model, type = "response")
 

### 2024-11-03 paper/paper.qmd

@@
 presidential winner using state-level poll averages for Kamala Harris and Donald Trump, weighted by poll quality and sample size.
-Applying a logistic regression model, we find a slight 0.8% lead for Trump over Harris, using polls taken since the reelection bid on July 21, 2024.
+Applying a Bayesian  model, we find a slight 0.5% lead for Trump over Harris, using polls taken since the reelection bid on July 21, 2024.
 Weighting the polls increases the influence of the most reliable data in shaping the forecast—limiting skewed results and showing a clearer voter sentiment amidst a particularly competitive election.
@@
 
-This study joins the growing body of forecast models—including media outlets’ own polls-of-polls approaches and various Bayesian methods—by refining the aggregation of poll data with logistic regression, weighted by sample size and pollster rating [@NYTimes2024_ElectionPolls; @AP2024_ElectionForecasts].
-In an environment where neither candidate shows a decisive lead, this model aims to add precision by filtering out lower-quality polls, improving forecast reliability for strategic decision-making and public confidence in election projections [@Bloomberg2024_KeyStates].
+This study joins the growing body of forecast models—including media outlets’ own polls-of-polls approaches and various Bayesian methods—by refining the aggregation of poll data with Bayesian regression, weighted by sample size and pollster rating [@NYTimes2024_ElectionPolls; @AP2024_ElectionForecasts].
+In an environment where neither candidate shows a decisive lead, this model aims to add precision by emphasizing out higher-quality polls, improving forecast reliability for strategic decision-making and public confidence in election projections [@Bloomberg2024_KeyStates].
 
-We estimate a slim 0.8% lead for Trump over Harris in the overall popular vote by aggregating state-level polling data, a close margin that highlights the importance of updated, representative data amidst competing forecasts and fluctuating polls.
+We estimate a slim 0.5% lead for Trump over Harris in the overall popular vote by aggregating state-level polling data, a close margin that highlights the importance of updated, representative data amidst competing forecasts and fluctuating polls.
 By using polls weighted by reliability and focusing on data collected after Harris' campaign declaration, this model maintains transparency, offering a robust perspective on public sentiment and helping to clarify perceptions in a highly competitive election.
 
-The following structure of this paper is as follows:  @sec-mydatasection presents the data sources and methodology, then  @sec-myresults followed by  @sec-mydisc, which discusses the results and their implications.
+The following structure of this paper is as follows:  @sec-mydatasection presents the data sources and methodology, @sec-mymodel presents the forecasting model, then  @sec-myresults followed by  @sec-mydisc, which discusses the results and their implications.
 @sec-mylimits concludes with a discussion of the limitations of the study and suggestions for future research.
@@
 # Data {#sec-mydatasection}
+
+TODO: PROPER FORMATTING FOR DATA SEC, ADD FINAL DETAILS AND FIGURES
+
+## Overview
+Our study uses state-level polling data from Donald Trump and Kamala Harris to predict the 2024 US presidential election.
+The polls used were sourced from FiveThirtyEight’s 2024 Presidential Election Forecast Database (FiveThirtyEight 2024), which compiles polling data from multiple organizations and rates each poll’s quality.
+These ratings assist in identifying and weighing the most reliable polls.
+The dataset used for this analysis spans polls starting from July 21, 2024, when Kamala Harris announced her reelection bid.
+
+Our analysis focuses on the following variables: 
+Polling rating (Pollscore): A numerical score reflecting the historical reliability of polling organizations used to weight polls by accuracy. 
+Sample size: The number of respondents per poll, influencing the poll’s margin of error and its weight in the forecast. 
+Percentage support (pct): The proportion of respondents who express support for each candidate. 
+State: The US state where the poll was conducted OR labeled “National” for nationwide polls. 
+Pollster: The organization that conducted the poll, influencing the credibility and weight of our analysis.
+
+Table X - Sample of the cleaned dataset, highlighting key variables crucial in the analysis. 
+[Table]
+
+### Measurement and Limitations
+We applied several measurement and limitations considerations to ensure the validity of our dataset:
+Poll Quality: 
+Temporal Dynamics:
+Geographic Coverage:
+Response Bias:
+
+### Outcome Variables
+The Percentage of support (pct) for Donald Trump and Kamala Harris is the main outcome variable.
+We forecast which candidate is leading by aggregating these results across states.
+Figure X represents the distribution of support percentages for the two candidates, showing clusters ranging from 40% to 60%, indicating the election’s competitive nature.
+
+[Figure X: Distribution of Candidate Support]
+
+### Predictor Variables
+The following predictor variables were used to build the logistic regression model and refine our prediction:
+Pollscore: Historical reliability score for pollster (range: )
+Numeric Grade: (scale: 0-4) 
+Transparency Score: (scale: 0-10) 
+Sample Size: Number of respondents in each poll (typically between 500-3000) 
+Methodology: Survey method (eg. phone interviews, online panels) 
+State: State-level indicators that capture regional variations in support.
+
+These selected variables are based on their relevance to polling accuracy and availability across the dataset.
+The analysis regarding their relationships with the outcome variables (percentage support) is further explored.
+
+### Cleaning Process and Analysis
+The data cleaning process involved the following steps:
+Filtering for High-Quality Polls: We included only polls with a numeric grade of X or higher to reduce the impact of unreliable data. 
+Standardizing Dates and Locations: 
+Variable Transformation:
+Converted percentages into absolute numbers using sample size 
+Created binary indicators for candidates (Trump = 0, Harris = 1)
 
@@
 ```
+# Model {#sec-mymodel}
+
+TODO: FINALIZE MODEL SECTION
+- VALIDATION
+- ALTERNATIVES
+- JUSTIFICATION
+- UPDATE MODEL BELOW
+
+$$
+\begin{aligned}
+\text{logit}(\mu_i) &= \beta_0 + \beta_1 \times \text{Pollster}_i + \beta_2 \times \text{State}_i + \beta_3 \times \text{Sample Size}_i + \beta_4 \times \text{Pct}_i
+\end{aligned}
+$$
+- $y_i$ is the dependent variable, representing the count of respondents who support Harris in a given poll, modeled as a binomial outcome.
+- $\beta_0$ is the intercept term, indicating the baseline log-odds of Harris support when all predictors are zero.
+- $\beta_1$, $\beta_2$, $\beta_3$, and $\beta_4$ are the coefficients for the predictor variables **Pollster**, **State**, **Sample Size**, and **Pct** (the percentage of support for Harris in a poll), respectively:
+  - $\beta_1$ represents the adjustment in log-odds based on the specific pollster conducting the survey.
+  - $\beta_2$ accounts for the impact of the state in which the poll is conducted.
+  - $\beta_3$ adjusts for the influence of the poll's sample size.
+  - $\beta_4$ represents the effect of the poll's percentage support for Harris on the log-odds.
+- The function $\text{logit}(\mu_i)$ transforms the linear combination of predictors into probabilities, providing the estimated probability of support for Harris.
 
@@
 
+TODO: FINISH RESULTS INCLUDING GRAPHS, TABLES, SUMMARY STATS
+
+Overall Percentage for Kamala Harris: 49.72895 %
+Overall Percentage for Donald Trump: 50.27105 %
+...
+  state_winner     n
+  <chr>        <int>
+1 Harris          19
+2 Trump           23
+
 # Discussion {#sec-mydisc}
+
+TO DO:
+
+FINISH DISCUSSION, EDIT CURRENT BITS
 
@@
 
-## Weaknesses and next steps {#sec-mylimits}
+## Weaknesses and next steps
 
-Weaknesses and next steps should also be included.
+What are some weaknesses of what was done? + implications
+
+Environmental and Societal Factors in Surveys, contributing to the – bias
+
+Recent envoronmental disasters have disruoted life in major swing states
+
+Didnt account for partisian pollsters and the potential bias ,
+
+Weaknesses in  forecasting itself, 
+How trump could lose despite pollster estimates
+
+Lack of data, could introduce bias, not entirely representative, uncertainty due to declaration for Kamala, 
+This can cause
+
+# Limitations {#sec-mylimits}
+
+What is left to learn or how should we proceed in the future?
+
+Future models could benefit from more complex models which account for other potentially influential factors such as poll methodology (i.e.
+online vs in-person surveying) and partisanship or respondents’ political affiliations [@cambridge_polling_factors_2024].
+Models using polls-of-polls approaches which use multi-level regression and post stratification tend to do particularly well in forecasting elections (REF tab 2), and increasing complexity to account for more influential factors like those mentioned above.
+They historically perform /….
+Better estimation .
+and accounts for uncertainty propagation, which is particularly relevant in this quarter’s election.
+(Reference in tab)
+Furthermore, using training data based on historical polls regarding democratic and republican election results and polling data would allow for a more robust model.
+This is especially because of Biden’s late declaration for Harris, thus substituting fot prior state poll behaviour will help account for a lack of polling data.
+Although many forecasting models aim to estimate the national popular vote,
+Our model similarly could benefit from estimations for electoral vote
 

### 2024-11-03 scripts/00-simulate_data.R
Final update of Simulation

@@ -1,6 +1,6 @@
 #### Preamble ####
 # Purpose: Simulates Data
-# Author: Cristina Su Lam, Karen Riani, Mariko Lee
+# Author: Mariko Lee, Karen Riani, Cristina Su Lam
 # Date: 10 October 2024 
 # License: MIT
 # Pre-requisites: None
@@ -13,13 +13,13 @@ set.seed(123)
 #### Simulate data ####
 # Defining the states and candidates
 states <- state.name  # U.S. state names
-candidates <- c("Trump", "Harris")  # Presidential candidates
+candidates <- c("Trump", "Harris") 
 
 # Define a list of pollsters
 pollsters <- c("YouGov", "Siena/NYT", "The Washington Post", "Ipsos", "Quinnipiac")
 
 # Number of simulated polls per state
-num_polls <- 10  # Simulating 10 polls per state
+num_polls <- 10  
 
 # Simulate sample sizes for each poll (between 500 and 2000 voters)
 sample_sizes <- sample(500:2000, num_polls * length(states), replace = TRUE)
@@ -27,14 +27,27 @@ sample_sizes <- sample(500:2000, num_polls * length(states), replace = TRUE)
 # Simulate percentage of support (pct) for each candidate (between 40% and 60%)
 pct_support <- runif(num_polls * length(states), min = 40, max = 60)
 
+# Date
+start_date <- as.Date("2024-07-21")
+end_date <- as.Date("2024-10-31")
+poll_dates <- sample(seq(start_date, end_date, by = "day"), num_polls * length(states), replace = TRUE)
+
+
 # Create the simulated data frame
 simulated_data <- tibble(
-  state = rep(states, each = num_polls),  # Repeat each state for the number of polls
-  pollster = sample(pollsters, num_polls * length(states), replace = TRUE),  # Random pollsters for each poll
-  candidate_preference = sample(candidates, num_polls * length(states), replace = TRUE, prob = c(0.5, 0.5)),  # Equal chance for each candidate
-  pct = pct_support,  # Simulated percentage support
-  sample_size = sample_sizes  # Simulated sample sizes
+  state = rep(states, each = num_polls),
+  pollster = sample(pollsters, num_polls * length(states), replace = TRUE),  
+  candidate_preference = sample(candidates, num_polls * length(states), replace = TRUE, prob = c(0.5, 0.5)), 
+  pct = pct_support, 
+  sample_size = sample_sizes,
+  date = poll_dates
[…]

### 2024-11-03 scripts/00-simulate_data.R
Final Update for Simulation and its Testing

@@ -1,9 +1,9 @@
 #### Preamble ####
-# Purpose: Simulates Data
+# Purpose: Simulates Data about US 2024 Elections.
 # Author: Mariko Lee, Karen Riani, Cristina Su Lam
 # Date: 10 October 2024 
 # License: MIT
-# Pre-requisites: None
+# Pre-requisites: The `tidyverse` package must be installed and loaded
 # Any other information needed? None
 
 #### Workspace setup ####

### 2024-11-03 scripts/01-test_simulated_data.R

@@ -1,6 +1,6 @@
 #### Preamble ####
-# Purpose: Tests the structure and validity of the simulated American elections dataset.
-# Author: Cristina Su Lam, Karen Riani, Mariko Lee
+# Purpose: Tests the structure and validity of the Simulated Data about US 2024 Elections.
+# Author: Mariko Lee, Karen Riani, Cristina Su Lam
 # Date: 10 October 2024 
 # License: MIT
 # Pre-requisites: 
@@ -28,22 +28,59 @@ expected_states <- state.name
 expected_pollsters <- c("YouGov", "Siena/NYT", "The Washington Post", "Ipsos", "Quinnipiac")
 
 # Test 1: Ensure there are no missing values
-stopifnot(!any(is.na(simulate_data)))
+stopifnot(!any(is.na(simulated_data)))
 
 # Test 2: Ensure `pct` values are within a valid range (0% to 100%)
-stopifnot(all(simulate_data$pct >= 0 & simulate_data$pct <= 100))
+stopifnot(all(simulated_data$pct >= 0 & simulated_data$pct <= 100))
 
 # Test 3: Ensure `candidate_preference` contains only "Trump" and "Harris"
-stopifnot(all(simulate_data$candidate_preference %in% expected_candidates))
+stopifnot(all(simulated_data$candidate_preference %in% expected_candidates))
 
 # Test 4: Ensure `state` contains valid U.S. states (this assumes all `state.name` values)
-stopifnot(all(simulate_data$state %in% expected_states))
+stopifnot(all(simulated_data$state %in% expected_states))
 
 # Test 5: Ensure `pollster` contains expected pollster names
-stopifnot(all(simulate_data$pollster %in% expected_pollsters))
+stopifnot(all(simulated_data$pollster %in% expected_pollsters))
 
 # Test 6: Check for duplicates in the simulated data
-stopifnot(nrow(simulate_data) == nrow(distinct(simulate_data)))
+stopifnot(nrow(simulated_data) == nrow(distinct(simulated_data)))
 
 # Test 7: Ensure `sample_size` is within a reasonable range (500 to 2000 as simulated)
-stopifnot(all(simulate_data$sample_size >= 500 & simulate_data$sample_size <= 2000))
+stopifnot(all(simulated_data$sample_size >= 500 & simulated_data$sample_size <= 2000))
+
+# Test 8: Ensure `date` is within the expected range
[…]

### 2024-11-03 scripts/02-download_data.R
Last update of dataset as of Nov 3

@@ -1,9 +1,9 @@
 #### Preamble ####
-# Purpose: Download the presidential poll data 
-# Author: Cristina Su Lam, Karen Riani, Mariko Lee
+# Purpose: Download the presidential poll data from website.
+# Author: Mariko Lee, Karen Riani, Cristina Su Lam
 # Date: 10 October 2024 
 # License: MIT
-# Pre-requisites: None
+# Pre-requisites: The `readr` package must be installed and loaded
 # Any other information needed? None
 
 #### Workspace setup ####

### 2024-11-03 scripts/03-clean_data.R
Final update of the clean dataset.

@@ -1,6 +1,6 @@
 #### Preamble ####
 # Purpose: Cleans the presidential poll data into an analysis data set
-# Author: Cristina Su Lam, Karen Riani, Mariko Lee
+# Author: Mariko Lee, Karen Riani, Cristina Su Lam
 # Date: 10 October 2024 
 # License: MIT
 # Pre-requisites: Run 01-download_data.R

### 2024-11-03 scripts/03-clean_data.R
Delete duplicate obs

@@ -66,5 +66,16 @@ clean_president_polls <- clean_president_polls %>%
     start_date >= ymd("2024-07-21")
   ) 
 
+#### Identify, Count, and Drop Duplicates ####
+# Identify duplicates
+duplicates <- clean_president_polls[duplicated(clean_president_polls), ]
+num_duplicates <- nrow(duplicates)
+message("Number of duplicate observations: ", num_duplicates)
+
+# Drop duplicates and keep only unique rows
+clean_president_polls <- clean_president_polls[!duplicated(clean_president_polls), ]
+message("Number of rows after removing duplicates: ", nrow(clean_president_polls))
+
+
 #### Save data ####
 write_parquet(clean_president_polls, "data/02-analysis_data/clean_president_polls.parquet")

### 2024-11-03 scripts/04-test_analysis_data.R
Add more testing

@@ -24,70 +24,48 @@ stopifnot(!any(is.na(clean_president_polls$pct)))
 stopifnot(!any(is.na(clean_president_polls$sample_size)))
 stopifnot(!any(is.na(clean_president_polls$state)))
 
-#### Test 2: Ensure numeric_grade values are within the expected range
+# Test 2: Ensure numeric_grade values are within the expected range
 stopifnot(all(clean_president_polls$numeric_grade >= 0.5 & clean_president_polls$numeric_grade <= 3.0))
 
-#### Test 3: Ensure state values are either valid U.S. state names
+# Test 3: Ensure state values are valid U.S. state names
 expected_states <- c(state.name)
 stopifnot(all(clean_president_polls$state %in% expected_states))
 
-#### Test 4: Ensure candidate_name contains only "Kamala Harris" and "Donald Trump"
+# Test 4: Ensure candidate_name contains only "Kamala Harris" and "Donald Trump"
 expected_candidates <- c("Kamala Harris", "Donald Trump")
 stopifnot(all(clean_president_polls$candidate_name %in% expected_candidates))
 
-#### Test 5: Ensure pct values are within the expected range (18 to 70)
+# Test 5: Ensure pct values are within the expected range (18 to 70)
 stopifnot(all(clean_president_polls$pct >= 18 & clean_president_polls$pct <= 70))
 
-#### Test 6: Ensure sample_size values are within the expected range (111 to 26230)
+# Test 6: Ensure sample_size values are within the expected range (111 to 26230)
 stopifnot(all(clean_president_polls$sample_size >= 111 & clean_president_polls$sample_size <= 26230))
 
-#### Test 7: Ensure pollster values are within the known list of pollsters
-expected_pollsters <- c(
-  "TIPP", "InsiderAdvantage", "Trafalgar Group", "co/efficient", "Emerson", 
-  "Siena/NYT", "YouGov", "Research Co.", "Ipsos", "Quinnipiac", 
-  "Marist", "The Washington Post", "Redfield & Wilton Strategies", 
-  "Morning Consult", "Pew", "Suffolk", "Big Village", "Noble Predictive Insights", 
-  "Glengariff Group Inc.", "Christopher Newport U.", "Data for Progress", 
-  "Cygnal", "Impact Research", "RMG Research", "St. Anselm", 
[…]

### 2024-11-03 scripts/01-test_simulated_data.R
Last changes

@@ -4,8 +4,8 @@
 # Date: 10 October 2024 
 # License: MIT
 # Pre-requisites: 
-  # - The `tidyverse` package must be installed and loaded
-  # - 00-simulate_data.R must have been run
+#   - 00-simulate_data.R must have been run
+#   - The `tidyverse` package must be installed and loaded
 # Any other information needed? None
 
 

### 2024-11-03 scripts/03-clean_data.R

@@ -3,7 +3,11 @@
 # Author: Mariko Lee, Karen Riani, Cristina Su Lam
 # Date: 10 October 2024 
 # License: MIT
-# Pre-requisites: Run 01-download_data.R
+# Pre-requisites: 
+#   - Run 02-download_data.R
+#   - The `arrow` package must be installed and loaded
+#   - The `janitor` package must be installed and loaded
+#   - The `lubridate` package must be installed and loaded
 # Any other information needed? None
 
 #### Workspace setup ####
@@ -12,7 +16,7 @@ library(tidyverse)
 library(arrow)
 library(janitor)
 library(lubridate)
-library(broom)
+
 
 #### Clean data ####
 clean_president_polls <- read_csv("data/01-raw_data/raw_president_polls.csv", 

### 2024-11-03 scripts/04-test_analysis_data.R

@@ -1,6 +1,6 @@
 #### Preamble ####
 # Purpose: Some tests for Cleaned Dataset
-# Author: Cristina Su Lam, Karen Riani, Mariko Lee
+# Author: Mariko Lee, Karen Riani, Cristina Su Lam
 # Date: 10 October 2024 
 # License: MIT
 # Pre-requisites: Run 03-clean_data.R script
@@ -9,7 +9,6 @@
 
 #### Workspace setup ####
 library(tidyverse)
-library(testthat)
 library(arrow)
 
 clean_president_polls <- read_parquet("data/02-analysis_data/clean_president_polls.parquet")

### 2024-11-03 scripts/06-model_data.R
renamed variables for clarity, added summaries for bayesian

@@ -40,11 +40,11 @@ logistic_model <- glm(
 )
 
 # Predict the probabilities of Harris winning for the entire dataset
-clean_president_polls <- clean_president_polls %>%
+clean_president_polls_log <- clean_president_polls %>%
   mutate(predicted_prob_harris = predict(logistic_model, type = "response"))
 
 # Calculate the weighted average predicted probability for Kamala Harris
-overall_predicted_prob_harris <- mean(clean_president_polls$predicted_prob_harris)
+overall_predicted_prob_harris <- mean(clean_president_polls_log$predicted_prob_harris)
 
 # Convert to percentage
 overall_percentage_harris <- overall_predicted_prob_harris * 100
@@ -56,7 +56,7 @@ cat("Overall Percentage for Donald Trump:", overall_percentage_trump, "%\n")
 
 
 # Calculate the average predicted probability by state
-state_predictions <- clean_president_polls %>%
+state_predictions <- clean_president_polls_log %>%
   group_by(state) %>%
   summarize(avg_predicted_prob_harris = mean(predicted_prob_harris))
 
@@ -107,12 +107,12 @@ test_data <- test_data %>%
 ### Step 5: Model Evaluation ###
 
 # Calculate RMSE between actual outcomes and predicted probabilities
-rmse_value <- sqrt(mean((test_data$is_harris - test_data$predicted_prob_harris)^2))
-rmse_value
+rmse_value1 <- sqrt(mean((test_data$is_harris - test_data$predicted_prob_harris)^2))
+rmse_value1
 
 # Calculate AUC 
-auc_value <- auc(test_data$is_harris, test_data$predicted_prob_harris)
-auc_value
+auc_value1 <- auc(test_data$is_harris, test_data$predicted_prob_harris)
+auc_value1
 
 
 ### Bayesian Model ###
@@ -138,18 +138,16 @@ bayesian_model <- stan_glmer(
   adapt_delta = 0.95
 )
 
-
-
 # Predict probabilities for Harris winning across all posterior draws
 predicted_probs_matrix <- posterior_predict(bayesian_model, type = "response")
 
 # Take the mean across posterior samples for each observation
 # Transpose the matrix to get one mean per observation
-clean_president_polls <- clean_president_polls %>%
+clean_president_polls_bay <- clean_president_polls %>%
   mutate(predicted_prob_harris = rowMeans(t(predicted_probs_matrix)))
 
[…]

### 2024-11-04 scripts/05-exploratory_data_analysis.R
Final changes for the model

@@ -1,16 +1,19 @@
 #### Preamble ####
 # Purpose: Exploratory Data Analysis for Cleaned Dataset
-# Author: Cristina Su Lam, Karen Riani, Mariko Lee
+# Author: Mariko Lee, Karen Riani, Cristina Su Lam
 # Date: 10 October 2024 
 # License: MIT
-# Pre-requisites: Run 03-clean_data.R script
+# Pre-requisites: 
+#   - Run 03-clean_data.R script
+#   - The `rstanarm` package must be installed and loaded
+#   - The `maps` package must be installed and loaded
 # Any other information needed? None
 
 #### Workspace setup ####
 library(tidyverse)
-library(rstanarm)
+library(rstanarm) 
 library(arrow)
-library(maps)
+library(maps) 
 
 #### Read upcoming presidential election forecast data ####
 clean_president_polls <- read_parquet("data/02-analysis_data/clean_president_polls.parquet")

### 2024-11-04 scripts/06-model_data.R

@@ -1,133 +1,44 @@
 #### Preamble ####
-# Purpose: Fitting Model and Making Predictions
-# Author: Cristina Su Lam, Karen Riani, Mariko Lee
+# Purpose: Fitting Model and Making Predictions.
+# Author: Mariko Lee, Karen Riani, Cristina Su Lam
 # Date: 23 October 2024
 # License: MIT
-# Pre-requisites: Run 03-clean_data.R script
+# Pre-requisites: 
+#   - Run 03-clean_data.R script
+#   - The `rsample` package must be installed and loaded
+#   - The `MLmetrics` package must be installed and loaded
+#   - The `modelsummary` package must be installed and loaded
+#   - The `performance` package must be installed and loaded
 # Any other information needed? None
 
 
 #### Workspace setup ####
 library(tidyverse)
-library(arrow)
 library(rstanarm)
-library(pROC)
-
+library(arrow)
+library(rsample)
+library(MLmetrics)
+library(modelsummary)
+library(performance)
 
 #### Read upcoming presidential election forecast data ####
 clean_president_polls <- read_parquet("data/02-analysis_data/clean_president_polls.parquet")
 
-# Convert categorical variables to factors
-clean_president_polls$pollster <- as.factor(clean_president_polls$pollster)
-clean_president_polls$state <- as.factor(clean_president_polls$state)
-
-# Set seed for reproducibility
-set.seed(123)
-
-### Step 1: Prepare Data and Define Initial Model ###
 
 # Add weights to the dataset
 clean_president_polls <- clean_president_polls %>%
-  mutate(weight = (numeric_grade * (sample_size / mean(sample_size))))
+  mutate(weight = (numeric_grade * (sample_size / mean(sample_size))),
+         pollster = as.factor(pollster),
+         state = as.factor(state)
+         )
 
+#### Bayesian Model ####
+# Define priors
+priors <- normal(0.5, 0.1, autoscale = TRUE)
 
-# Fit the initial logistic regression model
-logistic_model <- glm(
-  is_harris ~ pollster + state + sample_size + pct,
-  data = clean_president_polls,
-  family = binomial(link = "logit"),
-  weights = weight
-)
-
-# Predict the probabilities of Harris winning for the entire dataset
-clean_president_polls_log <- clean_president_polls %>%
-  mutate(predicted_prob_harris = predict(logistic_model, type = "response"))
-
[…]

### 2024-11-04 paper/paper.qmd
renaming + abs/intro update

@@
 ---
-title: "2024 US Presidential Election Forecasting"
-subtitle: "Trump forcasted to win popular vote by state over Harris by 0.8%"
+title: "Forecasting the 2024 Election: Tight Margins and Swing State Uncertainty"
+subtitle: "Harris Leads Trump by 0.8% as Voters Remain Divided%"
 author: 
@@
 presidential winner using state-level poll averages for Kamala Harris and Donald Trump, weighted by poll quality and sample size.
-Applying a Bayesian  model, we find a slight 0.5% lead for Trump over Harris, using polls taken since the reelection bid on July 21, 2024.
-Weighting the polls increases the influence of the most reliable data in shaping the forecast—limiting skewed results and showing a clearer voter sentiment amidst a particularly competitive election.
-These results show who may be leading the elections via polls, giving campaigns the insights they need and reinforcing public trust in the election process."
+Applying a logistic regression model, we find a slight 0.8% lead for Harris over Trump, using polls taken since the reelection bid on July 21, 2024.
+​​We weigh the polls with aims to increase the influence of the most reliable data in shaping the forecast—limiting biased results and showing a clearer voter sentiment amidst a particularly competitive election.
+These results show who may be leading the elections via polls, giving campaigns the insights they need and reinforcing public trust in the election process, especially as the outcome likely hinges on key swing states."
 format: pdf
@@
 Amid the chaotic, high-stakes 2024 U.S.
-presidential election, forecasting models offer crucial insights into voter sentiment as Americans face a tight race between top candidates Kamala Harris and Donald Trump.
-With the attempted assassination of Trump, Joe Biden’s endorsement of Harris, and recent crises such as hurricanes disrupting voter access in swing states, public opinion remains volatile [@CBSNews2024_CampaignPromises; @ABCNews2024_HurricaneImpact].
-Key events like Trump’s policy proposals on social security and public safety, Harris’s advocacy for reproductive rights, and other rising debates on gun control have further polarized the electorate, making forecasting models essential in offering clarity amid this competitive landscape [@HRW2022_RoeVWade].
+presidential election, forecasting models offer crucial insights into voter sentiment as Americans face a tight race between top candidates Donald Trump and Kamala Harris.
+With the attempted assassination of Trump, Joe Biden’s endorsement of Harris, and recent crises such as hurricanes disrupting voter access in swing states, public opinion remains volatile [ @ABCNews2024_HurricaneImpact].
+Key events like Trump’s policy proposals on social security and public safety, Harris’s advocacy for reproductive rights, and other rising debates on gun control have further polarized the electorate, making forecasting models essential in offering clarity amid this competitive landscape [@CBSNews2024_CampaignPromises; @HRW2022_RoeVWade].
 
-This study joins the growing body of forecast models—including media outlets’ own polls-of-polls approaches and various Bayesian methods—by refining the aggregation of poll data with Bayesian regression, weighted by sample size and pollster rating [@NYTimes2024_ElectionPolls; @AP2024_ElectionForecasts].
-In an environment where neither candidate shows a decisive lead, this model aims to add precision by emphasizing out higher-quality polls, improving forecast reliability for strategic decision-making and public confidence in election projections [@Bloomberg2024_KeyStates].
+Using presidential polling data from FiveThirtyEight, our Bayesian model’s estimand is the probability of support for Kamala Harris, weighted by pollster rating and sample size to improve the reliability of aggregated polling data.
+By refining the aggregation of polls through logistic regression, weighted by sample size and pollster rating, this study contributes to the growing body of forecast models—including media outlets’ own polls-of-polls and various Bayesian methods.
+Our model estimates only a slim 0.8% lead for Harris over Trump in the overall popular vote—a close margin that highlights the high uncertainty in public sentiment.
+In line with other forecasts that struggle to declare a clear frontrunner, this narrow lead reflects the volatility of the race and the importance of updated, representative polling data [@NYTimes2024_ElectionPolls].
 
-We estimate a slim 0.5% lead for Trump over Harris in the overall popular vote by aggregating state-level polling data, a close margin that highlights the importance of updated, representative data amidst competing forecasts and fluctuating polls.
-By using polls weighted by reliability and focusing on data collected after Harris' campaign declaration, this model maintains transparency, offering a robust perspective on public sentiment and helping to clarify perceptions in a highly competitive election.
+This uncertainty emphasizes the significance of swing states in election forecasts—particularly states with nearly equal support for each party and a history of alternating between Democratic and Republican candidates in recent elections [@Bloomberg2024_KeyStates].
+Existing research identifies seven critical swing states, with Pennsylvania standing out as a tipping point in more than 1 in 5 simulations [@AP2024_ElectionForecasts].
+With 19 electoral votes, Pennsylvania could prove decisive in the event of a close race.
+The competitive nature of these swing states shows the need for continued data updates to maintain forecast accuracy in a closely contested race.
 
-The following structure of this paper is as follows:  @sec-mydatasection presents the data sources and methodology, @sec-mymodel presents the forecasting model, then  @sec-myresults followed by  @sec-mydisc, which discusses the results and their implications.
-@sec-mylimits concludes with a discussion of the limitations of the study and suggestions for future research.
+By prioritizing higher-quality polls and focusing on data collected after Harris’s campaign announcement, the model seeks to mitigate biases and improve forecast reliability by using a polls-of-polls method over individual pollster data.
+This approach not only clarifies voter sentiment amid competing forecasts but also demonstrates the value of weighting variables such as pollster rating and sample size in enhancing prediction accuracy.
+While some states lack sufficient polling data—introducing potential limitations that could be improved with more data and model complexity—the overall findings provide strategic insights for decision-makers and reinforce public confidence in an election where no candidate holds a decisive lead.
+Appendix A provides a detailed assessment on The Washington Post, one of the polling methodologies used by prominent sources, including sampling approaches, weighting, and non-response handling strategies to ensure robust data quality.
+
+The rest of this paper is structured as follows:  @sec-mydatasection presents the data sources and methodology, @sec-mymodel presents the forecasting model, then  @sec-myresults followed by  @sec-mydisc, which discusses the results and their implications.
+@sec-mylimits concludes with a discussion of the limitations of the study and suggestions for future research, followed by @sec-appenx with The Washington Post’s pollster methodology and a budgeted idealized methodology.
 
@@
 
-# Appendix {-}
+# Appendix {#sec-appenx} {-}
 

### 2024-11-04 scripts/06-model_data.R
Added seed, no more changes

@@ -37,6 +37,7 @@ clean_president_polls <- clean_president_polls %>%
 priors <- normal(0.5, 0.1, autoscale = TRUE)
 
 # Fit model with complete dataset
+set.seed(123)
 bayesian_model <- stan_glmer(
   formula = is_harris ~ pct + (1 | pollster) + (1 | state),
   data = clean_president_polls,
@@ -50,14 +51,13 @@ bayesian_model <- stan_glmer(
 )
 
 # Predict probabilities for Harris winning 
-bayesian_predicted <- clean_president_polls |> 
-  select(pollster, state, start_date, end_date, is_harris, pct) |> 
+clean_president_polls <- clean_president_polls |> 
   mutate(
     predicted_prob_harris = posterior_predict(bayesian_model, newdata = clean_president_polls, type = "response") |> colMeans(),
     winner_harris = ifelse(predicted_prob_harris > 0.500, 1, 0)
   )
 # Calculate the unweighted average predicted probability for Kamala Harris
-overall_predicted_prob_harris <- mean(bayesian_predicted$predicted_prob_harris)
+overall_predicted_prob_harris <- mean(clean_president_polls$predicted_prob_harris)
 
 # Convert to percentage
 overall_percentage_harris <- overall_predicted_prob_harris * 100
@@ -68,7 +68,7 @@ cat("Overall Percentage for Kamala Harris:", overall_percentage_harris, "%\n")
 cat("Overall Percentage for Donald Trump:", overall_percentage_trump, "%\n")
 
 # Calculate the average predicted probability by state for Harris
-state_predictions <- bayesian_predicted %>%
+state_predictions <- clean_president_polls %>%
   group_by(state) %>%
   summarize(
     avg_predicted_prob_harris = mean(predicted_prob_harris)

### 2024-11-04 scripts/03-clean_data.R
Final changes

@@ -80,6 +80,5 @@ message("Number of duplicate observations: ", num_duplicates)
 clean_president_polls <- clean_president_polls[!duplicated(clean_president_polls), ]
 message("Number of rows after removing duplicates: ", nrow(clean_president_polls))
 
-
 #### Save data ####
 write_parquet(clean_president_polls, "data/02-analysis_data/clean_president_polls.parquet")

### 2024-11-04 paper/paper.qmd
Appendix A Update

@@
 # Methodology Analysis of The Washington Post Polling
-With an evaluation of sampling methodology, recruitment, handling non-response, and questionnaire design, this appendix offers an analysis of the polling methodology used by The Washington Post in collaboration with ABC News.
-The objective is to analyze these approaches’ strengths and weaknesses and determine how they affect polling accuracy.
+With an evaluation of sampling methodology, recruitment, handling non-response, and questionnaire design, this appendix analyzes the polling methodology used by The Washington Post in collaboration with ABC News.
+The objective is to assess the strengths and weaknesses of these approaches and understand their impact on polling accuracy.
 
 ## Population, Frame, and Sample
-The Washington Post, in partnership with ABC News, employs a combination of text-to-web polls and random digit dialing (RDD) for landlines and mobile phones to reach a large and representative sample of American adults and registered voters [@WashPost2024_ABCMethodology].
+The target population for The Washington Post and ABC News polls includes U.S.
+adults and registered voters.
+The sampling frame, or list from which the sample is drawn, includes landlines, mobile phones, and internet users, covering a broad demographic range.
+Using a combination of text-to-web polls and random digit dialing (RDD) for landlines and mobile phones, they aim to reach a probability sample where each individual in the population has a known chance of selection.
+This minimizes sampling bias and improves the representativeness of the sample [@WashPost2024_ABCMethodology].
 
@@
 These surveys were chosen because they employ suitable stratification and weighting strategies in addition to random sample approaches [@WashPost2024_PollingAverages].
-To represent critical demographics such as age, race, gender, and education, the samples are meticulously weighted [@WashPost2024_ABCMethodology] [@WashPost2024_PollingAverages].
+To represent key demographics such as age, race, gender, and education, the samples are weighted [@WashPost2024_ABCMethodology] [@WashPost2024_PollingAverages].
 
 ## Sample Recruitment
-Live phone interviews and text-to-web surveys collect samples for The Washington Post polls, focusing on ensuring comprehensive demographic coverage.
-In a typical 2024 poll, text-to-web invites reached 21% of respondents, landlines reached 15%, and mobile phones reached 64% of respondents [@WashPost2024_ABCMethodology].
-Younger and minority voters, who might not be well represented in conventional landline-based surveys, can be efficiently reached by pollsters using this technique.
+Data collection includes live phone interviews and text-to-web surveys, ensuring diverse demographic representation.
+In a typical 2024 poll, sample collection consisted of 21% text-to-web invites, 15% landlines, and 64% mobile phones [@WashPost2024_ABCMethodology].
+This technique facilitates access to younger and minority voters, who may not be well represented in conventional landline-based surveys.
 
-By using address-based sampling from the Delivery Sequence File of the US Postal Service, ABC News also leverages probability-based recruiting through the Ipos KnowledgePanel.
+ABC News also uses probability-based recruiting through the Ipos KnowledgePanel by using address-based sampling from the Delivery Sequence File of the US Postal Service.
 Since internet connections and equipment are offered at no cost, this guarantees that even households without internet connections or digital devices are involved [@ABCNews2024_Methodology].
@@
 ## Sampling Approaches and Trade-offs
-Using stratified random sampling, The Washington Post ensures that important demographic groups are represented proportionately to their voter base.
-By using stratified sampling, the polls are more likely to represent the diversity of the voting population accurately.
-To account for over- or under-representation of particular groups, samples are further weighted [@WashPost2024_ABCMethodology] [@WashPost2024_PollingAverages].
+The Washington Post uses stratified random sampling to ensure that important demographics are represented proportionately to their voter base.
+Stratified sampling increases the likelihood of a balanced representation, while weighting addresses over- or under-representation of particular groups [@WashPost2024_ABCMethodology] [@WashPost2024_PollingAverages].
 
-Particularly in situations where state-level polling data is scarce, The Washington Post’s polling averages consider the state’s voting record in the last two presidential elections [@WashPost2024_PollingAverages].
-This adjustment offers a more accurate representation of voters' preferences in states with fewer high-quality polls.
-However, there may be a trade-off since, depending solely on historical data, we may miss recent shifts in voter sentiment [@WashPost2024_ABCMethodology].
+When state-level polling data is scarce, The Washington Post incorporates historical voting records from the last two presidential elections [@WashPost2024_PollingAverages].
+This adjustment helps estimate current preferences but depends on past trends, which may miss recent shifts in voter sentiment [@WashPost2024_ABCMethodology].
 
 ## Non-response Handling
-The Washington Post uses response weighting, which modifies the results according to demographic variables such as age, race, and education, in order to address non-response bias.
-In spite of variations in response rates among demographic groupings, this ensures that the final sample more accurately represents the population [@WashPost2023_Standards].
+Non-response can introduce non-response bias, where the views of non-respondents may differ from respondents, potentially skewing results.
+The Washington Post addresses this issue using response weighting based on age, race, and education to align the final sample with the overall population distribution [@WashPost2023_Standard].
 
 ABC News also addressed non-response bias by applying post-stratification adjustments and sending email reminders to non-respondents.
-In addition, The Washington Post and ABC News both ensure that their samples are weighted to account for any anomalies in non-response [@ABCNews2024_Methodology] [@WashPost2023_Standards].
+In addition, The Washington Post and ABC News ensure that their samples are weighted to account for any anomalies in non-response [@ABCNews2024_Methodology] [@WashPost2023_Standards].
 
@@
 ## Questionnaire Design
-To prevent respondents from being guided toward predetermined responses, The Washington Post creates its surveys with neutrality and clarity in mind.
-The questions are randomized, and respondents are given multiple choices, including “No Opinion,” to avoid pressuring answers [@WashPost2024_ABCMethodology].
-Question order bias can affect how respondents understand and respond to follow-up questions.
-Therefore, rotation helps mitigate this effect [@WashPost2023_Standards].
+The Washington Post creates its surveys with neutrality and clarity to avoid influencing responses.
+Questions are rotated to minimize question order bias, and multiple-choice options, including “No opinion,” are provided [@WashPost2024_ABCMethodology].
+Randomizing questions prevents any unintentional influence on how respondents interpret and answer follow-up questions [@WashPost2023_Standards].
 
-Similar ideas are utilized by ABC News, which offers surveys in both Spanish and English to reach a more representative sample of the general public.
-Leading questions are purposefully omitted from the questionnaires to ensure that the information gathered accurately reflects public opinion [@ABCNews2024_Methodology].
+ABC News follows similar principles, offering surveys in both Spanish and English to ensure inclusivity.
+Leading questions are purposefully omitted from the questionnaires to better capture genuine public opinion across language barriers [@ABCNews2024_Methodology].
 
@@
 **Strength:** \
-**Comprehensive Sampling Method:** The Washington Post can reach a broad demographic, including younger and more difficult-to-reach voters, by combining RDD, text-to-web polls, and live phone interviews [@WashPost2024_ABCMethodology] [@WashPost2023_Standards].\
-**Post-stratification Weighting:** To account for demographic imbalances and increase the accuracy of their polls, The Washington Post and ABC News both use strong post-stratification weighting [@WashPost2024_ABCMethodology] [@ABCNews2024_Methodology].\
-**Transparent Approach:** The Washington Post’s polling data is more credible since they only employ high-quality polls in their averages and is transparent about their methodology [@WashPost2023_Standards] [@WashPost2024_PollingAverages].
+**Sampling Method:** The combination of RDD, text-to-web polls, and live phone interviews enable access to a wide demographic, including younger and harder-to-reach voters [@WashPost2024_ABCMethodology] [@WashPost2023_Standards].
+\
+**Post-stratification Weighting:** Effective weighting adjusts for demographic imbalances, enhancing poll accuracy and representativeness [@WashPost2024_ABCMethodology] [@ABCNews2024_Methodology]. \
+**Transparent Approach:** Using only high-quality polls and transparency in methodology increases the credibility of The Washington Post’s polling data [@WashPost2023_Standards] [@WashPost2024_PollingAverages].
 \
@@
 **Weaknesses:** \
-**Non-response Bias:** Even if both organizations use weighting adjustments, non-response bias still remains a challenge, particularly when it comes to groups that are less likely to respond to surveys [@WashPost2023_Standards] [@WashPost2024_ABCMethodology].
+**Non-response Bias:** Despite response weighting, non-response remains an issue, particularly with groups less inclined to participate, such as younger individuals [@WashPost2023_Standards] [@WashPost2024_ABCMethodology].
 \
-**Dependency on Historical Data:** In states with fewer polls, The Washington Post relies on historical data (the last two presidential elections), which raises the possibility that the polling averages might not accurately reflect current changes in voters preferences [@WashPost2024_PollingAverages].
-\
+**Dependency on Historical Data:** In states with limited polling, depending on past election data may not accurately reflect recent shifts in voter preferences [@WashPost2024_PollingAverages]. \
 
 ## Conclusion
-The polling methodologies used by The Washington Post and ABC News offer a strict framework for gauging popular sentiment in the 2024 US presidential election.
-Their surveys often represent the electorate since they employ various sampling strategies, stratification, and weighting methodologies.
-However, obstacles such as non-response bias and the use of historical data in some states must be addressed appropriately to protect the accuracy and reliability of their polling averages.
+The Washington Post and ABC News polling methodologies offer a framework for measuring public opinion in the 2024 U.S.
+presidential election.
+Employing various sampling strategies, stratification, and weighting improves sample representation.
+However, challenges such as non-response bias and reliance on historical data in under-polled states need continued adjustments to maintain polling reliability.
 

### 2024-11-04 paper/paper.qmd
Appendix B update

@@
 ## Overview
-Using a $100K budget, this appendix outlines a carefully designed survey methodology for predicting the 2024 US Presidential Election.
-The objective is to collect representative, high-quality data using recruiting, poll aggregation, and selective sample methods.
-Through rigorous validation, this approach ensures data accuracy and reduces common survey research errors.
+With a budget of $100K, this appendix outlines the survey methodology designed to predict the 2024 U.S.
+Presidential Election.
+The goal is to collect representative, high-quality data using recruiting, poll aggregation, and selective sample methods.
+This approach aims to minimize errors commonly encountered in survey research and to produce reliable insights.
 
 ## Sampling Approach
-We will implement stratified random sampling to ensure that key demographic and geographic subgroups are fairly represented.
-This approach reduces bias and offers more reliable insights into voter preference.
+We will use stratified random sampling to ensure fair representation across key demographic and geographic subgroups. This approach reduces bias and provides a more accurate view of voter preferences.
 
@@
 - Education Levels  
-- Geographic Representation  
+- Geographic Regions  
 - Political Affiliation
 
-**Sample Size Goal:** 10,000 respondents across states and demographics to achieve **high statistical power** with a margin of error below ±1%
+**Sample Size Goal:**   
+- 10,000 respondents across states and demographics to achieve **high statistical power** with a margin of error below ±1%
 
 **Trade-offs:**  
-- Although stratified sampling increases representativeness, it necessitates accurate demographic information and may raise operating expenses.  
-- **Missing Data:** It’s possible that some demographics (e.g. men) may be less likely to respond. Post-stratification weighting and data imputation will be used to address this issue.
+- While stratified sampling improves representativeness, it requires precise demographic data, which can increase costs.
+
+**Missing Data:**  
+- Specific demographics (e.g., younger males) may have lower response rates. To address this, we will use **post-stratification weighting** and **data imputation**.
 
 ## Recruitment Strategy
-Outline outreach and telephone surveys will be combined in our recruitment strategy to ensure widespread participation from various demographic groups.
+To ensure a broad demographic reach, we will combine **online recruitment** and **random-digit dialing (RDD)** for telephone surveys.
+This approach targets both urban and rural populations, increasing inclusivity.
 
@@
 **Random-Digit Dialing (RDD):**  
-- Phone outreach to reach older, rural voters with poor internet connection. 
+- Phone surveys to engage older and rural voters with limited internet access. 
 - Budget Allocation: $30,000
@@
 **Incentives:**  
-- Participants are offered $5 gift cards to increase response rates. 
+- $5 gift cards offered to participants to increase response rates.
 - Budget Allocation: $20,000
@@
 **Non-Response Handling:**  
-- Increase recruitment incentives for underrepresented groups and use numerous follow-up reminders.
+- Increase incentives for underrepresented groups and send multiple follow-up reminders.
 
 ## Data Validation
-To ensure the accuracy and reliability of responses, we will implement several data validation techniques:
+To ensure data quality, several validation techniques will be implemented:
 
 **Survey Logic Check:**  
-- Recognize and flag responses contradicting one another (e.g., reporting under 18 but registered to vote).  
+- Identifies and flags contradictory responses (e.g., respondents under 18 claiming voter registration).   
 **Attention Check:**  
-- Utilize questions to confirm respondents are actively engaged (e.g., “Select ‘Confirm’ to start questionnaire”)  
+- Includes questions to confirm active engagement (e.g., “Select ‘Confirm’ to proceed”) 
 **Post-Stratification Weighting: **  
-- Adjusting for over- and under-enumeration and weighting the sample to reflect the demographic of the US population.
+- Adjusts the sample for over- and under-enumeration and weighting to reflect the demographic composition of the U.S. population.
 
 **Mode and Measurement Errors:**   
-- We mitigate the impact of using mixed modes (online and telephones) by training enumerators and reducing enumerator bias. Misreporting will be reduced through straightforward questions.
+- Enumerators will be trained to reduce inconsistencies between online and phone responses, and survey questions will be straightforward to minimize misreporting.
 
 ## Poll Aggregation Methodology
-We will employ a poll-of-polls aggregation method to reduce bias and smooth fluctuation across individual polls.
+A **poll-of-polls aggregation** method will be applied to reduce bias and balance fluctuation between individual polls.
 
 **Weighting Criteria:**  
-**Sample Size:** larger samples receive more weight to mirror greater reliability.
+**Sample Size:** larger samples receive higher weight for greater reliability.
 
@@
 ## Survey Implementation 
-Google Forms was used to create and implement the survey, allowing for efficient data collection and safe storage. The main section and sample questions are listed below.
+The survey will be created and administered through Google Forms for efficient data collection and secure storage. The main survey sections and example questions are outlined below.
 
@@
 **Title:** 2024 US Presidential Election Poll \
-**Purpose:** To gather public sentiment and predict election outcomes. \
+**Purpose:** To collect public opinions and forecast election outcomes. \
 **Estimated Time:** Less than 5 minutes \
-**Confidentiality:** All responses are anonymous and used only for research purposes.
+**Confidentiality:** Responses are anonymous and used solely for research
 
-1. What is your age?
+Consent to Participate
+- I confirm that I understand the purpose of the survey, and I agree to participate.
+
+1. What is your age group?
 - Under 18
@@
 - 55-64
-- 65+
+- 65-74
+- 75 and older
 
-2. What is your gender?
+2. What is your gender identity?
 - Male
 - Female
-- Non-binary / Prefer not to say
+- Non-binary
+- Prefer not to disclose
 
@@
 - High school diploma or GED
-- Some college
+- Some college, no degree
+- Associate’s degree
 - Bachelor’s degree
-- Master’s degree or higher
+- Master’s degree
+- Doctorate or professional degree (e.g., MD, JD, PhD)
 
-4. Are you registered to vote?
+4. Are you currently registered to vote in the United States?
 - Yes
-- No
+- No, but I am eligible and plan to register before the election
+- No, and I am not planning to register
+- No, I am not eligible to vote in the United States
 
-5. Who do you intend to vote for in the upcoming presidential election?
+5. What is your current political affiliation?
+- Republican
+- Democrat
+- Independent 
+- Libertarian
+- Green Party
+- Prefer not to disclose
+
+6. How would you describe your political views?
+- Very conservative
+- Somewhat conservative
+- Moderate
+- Somewhat liberal
+- Very liberal
+- Prefer not to disclose
+
+7. Who do you currently support in the 2024 U.S. presidential election?
 - Donald Trump (Republican)
 - Kamala Harris (Democrat)
+- Undecided
 - Other
-- Undecided
 
-6. How likely are you to vote in the upcoming election?
-- Very Likely
-- Somewhat Likely
-- Not Likely
+8. How likely are you to vote in the 2024 presidential election?
+- Extremely likely 
+- Very likely 
+- Moderately likely 
+- Somewhat likely 
+- Not at all likely
 
@@
 ## Conclusion
-This survey methodology uses stratified sampling, multi-channel recruitment, and rigorous data validation procedures to ensure accurate forecasting of the 2024 US Presidential Election.
-We provide a more stable and reliable prediction through poll-polls aggregation, smoothing out fluctuations across polls.
-This design balances accuracy, inclusivity, and efficiency with a carefully considered $100K budget, ensuring the poll gathers meaningful insights into voter sentiment and behavior.
+This survey methodology uses stratified sampling, multi-channel recruitment, and data validation procedures to ensure accurate forecasting of the 2024 US Presidential Election.
+Through poll-of-polls aggregation, it offers a stable and reliable prediction by mitigating fluctuations in individual polls.
+The design aims to balance accuracy, inclusivity, and efficiency within a $100K budget, offering meaningful insights into voter behavior and sentiment.
 

### 2024-11-04 scripts/03-clean_data.R
Did some changes

@@ -16,6 +16,7 @@ library(tidyverse)
 library(arrow)
 library(janitor)
 library(lubridate)
+library(dplyr)
 
 
 #### Clean data ####
@@ -49,7 +50,7 @@ clean_president_polls <- read_csv("data/01-raw_data/raw_president_polls.csv",
                                                    answer = col_skip(), candidate_id = col_skip())) |>
   clean_names()
 
-# Create binary variable for state or national polls
+# Create binary variable, drop missing values
 clean_president_polls <- clean_president_polls %>%
   mutate(
     state = str_replace(state, "Maine CD-[12]", "Maine"),
@@ -58,7 +59,7 @@ clean_president_polls <- clean_president_polls %>%
     state = str_replace(state, "Nebraska CD-3", "Nebraska"),
     end_date = mdy(end_date),
     start_date = mdy(start_date),
-    is_harris = ifelse(candidate_name == "Kamala Harris", 1, 0)) %>%
+    candidate_chosen = ifelse(candidate_name == "Kamala Harris", 1, 0)) %>%
   drop_na(sample_size) %>% 
   drop_na(numeric_grade) %>%
   drop_na(state)

### 2024-11-04 scripts/05-exploratory_data_analysis.R

@@ -18,7 +18,7 @@ library(maps)
 #### Read upcoming presidential election forecast data ####
 clean_president_polls <- read_parquet("data/02-analysis_data/clean_president_polls.parquet")
 
-# Summary statistics for pct and sample_size
+# Summary statistics for pct, sample_size, and numeric_grade
 summary_stats <- clean_president_polls %>%
   summarize(
     mean_pct = mean(pct, na.rm = TRUE),
@@ -28,21 +28,39 @@ summary_stats <- clean_president_polls %>%
     mean_sample_size = mean(sample_size, na.rm = TRUE),
     sd_sample_size = sd(sample_size, na.rm = TRUE),
     min_sample_size = min(sample_size, na.rm = TRUE),
-    max_sample_size = max(sample_size, na.rm = TRUE)
+    max_sample_size = max(sample_size, na.rm = TRUE),
+    mean_numeric_grade = mean(numeric_grade, na.rm = TRUE),
+    sd_numeric_grade = sd(numeric_grade, na.rm = TRUE),
+    min_numeric_grade = min(numeric_grade, na.rm = TRUE),
+    max_numeric_grade = max(numeric_grade, na.rm = TRUE)
   )
 
 # Create a summary table in the desired format
 summary_table <- data.frame(
-  `Statistic` = c("Percentage Support", "Sample Size"),
-  `Mean` = c(format(round(summary_stats$mean_pct, 2), big.mark = ",", scientific = FALSE),
-             format(round(summary_stats$mean_sample_size, 2), big.mark = ",", scientific = FALSE)),
-  `SD` = c(format(round(summary_stats$sd_pct, 2), big.mark = ",", scientific = FALSE),
-           format(round(summary_stats$sd_sample_size, 2), big.mark = ",", scientific = FALSE)),
-  `Min` = c(format(summary_stats$min_pct, big.mark = ",", scientific = FALSE),
-            format(summary_stats$min_sample_size, big.mark = ",", scientific = FALSE)),
-  `Max` = c(format(summary_stats$max_pct, big.mark = ",", scientific = FALSE),
-            format(summary_stats$max_sample_size, big.mark = ",", scientific = FALSE))
+  `Statistic` = c("Percentage Support", "Sample Size", "Numeric Grade"),
+  `Mean` = c(
+    format(round(summary_stats$mean_pct, 2), big.mark = ",", scientific = FALSE),
[…]

### 2024-11-04 scripts/06-model_data.R

@@ -39,7 +39,7 @@ priors <- normal(0.5, 0.1, autoscale = TRUE)
 # Fit model with complete dataset
 set.seed(123)
 bayesian_model <- stan_glmer(
-  formula = is_harris ~ pct + (1 | pollster) + (1 | state),
+  formula = candidate_chosen ~ pct + (1 | pollster) + (1 | state),
   data = clean_president_polls,
   family = binomial(link = "logit"),
   prior = priors,
@@ -54,7 +54,7 @@ bayesian_model <- stan_glmer(
 clean_president_polls <- clean_president_polls |> 
   mutate(
     predicted_prob_harris = posterior_predict(bayesian_model, newdata = clean_president_polls, type = "response") |> colMeans(),
-    winner_harris = ifelse(predicted_prob_harris > 0.500, 1, 0)
+    winner_harris = ifelse(predicted_prob_harris > 0.5 & candidate_chosen == 1, 1, 0)
   )
 # Calculate the unweighted average predicted probability for Kamala Harris
 overall_predicted_prob_harris <- mean(clean_president_polls$predicted_prob_harris)
@@ -105,7 +105,7 @@ testing_data <- testing_data |>
 
 # Fit `Bayesian model` with training dataset
 model_validation_train <- stan_glmer(
-  formula = is_harris ~ pct + (1 | pollster) + (1 | state),
+  formula = candidate_chosen ~ pct + (1 | pollster) + (1 | state),
   data = training_data,
   family = binomial(link = "logit"),
   prior = priors,
@@ -122,22 +122,22 @@ pp_check(model_validation_train)
 # Make predictions on the test set for the Bayesian model
 set.seed(123)
 bayesian_predicted_data <- testing_data |> 
-  select(pollster, state, start_date, end_date, is_harris, pct) |> 
+  select(pollster, state, start_date, end_date, candidate_chosen, pct) |> 
   mutate(
     predicted_prob_harris = posterior_predict(model_validation_train, newdata = testing_data, type = "response") |> colMeans(),
     winner_harris = ifelse(predicted_prob_harris > 0.500, 1, 0)
   )
 
 # Model evaluation metrics for Bayesian model
-f1_score_b <- F1_Score(bayesian_predicted_data$is_harris, bayesian_predicted_data$winner_harris)
-auc_value_b <- AUC(bayesian_predicted_data$is_harris, bayesian_predicted_data$winner_harris)
[…]

### 2024-11-04 paper/paper.qmd
Edited Data Section

--- 2024-11-04 paper/paper.qmd
@@
 library(arrow)
-
-#data <- read_csv(here::here("data/01-raw_data/raw_president_polls.csv"))
-# cleaned_data <- read_parquet("data/02-analysis_data/clean_president_polls.parquet")
-
+library(dplyr)
+library(kableExtra)
 ```
@@
 presidential election, forecasting models offer crucial insights into voter sentiment as Americans face a tight race between top candidates Donald Trump and Kamala Harris.
-With the attempted assassination of Trump, Joe Biden’s endorsement of Harris, and recent crises such as hurricanes disrupting voter access in swing states, public opinion remains volatile [ @ABCNews2024_HurricaneImpact].
-Key events like Trump’s policy proposals on social security and public safety, Harris’s advocacy for reproductive rights, and other rising debates on gun control have further polarized the electorate, making forecasting models essential in offering clarity amid this competitive landscape [@CBSNews2024_CampaignPromises; @HRW2022_RoeVWade].
+With the attempted assassination of Trump, Joe Biden's endorsement of Harris, and recent crises such as hurricanes disrupting voter access in swing states, public opinion remains volatile [@ABCNews2024_HurricaneImpact].
+Key events like Trump's policy proposals on social security and public safety, Harris's advocacy for reproductive rights, and other rising debates on gun control have further polarized the electorate, making forecasting models essential in offering clarity amid this competitive landscape [@CBSNews2024_CampaignPromises; @HRW2022_RoeVWade].
 
-Using presidential polling data from FiveThirtyEight, our Bayesian model’s estimand is the probability of support for Kamala Harris, weighted by pollster rating and sample size to improve the reliability of aggregated polling data.
-By refining the aggregation of polls through logistic regression, weighted by sample size and pollster rating, this study contributes to the growing body of forecast models—including media outlets’ own polls-of-polls and various Bayesian methods.
-Our model estimates only a slim 0.8% lead for Harris over Trump in the overall popular vote—a close margin that highlights the high uncertainty in public sentiment.
+Using presidential polling data from FiveThirtyEight, our Bayesian model's estimand is the probability of support for Kamala Harris, weighted by pollster rating and sample size to improve the reliability of aggregated polling data.
+By refining the aggregation of polls through logistic regression, weighted by sample size and pollster rating, this study contributes to the growing body of forecast models---including media outlets' own polls-of-polls and various Bayesian methods.
+Our model estimates only a slim 0.8% lead for Harris over Trump in the overall popular vote---a close margin that highlights the high uncertainty in public sentiment.
 In line with other forecasts that struggle to declare a clear frontrunner, this narrow lead reflects the volatility of the race and the importance of updated, representative polling data [@NYTimes2024_ElectionPolls].
 
-This uncertainty emphasizes the significance of swing states in election forecasts—particularly states with nearly equal support for each party and a history of alternating between Democratic and Republican candidates in recent elections [@Bloomberg2024_KeyStates].
+This uncertainty emphasizes the significance of swing states in election forecasts---particularly states with nearly equal support for each party and a history of alternating between Democratic and Republican candidates in recent elections [@Bloomberg2024_KeyStates].
 Existing research identifies seven critical swing states, with Pennsylvania standing out as a tipping point in more than 1 in 5 simulations [@AP2024_ElectionForecasts].
@@
 
-By prioritizing higher-quality polls and focusing on data collected after Harris’s campaign announcement, the model seeks to mitigate biases and improve forecast reliability by using a polls-of-polls method over individual pollster data.
+By prioritizing higher-quality polls and focusing on data collected after Harris's campaign announcement, the model seeks to mitigate biases and improve forecast reliability by using a polls-of-polls method over individual pollster data.
 This approach not only clarifies voter sentiment amid competing forecasts but also demonstrates the value of weighting variables such as pollster rating and sample size in enhancing prediction accuracy.
-While some states lack sufficient polling data—introducing potential limitations that could be improved with more data and model complexity—the overall findings provide strategic insights for decision-makers and reinforce public confidence in an election where no candidate holds a decisive lead.
+While some states lack sufficient polling data---introducing potential limitations that could be improved with more data and model complexity---the overall findings provide strategic insights for decision-makers and reinforce public confidence in an election where no candidate holds a decisive lead.
 Appendix A provides a detailed assessment on The Washington Post, one of the polling methodologies used by prominent sources, including sampling approaches, weighting, and non-response handling strategies to ensure robust data quality.
@@
 The rest of this paper is structured as follows: @sec-mydatasection presents the data sources and methodology, @sec-mymodel presents the forecasting model, then @sec-myresults followed by @sec-mydisc, which discusses the results and their implications.
-@sec-mylimits concludes with a discussion of the limitations of the study and suggestions for future research, followed by @sec-appenx with The Washington Post’s pollster methodology and a budgeted idealized methodology.
+@sec-mylimits concludes with a discussion of the limitations of the study and suggestions for future research, followed by @sec-appenx with The Washington Post's pollster methodology and a budgeted idealized methodology.
 
-# Data {#sec-mydatasection}
+# Data Analysis {#sec-mydatasection}
 
-TODO: PROPER FORMATTING FOR DATA SEC, ADD FINAL DETAILS AND FIGURES
+```{r}
+#| include: false
+#| warning: false
+#| message: false
+
+# Read Cleaned Dataset
+clean_president_polls <- read_parquet("/Users/cristinasulam/Forecasting-the-2024-US-Presidential-Elections/data/02-analysis_data/clean_president_polls.parquet")
+```
 
 ## Overview
-Our study uses state-level polling data from Donald Trump and Kamala Harris to predict the 2024 US presidential election.
-The polls used were sourced from FiveThirtyEight’s 2024 Presidential Election Forecast Database (FiveThirtyEight 2024), which compiles polling data from multiple organizations and rates each poll’s quality.
-These ratings assist in identifying and weighing the most reliable polls.
-The dataset used for this analysis spans polls starting from July 21, 2024, when Kamala Harris announced her reelection bid.
 
-Our analysis focuses on the following variables: 
-Polling rating (Pollscore): A numerical score reflecting the historical reliability of polling organizations used to weight polls by accuracy. 
-Sample size: The number of respondents per poll, influencing the poll’s margin of error and its weight in the forecast. 
-Percentage support (pct): The proportion of respondents who express support for each candidate. 
-State: The US state where the poll was conducted OR labeled “National” for nationwide polls. 
-Pollster: The organization that conducted the poll, influencing the credibility and weight of our analysis.
+To forecast the 2024 U.S.
+presidential election, this analysis uses state-level polling data for candidates Kamala Harris and Donald Trump.
+Our primary data source is FiveThirtyEight's 2024 Presidential Election Forecast Database [@FiveThirtyEight], which aggregates polling data from multiple organizations and provides quality ratings for each poll, helping to identify and weigh the most reliable sources.
+The dataset covers polls conducted from July 21, 2024—when Kamala Harris announced her reelection bid—through to November 3, 2024.
 
-Table X - Sample of the cleaned dataset, highlighting key variables crucial in the analysis. 
-[Table]
+Data cleaning was conducted in R [@citeR] using several packages: tidyverse [@tidyverse], janitor [@janitor], and lubridate [@lubridate].
+Duplicate rows were removed, and the dataset was filtered to include only polls for Harris and Trump.
+Observations with missing values in key variables such as 'sample size,' 'numeric grade,' and 'state' were excluded.
+A binary variable, 'Chosen Candidate' (1 for Harris, 0 for Trump), was created to indicate support for each candidate.
+After these adjustments, the dataset was refined from its original 17,765 entries to 2,367 observations.
+The primary variables of interest include pollster, numeric grade, state, sample size, percentage, and chosen candidate.
+A sample of the cleaned dataset is displayed in [@tbl-presidentpolls].
 
-### Measurement and Limitations
-We applied several measurement and limitations considerations to ensure the validity of our dataset:
-Poll Quality: 
-Temporal Dynamics:
-Geographic Coverage:
-Response Bias:
+For visualization and analysis, the knitr (Xie 2014), kableExtra (Zhu 2024), and ggplot2 (Wickham 2016) packages were utilized to generate tables, graphs, and maps, presenting the data clearly and effectively.
+[[[UPDATE THIS LATER]]]
 
-### Outcome Variables
+```{r}
+#| label: tbl-presidentpolls
+#| tbl-cap: Sample of President Polls Dataset
+#| echo: false
+#| warning: false
+#| fig-align: center
+clean_president_polls |>   
+  select(pollster, numeric_grade, state, sample_size, pct, candidate_chosen) |> 
+  head(5) |> 
+  kable(
+    col.names = c("Pollster", "Numeric Grade", "State", "Sample Size", "Percentage", "Chosen Candidate"),
+    booktabs = TRUE
+  )
+```
+
+[@tbl-summary] presents summary statistics of the three numerical variables used.
+The **Percentage Support** variable, representing the proportion of respondents supporting a candidate, has a mean of 47.25% with a standard deviation of 4.41%, ranging from a minimum of 25% to a maximum of 70%.
+This spread reflects varying levels of support for each candidate across different polls.
+
+The **Sample Size** variable, indicating the number of respondents in each poll, has a mean of 930.92 with a substantial standard deviation of 550.61, indicating variability in poll sizes.
+The minimum sample size is 301, while the maximum reaches 6,526, highlighting the range in poll reliability, as larger samples tend to yield more accurate reflections of public sentiment.
+
+Lastly, the **Numeric Grade** variable, which rates pollster quality, shows an average score of 2.28 with a standard deviation of 0.63, spanning from 0.5 to 3.
+This distribution suggests that the dataset includes a mix of polls with varying reliability, with higher-rated polls contributing more heavily to the overall forecast due to weighted adjustments (further discussed in @sec-mymodel).
+```{r}
+#| label: tbl-summary
+#| tbl-cap: Summary Statistics
+#| echo: false
+#| warning: false
+#| fig-align: center
+# Summary statistics for pct, sample_size, and numeric_grade
+summary_stats <- clean_president_polls %>%
+  summarize(
+    mean_pct = mean(pct, na.rm = TRUE),
+    sd_pct = sd(pct, na.rm = TRUE),
+    min_pct = min(pct, na.rm = TRUE),
+    max_pct = max(pct, na.rm = TRUE),
+    mean_sample_size = mean(sample_size, na.rm = TRUE),
+    sd_sample_size = sd(sample_size, na.rm = TRUE),
+    min_sample_size = min(sample_size, na.rm = TRUE),
+    max_sample_size = max(sample_size, na.rm = TRUE),
+    mean_numeric_grade = mean(numeric_grade, na.rm = TRUE),
+    sd_numeric_grade = sd(numeric_grade, na.rm = TRUE),
+    min_numeric_grade = min(numeric_grade, na.rm = TRUE),
+    max_numeric_grade = max(numeric_grade, na.rm = TRUE)
+  )
+
+# Create a summary table in the desired format
+summary_table <- data.frame(
+  `Statistic` = c("Percentage Support", "Sample Size", "Numeric Grade"),
+  `Mean` = c(
+    format(round(summary_stats$mean_pct, 2), big.mark = ",", scientific = FALSE),
+    format(round(summary_stats$mean_sample_size, 2), big.mark = ",", scientific = FALSE),
+    format(round(summary_stats$mean_numeric_grade, 2), big.mark = ",", scientific = FALSE)
+  ),
+  `SD` = c(
+    format(round(summary_stats$sd_pct, 2), big.mark = ",", scientific = FALSE),
+    format(round(summary_stats$sd_sample_size, 2), big.mark = ",", scientific = FALSE),
+    format(round(summary_stats$sd_numeric_grade, 2), big.mark = ",", scientific = FALSE)
+  ),
+  `Min` = c(
+    format(summary_stats$min_pct, big.mark = ",", scientific = FALSE),
+    format(summary_stats$min_sample_size, big.mark = ",", scientific = FALSE),
+    format(summary_stats$min_numeric_grade, big.mark = ",", scientific = FALSE)
+  ),
+  `Max` = c(
+    format(summary_stats$max_pct, big.mark = ",", scientific = FALSE),
+    format(summary_stats$max_sample_size, big.mark = ",", scientific = FALSE),
+    format(summary_stats$max_numeric_grade, big.mark = ",", scientific = FALSE)
+  )
+)
+
+kable(summary_table, format = "pipe")
+```
+
+## Measurement and Limitations
+
+The Numeric Grade variable in the dataset, ranging from 0 ("Poor or not done") to 4 ("Exceptional"), serves as an indicator of poll quality, reflecting the reliability of each poll in capturing true public sentiment on candidate support.
+Based on FiveThirtyEight’s methodology [@ABCNews2024_Methodology], this grading system allows the model to prioritize data from more rigorous and representative polls, enhancing forecast accuracy.
+
+The transformation from real-world public opinion into a poll entry involves several steps.
+Polling agencies survey segments of the population to gauge support for candidates such as Donald Trump and Kamala Harris, with the ultimate goal of turning these sentiments into quantifiable data points that represent broader public opinion.
+Polls employ various sampling techniques—including random digit dialing, online panels, and text-to-web surveys—to capture a representative cross-section of voters.
+However, the process faces inherent challenges, including sampling bias, survey design limitations, and non-response bias, each of which can affect the accuracy of the data collected.
+
+Sampling Bias: Accurate representation requires careful respondent selection.
+If certain demographics (e.g., younger people, rural residents) are underrepresented, the poll may not reflect overall public opinion accurately.
+
+Survey Design: Question framing and order can influence responses. Leading questions or lack of clarity may skew results, affecting the validity of the data collected.
+
+Non-Response Bias: When certain individuals do not participate, it can lead to unrepresentative data.
+Polling agencies attempt to mitigate this through weighting adjustments that account for demographics less likely to respond, such as younger or politically disengaged individuals.
+
+To address these challenges, FiveThirtyEight assigns a Numeric Grade to each poll based on factors such as sampling method, sample size, methodological transparency, and historical pollster accuracy.
+Higher grades (closer to 4) are given to polls with rigorous data collection practices, large representative samples, and transparent methodologies, indicating a high likelihood of reliability.
+Lower grades (closer to 0) suggest methodological concerns that may compromise accuracy, thus marking the poll as less reliable.
+
+In this dataset, each poll’s Numeric Grade directly influences its weight in the forecast model.
+Polls rated higher contribute more significantly to the forecast, while those rated lower have reduced impact.
+By applying weighted adjustments according to poll quality, the model mitigates the influence of less reliable data, ultimately improving the robustness of the prediction.
+This weighting approach helps ensure that the forecast reflects an accurate picture of public sentiment, especially in a high-stakes election where competition is tight and reliable data is crucial.
+
+## Outcome Variable
+### Distribution of Chosen Candidate
 The Percentage of support (pct) for Donald Trump and Kamala Harris is the main outcome variable.
 We forecast which candidate is leading by aggregating these results across states.
-Figure X represents the distribution of support percentages for the two candidates, showing clusters ranging from 40% to 60%, indicating the election’s competitive nature.
+Figure X represents the distribution of support percentages for the two candidates, showing clusters ranging from 40% to 60%, indicating the election's competitive nature.
 
-[Figure X: Distribution of Candidate Support]
+```{r}
+#| label: fig-candidate
+#| tbl-cap: Distribution of Chosen Candidate
+#| echo: false
+#| warning: false
+#| fig-align: center
+ggplot(clean_president_polls, aes(x = factor(candidate_chosen), fill = factor(candidate_chosen))) +
+  geom_bar(alpha = 0.7) +
+  labs(title = "Distribution of Polls by Harris vs Trump", 
+       x = "Chosen Candidate", 
+       y = "Number of Polls",
+       fill = "Chosen Candidate") +  # This sets the legend title
+  scale_x_discrete(labels = c("0" = "Trump", "1" = "Harris")) +
+  scale_fill_manual(values = c("0" = "red", "1" = "blue"), labels = c("Trump", "Harris")) +
+  geom_text(stat = "count", aes(label = scales::percent(..count../sum(..count..), 
+                                                        accuracy = 0.2)), vjust = -0.5) +
+  theme_classic()
+```
 
-### Predictor Variables
-The following predictor variables were used to build the logistic regression model and refine our prediction:
-Pollscore: Historical reliability score for pollster (range: )
-Numeric Grade: (scale: 0-4) 
-Transparency Score: (scale: 0-10) 
-Sample Size: Number of respondents in each poll (typically between 500-3000) 
-Methodology: Survey method (eg. phone interviews, online panels) 
-State: State-level indicators that capture regional variations in support.
+```{r}
+#| label: fig-candbystate
+#| tbl-cap: Distribution of Candidate Preference by State
+#| echo: false
+#| warning: false
+#| fig-align: center
+ggplot(clean_president_polls, aes(x = state, fill = factor(candidate_chosen, labels = c("Trump", "Harris")))) +
+  geom_bar(position = "fill") +
+  labs(title = "Candidate Preference by State", x = "State", y = "Proportion", fill = "Chosen Candidate") +
+  scale_fill_manual(values = c("red", "blue")) +
+  theme_classic() +
+  theme(
+    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, margin = margin(t = 5), size = 8)
+  )
+```
+
+## Predictor Variables
+### Distribution of Pollsters
+```{r}
+#| label: fig-pollsters
+#| tbl-cap: Pollsters by Count and Numeric Grade
+#| echo: false
+#| warning: false
+#| fig-align: center
+# Calculate the number of polls for each pollster with more than 20 polls
+pollsters_over_20 <- clean_president_polls %>%
+  group_by(pollster) %>%
+  filter(n() > 20) %>%
+  summarize(
+    num_polls = n(),
+    avg_numeric_grade = mean(numeric_grade, na.rm = TRUE)
+  )
+
+# Plot 
+ggplot(pollsters_over_20, aes(x = reorder(pollster, -num_polls), y = num_polls)) +
+  geom_bar(stat = "identity", aes(fill = avg_numeric_grade), alpha = 0.7) +
+  scale_fill_gradient(low = "lightblue", high = "darkblue", name = "Numeric Grade") +
+  labs(title = "Distribution of Polls by Pollster", x = "Pollster", y = "Number of Polls") +
+  geom_text(aes(label = round(avg_numeric_grade, 2)), vjust = -0.5, size = 2.5) +
+  theme_classic() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8))
+```
+
+### Distribution of Numeric Grades
+```{r}
+ggplot(clean_president_polls, aes(x = numeric_grade)) +
+  geom_histogram(binwidth = 0.5, fill = "darkblue", alpha = 0.7) +
+  labs(title = "Distribution of Pollster Grades", x = "Numeric Grade", y = "Frequency") +
+  theme_classic()
+```
+
+The following predictor variables were used to build the logistic regression model and refine our prediction: Pollscore: Historical reliability score for pollster (range: ) Numeric Grade: (scale: 0-4) Transparency Score: (scale: 0-10) Sample Size: Number of respondents in each poll (typically between 500-3000) Methodology: Survey method (eg.
+phone interviews, online panels) State: State-level indicators that capture regional variations in support.
 
@@
 
-### Cleaning Process and Analysis
-The data cleaning process involved the following steps:
-Filtering for High-Quality Polls: We included only polls with a numeric grade of X or higher to reduce the impact of unreliable data. 
-Standardizing Dates and Locations: 
-Variable Transformation:
-Converted percentages into absolute numbers using sample size 
-Created binary indicators for candidates (Trump = 0, Harris = 1)
-
-Some of our data is of marriage license (@fig-marriage), from @TorontoOpenData
-
-```{r}
-#| label: fig-marriage
-#| fig-cap: Marriage License
-#| echo: false
-
-```
 # Model {#sec-mymodel}
 
-TODO: FINALIZE MODEL SECTION
-- VALIDATION
-- ALTERNATIVES
-- JUSTIFICATION
-- UPDATE MODEL BELOW
+TODO: FINALIZE MODEL SECTION - VALIDATION - ALTERNATIVES - JUSTIFICATION - UPDATE MODEL BELOW
 
@@
 \end{aligned}
-$$
-- $y_i$ is the dependent variable, representing the count of respondents who support Harris in a given poll, modeled as a binomial outcome.
+$$ - $y_i$ is the dependent variable, representing the count of respondents who support Harris in a given poll, modeled as a binomial outcome.
 - $\beta_0$ is the intercept term, indicating the baseline log-odds of Harris support when all predictors are zero.
-- $\beta_1$, $\beta_2$, $\beta_3$, and $\beta_4$ are the coefficients for the predictor variables **Pollster**, **State**, **Sample Size**, and **Pct** (the percentage of support for Harris in a poll), respectively:
-  - $\beta_1$ represents the adjustment in log-odds based on the specific pollster conducting the survey.
+- $\beta_1$, $\beta_2$, $\beta_3$, and $\beta_4$ are the coefficients for the predictor variables **Pollster**, **State**, **Sample Size**, and **Pct** (the percentage of support for Harris in a poll), respectively: - $\beta_1$ represents the adjustment in log-odds based on the specific pollster conducting the survey.
 - $\beta_2$ accounts for the impact of the state in which the poll is conducted.
@@
 
-Overall Percentage for Kamala Harris: 49.72895 %
-Overall Percentage for Donald Trump: 50.27105 %
-...
-  state_winner     n
-  <chr>        <int>
-1 Harris          19
-2 Trump           23
+Overall Percentage for Kamala Harris: 49.72895 % Overall Percentage for Donald Trump: 50.27105 % ... state_winner n <chr> <int> 1 Harris 19 2 Trump 23
 
@@
 
-Environmental and Societal Factors in Surveys, contributing to the – bias
+Environmental and Societal Factors in Surveys, contributing to the -- bias
 
@@
 
-Weaknesses in  forecasting itself, 
-How trump could lose despite pollster estimates
+Weaknesses in forecasting itself, How trump could lose despite pollster estimates
 
-Lack of data, could introduce bias, not entirely representative, uncertainty due to declaration for Kamala, 
-This can cause
+Lack of data, could introduce bias, not entirely representative, uncertainty due to declaration for Kamala, This can cause
 
@@
 Future models could benefit from more complex models which account for other potentially influential factors such as poll methodology (i.e.
-online vs in-person surveying) and partisanship or respondents’ political affiliations [@cambridge_polling_factors_2024].
+online vs in-person surveying) and partisanship or respondents' political affiliations [@cambridge_polling_factors_2024].
 Models using polls-of-polls approaches which use multi-level regression and post stratification tend to do particularly well in forecasting elections (REF tab 2), and increasing complexity to account for more influential factors like those mentioned above.
-They historically perform /….
+They historically perform /....
 Better estimation .
-and accounts for uncertainty propagation, which is particularly relevant in this quarter’s election.
-(Reference in tab)
-Furthermore, using training data based on historical polls regarding democratic and republican election results and polling data would allow for a more robust model.
-This is especially because of Biden’s late declaration for Harris, thus substituting fot prior state poll behaviour will help account for a lack of polling data.
-Although many forecasting models aim to estimate the national popular vote,
-Our model similarly could benefit from estimations for electoral vote
+and accounts for uncertainty propagation, which is particularly relevant in this quarter's election.
+(Reference in tab) Furthermore, using training data based on historical polls regarding democratic and republican election results and polling data would allow for a more robust model.
+This is especially because of Biden's late declaration for Harris, thus substituting fot prior state poll behaviour will help account for a lack of polling data.
+Although many forecasting models aim to estimate the national popular vote, Our model similarly could benefit from estimations for electoral vote
 
@@
 
-# Appendix {#sec-appenx} {-}
+# Appendix {#sec-appenx} {.unnumbered}
 
 # Methodology Analysis of The Washington Post Polling
+
 With an evaluation of sampling methodology, recruitment, handling non-response, and questionnaire design, this appendix offers an analysis of the polling methodology used by The Washington Post in collaboration with ABC News.
-The objective is to analyze these approaches’ strengths and weaknesses and determine how they affect polling accuracy.
+The objective is to analyze these approaches' strengths and weaknesses and determine how they affect polling accuracy.
 
 ## Population, Frame, and Sample
+
 The Washington Post, in partnership with ABC News, employs a combination of text-to-web polls and random digit dialing (RDD) for landlines and mobile phones to reach a large and representative sample of American adults and registered voters [@WashPost2024_ABCMethodology].
@@
 ## Sample Recruitment
+
 Live phone interviews and text-to-web surveys collect samples for The Washington Post polls, focusing on ensuring comprehensive demographic coverage.
@@
 ## Sampling Approaches and Trade-offs
+
 Using stratified random sampling, The Washington Post ensures that important demographic groups are represented proportionately to their voter base.
@@
 
-Particularly in situations where state-level polling data is scarce, The Washington Post’s polling averages consider the state’s voting record in the last two presidential elections [@WashPost2024_PollingAverages].
+Particularly in situations where state-level polling data is scarce, The Washington Post's polling averages consider the state's voting record in the last two presidential elections [@WashPost2024_PollingAverages].
 This adjustment offers a more accurate representation of voters' preferences in states with fewer high-quality polls.
@@
 ## Non-response Handling
+
 The Washington Post uses response weighting, which modifies the results according to demographic variables such as age, race, and education, in order to address non-response bias.
@@
 ## Questionnaire Design
+
 To prevent respondents from being guided toward predetermined responses, The Washington Post creates its surveys with neutrality and clarity in mind.
-The questions are randomized, and respondents are given multiple choices, including “No Opinion,” to avoid pressuring answers [@WashPost2024_ABCMethodology].
+The questions are randomized, and respondents are given multiple choices, including "No Opinion," to avoid pressuring answers [@WashPost2024_ABCMethodology].
 Question order bias can affect how respondents understand and respond to follow-up questions.
@@
 ## Strengths and Weaknesses of the Methodology
-**Strength:** \
+
+**Strength:**\
 **Comprehensive Sampling Method:** The Washington Post can reach a broad demographic, including younger and more difficult-to-reach voters, by combining RDD, text-to-web polls, and live phone interviews [@WashPost2024_ABCMethodology] [@WashPost2023_Standards].\
 **Post-stratification Weighting:** To account for demographic imbalances and increase the accuracy of their polls, The Washington Post and ABC News both use strong post-stratification weighting [@WashPost2024_ABCMethodology] [@ABCNews2024_Methodology].\
-**Transparent Approach:** The Washington Post’s polling data is more credible since they only employ high-quality polls in their averages and is transparent about their methodology [@WashPost2023_Standards] [@WashPost2024_PollingAverages].
-\
+**Transparent Approach:** The Washington Post's polling data is more credible since they only employ high-quality polls in their averages and is transparent about their methodology [@WashPost2023_Standards] [@WashPost2024_PollingAverages].\
 
-**Weaknesses:** \
-**Non-response Bias:** Even if both organizations use weighting adjustments, non-response bias still remains a challenge, particularly when it comes to groups that are less likely to respond to surveys [@WashPost2023_Standards] [@WashPost2024_ABCMethodology].
-\
-**Dependency on Historical Data:** In states with fewer polls, The Washington Post relies on historical data (the last two presidential elections), which raises the possibility that the polling averages might not accurately reflect current changes in voters preferences [@WashPost2024_PollingAverages].
-\
+**Weaknesses:**\
+**Non-response Bias:** Even if both organizations use weighting adjustments, non-response bias still remains a challenge, particularly when it comes to groups that are less likely to respond to surveys [@WashPost2023_Standards] [@WashPost2024_ABCMethodology].\
+**Dependency on Historical Data:** In states with fewer polls, The Washington Post relies on historical data (the last two presidential elections), which raises the possibility that the polling averages might not accurately reflect current changes in voters preferences [@WashPost2024_PollingAverages].\
 
 ## Conclusion
+
 The polling methodologies used by The Washington Post and ABC News offer a strict framework for gauging popular sentiment in the 2024 US presidential election.
@@
 
-# Idealized Survey & Methodology - $100K Budget
+# Idealized Survey & Methodology - \$100K Budget
 
 ## Overview
-Using a $100K budget, this appendix outlines a carefully designed survey methodology for predicting the 2024 US Presidential Election.
+
+Using a \$100K budget, this appendix outlines a carefully designed survey methodology for predicting the 2024 US Presidential Election.
 The objective is to collect representative, high-quality data using recruiting, poll aggregation, and selective sample methods.
@@
 ## Sampling Approach
+
 We will implement stratified random sampling to ensure that key demographic and geographic subgroups are fairly represented.
@@
 
-**Stratification Criteria:**  
-- Age Groups  
-- Gender  
-- Education Levels  
-- Geographic Representation  
+**Stratification Criteria:**\
+- Age Groups\
+- Gender\
+- Education Levels\
+- Geographic Representation\
 - Political Affiliation
@@
 
-**Trade-offs:**  
-- Although stratified sampling increases representativeness, it necessitates accurate demographic information and may raise operating expenses.  
-- **Missing Data:** It’s possible that some demographics (e.g. men) may be less likely to respond. Post-stratification weighting and data imputation will be used to address this issue.
+**Trade-offs:**\
+- Although stratified sampling increases representativeness, it necessitates accurate demographic information and may raise operating expenses.\
+- **Missing Data:** It's possible that some demographics (e.g. men) may be less likely to respond. Post-stratification weighting and data imputation will be used to address this issue.
 
 ## Recruitment Strategy
+
 Outline outreach and telephone surveys will be combined in our recruitment strategy to ensure widespread participation from various demographic groups.
 
-**Online Recruitment:**  
-- Target ads on Google, Facebook, and Twitter to engage younger voters and urban populations.  
-- Budget Allocation: $25,000
+**Online Recruitment:**\
+- Target ads on Google, Facebook, and Twitter to engage younger voters and urban populations.\
+- Budget Allocation: \$25,000
 
-**Random-Digit Dialing (RDD):**  
-- Phone outreach to reach older, rural voters with poor internet connection. 
-- Budget Allocation: $30,000
+**Random-Digit Dialing (RDD):**\
+- Phone outreach to reach older, rural voters with poor internet connection. - Budget Allocation: \$30,000
 
-**Incentives:**  
-- Participants are offered $5 gift cards to increase response rates. 
-- Budget Allocation: $20,000
+**Incentives:**\
+- Participants are offered \$5 gift cards to increase response rates. - Budget Allocation: \$20,000
 
-**Non-Response Handling:**  
+**Non-Response Handling:**\
 - Increase recruitment incentives for underrepresented groups and use numerous follow-up reminders.
@@
 ## Data Validation
+
 To ensure the accuracy and reliability of responses, we will implement several data validation techniques:
 
-**Survey Logic Check:**  
-- Recognize and flag responses contradicting one another (e.g., reporting under 18 but registered to vote).  
-**Attention Check:**  
-- Utilize questions to confirm respondents are actively engaged (e.g., “Select ‘Confirm’ to start questionnaire”)  
-**Post-Stratification Weighting: **  
+**Survey Logic Check:**\
+- Recognize and flag responses contradicting one another (e.g., reporting under 18 but registered to vote).\
+**Attention Check:**\
+- Utilize questions to confirm respondents are actively engaged (e.g., "Select 'Confirm' to start questionnaire")\
+**Post-Stratification Weighting:**\
 - Adjusting for over- and under-enumeration and weighting the sample to reflect the demographic of the US population.
 
-**Mode and Measurement Errors:**   
+**Mode and Measurement Errors:**\
 - We mitigate the impact of using mixed modes (online and telephones) by training enumerators and reducing enumerator bias. Misreporting will be reduced through straightforward questions.
@@
 ## Poll Aggregation Methodology
+
 We will employ a poll-of-polls aggregation method to reduce bias and smooth fluctuation across individual polls.
 
-**Weighting Criteria:**  
+**Weighting Criteria:**\
 **Sample Size:** larger samples receive more weight to mirror greater reliability.
@@
 ## Survey Implementation
+
 Google Forms was used to create and implement the survey, allowing for efficient data collection and safe storage. The main section and sample questions are listed below.
@@
 
-**Survey Overview** \
-**Title:** 2024 US Presidential Election Poll \
-**Purpose:** To gather public sentiment and predict election outcomes. \
-**Estimated Time:** Less than 5 minutes \
+**Survey Overview**\
+**Title:** 2024 US Presidential Election Poll\
+**Purpose:** To gather public sentiment and predict election outcomes.\
+**Estimated Time:** Less than 5 minutes\
 **Confidentiality:** All responses are anonymous and used only for research purposes.
@@
 1.  What is your age?
+
 -   Under 18
@@
 2.  What is your gender?
+
 -   Male
@@
 3.  What is the highest level of education you have completed?
+
 -   Less than high school
@@
 -   Some college
-- Bachelor’s degree
-- Master’s degree or higher
+-   Bachelor's degree
+-   Master's degree or higher
 
 4.  Are you registered to vote?
+
 -   Yes
@@
 5.  Who do you intend to vote for in the upcoming presidential election?
+
 -   Donald Trump (Republican)
@@
 6.  How likely are you to vote in the upcoming election?
+
 -   Very Likely
@@
 ## Budget Breakdown
-- **Online Recruitment** : $25,000
-- **RDD Recruitment**: $30,000
-- **Incentives for Participants**: $20,000
-- **Data Processing & Validation**: $15,000
-- **Miscellaneous Expenses**: $10,000
 
-**Total**: $100,000
+-   **Online Recruitment** : \$25,000
+-   **RDD Recruitment**: \$30,000
+-   **Incentives for Participants**: \$20,000
+-   **Data Processing & Validation**: \$15,000
+-   **Miscellaneous Expenses**: \$10,000
+
+**Total**: \$100,000
 
 ## Conclusion
+
 This survey methodology uses stratified sampling, multi-channel recruitment, and rigorous data validation procedures to ensure accurate forecasting of the 2024 US Presidential Election.
 We provide a more stable and reliable prediction through poll-polls aggregation, smoothing out fluctuations across polls.
-This design balances accuracy, inclusivity, and efficiency with a carefully considered $100K budget, ensuring the poll gathers meaningful insights into voter sentiment and behavior.
+This design balances accuracy, inclusivity, and efficiency with a carefully considered \$100K budget, ensuring the poll gathers meaningful insights into voter sentiment and behavior.
 

### 2024-11-04 scripts/06-model_data.R
Updates

@@ -119,6 +119,7 @@ model_validation_train <- stan_glmer(
 # Posterior predictive check
 pp_check(model_validation_train)
 
+
 # Make predictions on the test set for the Bayesian model
 set.seed(123)
 bayesian_predicted_data <- testing_data |> 

### 2024-11-04 paper/paper.qmd
Final Update Appendix

--- 2024-11-04 paper/paper.qmd
@@
 title: "Forecasting the 2024 Election: Tight Margins and Swing State Uncertainty"
-subtitle: "Harris Leads Trump by 0.8% as Voters Remain Divided%"
+subtitle: "Harris Leads Trump by 0.8% as Voters Remain Divided"
 author: 
@@
 Applying a logistic regression model, we find a slight 0.8% lead for Harris over Trump, using polls taken since the reelection bid on July 21, 2024.
-​​We weigh the polls with aims to increase the influence of the most reliable data in shaping the forecast—limiting biased results and showing a clearer voter sentiment amidst a particularly competitive election.
+We weigh the polls with aims to increase the influence of the most reliable data in shaping the forecast—limiting biased results and showing a clearer voter sentiment amidst a particularly competitive election.
 These results show who may be leading the elections via polls, giving campaigns the insights they need and reinforcing public trust in the election process, especially as the outcome likely hinges on key swing states."
-format: pdf
+toc: true
+fig_caption: yes
 number-sections: true
 bibliography: references.bib
+format: pdf
 ---
@@
 library(arrow)
-
-#data <- read_csv(here::here("data/01-raw_data/raw_president_polls.csv"))
-# cleaned_data <- read_parquet("data/02-analysis_data/clean_president_polls.parquet")
-
+library(kableExtra)
+library(maps) 
 ```
@@
 The rest of this paper is structured as follows: @sec-mydatasection presents the data sources and methodology, @sec-mymodel presents the forecasting model, then @sec-myresults followed by @sec-mydisc, which discusses the results and their implications.
-@sec-mylimits concludes with a discussion of the limitations of the study and suggestions for future research, followed by @sec-appenx with The Washington Post’s pollster methodology and a budgeted idealized methodology.
+@sec-mylimits concludes with a discussion of the limitations of the study and suggestions for future research, followed by an Appendix in @sec-appenx with The Washington Post's pollster methodology and a budgeted idealized methodology.
 
-# Data {#sec-mydatasection}
+# Data Analysis {#sec-mydatasection}
 
-TODO: PROPER FORMATTING FOR DATA SEC, ADD FINAL DETAILS AND FIGURES
+```{r}
+#| include: false
+#| warning: false
+#| message: false
+
+# Load Parquet file and model
+clean_president_polls <- read_parquet("../data/02-analysis_data/clean_president_polls.parquet")
+bayesian_model <- readRDS("../models/first_model.rds")
+
+```
 
 ## Overview
-Our study uses state-level polling data from Donald Trump and Kamala Harris to predict the 2024 US presidential election.
-The polls used were sourced from FiveThirtyEight’s 2024 Presidential Election Forecast Database (FiveThirtyEight 2024), which compiles polling data from multiple organizations and rates each poll’s quality.
-These ratings assist in identifying and weighing the most reliable polls.
-The dataset used for this analysis spans polls starting from July 21, 2024, when Kamala Harris announced her reelection bid.
+To forecast the 2024 U.S.
+presidential election, this analysis uses state-level polling data for candidates Kamala Harris and Donald Trump.
+Our primary data source is FiveThirtyEight's 2024 Presidential Election Forecast Database [@FiveThirtyEight], which aggregates polling data from multiple organizations and provides quality ratings for each poll, helping to identify and weigh the most reliable sources.
+The dataset covers polls conducted from July 21, 2024—when Kamala Harris announced her reelection bid—through to November 3, 2024.
 
-Our analysis focuses on the following variables: 
-Polling rating (Pollscore): A numerical score reflecting the historical reliability of polling organizations used to weight polls by accuracy. 
-Sample size: The number of respondents per poll, influencing the poll’s margin of error and its weight in the forecast. 
-Percentage support (pct): The proportion of respondents who express support for each candidate. 
-State: The US state where the poll was conducted OR labeled “National” for nationwide polls. 
-Pollster: The organization that conducted the poll, influencing the credibility and weight of our analysis.
+Data cleaning was conducted in R [@citeR] using several packages: tidyverse [@tidyverse], janitor [@janitor], and lubridate [@lubridate].
+Duplicate rows were removed, and the dataset was filtered to include only polls for Harris and Trump.
+Observations with missing values in key variables such as 'sample size,' 'numeric grade,' and 'state' were excluded.
+A binary variable, 'Chosen Candidate' (1 for Harris, 0 for Trump), was created to indicate support for each candidate.
+After these adjustments, the dataset was refined from its original 17,765 entries to 2,367 observations.
+The primary variables of interest include pollster, numeric grade, state, sample size, percentage, and chosen candidate.
+A sample of the cleaned dataset is displayed in [@tbl-presidentpolls].
 
-Table X - Sample of the cleaned dataset, highlighting key variables crucial in the analysis. 
-[Table]
-
-### Measurement and Limitations
-We applied several measurement and limitations considerations to ensure the validity of our dataset:
-Poll Quality: 
-Temporal Dynamics:
-Geographic Coverage:
-Response Bias:
-
-### Outcome Variables
-The Percentage of support (pct) for Donald Trump and Kamala Harris is the main outcome variable.
-We forecast which candidate is leading by aggregating these results across states.
-Figure X represents the distribution of support percentages for the two candidates, showing clusters ranging from 40% to 60%, indicating the election’s competitive nature.
-
-[Figure X: Distribution of Candidate Support]
-
-### Predictor Variables
-The following predictor variables were used to build the logistic regression model and refine our prediction:
-Pollscore: Historical reliability score for pollster (range: )
-Numeric Grade: (scale: 0-4) 
-Transparency Score: (scale: 0-10) 
-Sample Size: Number of respondents in each poll (typically between 500-3000) 
-Methodology: Survey method (eg. phone interviews, online panels) 
-State: State-level indicators that capture regional variations in support.
-
-These selected variables are based on their relevance to polling accuracy and availability across the dataset.
-The analysis regarding their relationships with the outcome variables (percentage support) is further explored.
-
-### Cleaning Process and Analysis
-The data cleaning process involved the following steps:
-Filtering for High-Quality Polls: We included only polls with a numeric grade of X or higher to reduce the impact of unreliable data. 
-Standardizing Dates and Locations: 
-Variable Transformation:
-Converted percentages into absolute numbers using sample size 
-Created binary indicators for candidates (Trump = 0, Harris = 1)
-
-Some of our data is of marriage license (@fig-marriage), from @TorontoOpenData
+For visualization and analysis, the knitr (Xie 2014), kableExtra (Zhu 2024), and ggplot2 (Wickham 2016) packages were utilized to generate tables, graphs, and maps, presenting the data clearly and effectively.
+[[[UPDATE THIS LATER]]]
 
 ```{r}
-#| label: fig-marriage
-#| fig-cap: Marriage License
+#| label: tbl-presidentpolls
+#| tbl-cap: Sample of President Polls Dataset
 #| echo: false
+#| warning: false
+#| fig-align: center
+clean_president_polls |>   
+  select(pollster, numeric_grade, state, sample_size, pct, candidate_chosen) |> 
+  head(5) |> 
+  kable(
+    col.names = c("Pollster", "Numeric Grade", "State", "Sample Size", "Percentage", "Chosen Candidate"),
+    booktabs = TRUE
+  )
+```
 
+[@tbl-summary] presents summary statistics of the three numerical variables used.
+The **Percentage Support** variable, representing the proportion of respondents supporting a candidate, has a mean of 47.25% with a standard deviation of 4.41%, ranging from a minimum of 25% to a maximum of 70%.
+This spread reflects varying levels of support for each candidate across different polls.
+
+The **Sample Size** variable, indicating the number of respondents in each poll, has a mean of 930.92 with a substantial standard deviation of 550.61, indicating variability in poll sizes.
+The minimum sample size is 301, while the maximum reaches 6,526, highlighting the range in poll reliability, as larger samples tend to yield more accurate reflections of public sentiment.
+
+Lastly, the **Numeric Grade** variable, which rates pollster quality, shows an average score of 2.28 with a standard deviation of 0.63, spanning from 0.5 to 3.
+This distribution suggests that the dataset includes a mix of polls with varying reliability, with higher-rated polls contributing more heavily to the overall forecast due to weighted adjustments (further discussed in @sec-mymodel).
+```{r}
+#| label: tbl-summary
+#| tbl-cap: Summary Statistics
+#| echo: false
+#| warning: false
+#| fig-align: center
+# Summary statistics for pct, sample_size, and numeric_grade
+summary_stats <- clean_president_polls %>%
+  summarize(
+    mean_pct = mean(pct, na.rm = TRUE),
+    sd_pct = sd(pct, na.rm = TRUE),
+    min_pct = min(pct, na.rm = TRUE),
+    max_pct = max(pct, na.rm = TRUE),
+    mean_sample_size = mean(sample_size, na.rm = TRUE),
+    sd_sample_size = sd(sample_size, na.rm = TRUE),
+    min_sample_size = min(sample_size, na.rm = TRUE),
+    max_sample_size = max(sample_size, na.rm = TRUE),
+    mean_numeric_grade = mean(numeric_grade, na.rm = TRUE),
+    sd_numeric_grade = sd(numeric_grade, na.rm = TRUE),
+    min_numeric_grade = min(numeric_grade, na.rm = TRUE),
+    max_numeric_grade = max(numeric_grade, na.rm = TRUE)
+  )
+
+# Create a summary table in the desired format
+summary_table <- data.frame(
+  `Statistic` = c("Percentage Support", "Sample Size", "Numeric Grade"),
+  `Mean` = c(
+    format(round(summary_stats$mean_pct, 2), big.mark = ",", scientific = FALSE),
+    format(round(summary_stats$mean_sample_size, 2), big.mark = ",", scientific = FALSE),
+    format(round(summary_stats$mean_numeric_grade, 2), big.mark = ",", scientific = FALSE)
+  ),
+  `SD` = c(
+    format(round(summary_stats$sd_pct, 2), big.mark = ",", scientific = FALSE),
+    format(round(summary_stats$sd_sample_size, 2), big.mark = ",", scientific = FALSE),
+    format(round(summary_stats$sd_numeric_grade, 2), big.mark = ",", scientific = FALSE)
+  ),
+  `Min` = c(
+    format(summary_stats$min_pct, big.mark = ",", scientific = FALSE),
+    format(summary_stats$min_sample_size, big.mark = ",", scientific = FALSE),
+    format(summary_stats$min_numeric_grade, big.mark = ",", scientific = FALSE)
+  ),
+  `Max` = c(
+    format(summary_stats$max_pct, big.mark = ",", scientific = FALSE),
+    format(summary_stats$max_sample_size, big.mark = ",", scientific = FALSE),
+    format(summary_stats$max_numeric_grade, big.mark = ",", scientific = FALSE)
+  )
+)
+
+kable(summary_table, format = "pipe")
 ```
+
+## Measurement and Limitations
+Converting real-world electoral dynamics into data requires careful measurement of public sentiment towards presidential candidates for the 2024 U.S.
+presidential election.
+This dataset is compiled from polling data collected by established pollsters such as Siena/NYT and YouGov, employing diverse methodologies to gauge voter preferences.
+Key variables captured include the percentage of respondents favoring each candidate, sample sizes, and demographic information about the participants.
+
+The measurement process begins with pollsters designing surveys aimed at accurately reflecting voter attitudes.
+To ensure a representative sample, stratified random sampling methods are employed, accounting for the electorate's diverse demographics.
+Poll responses are gathered through various channels, including phone interviews, online surveys, and face-to-face interactions, ultimately quantifying voter support as a percentage for each candidate.
+
+After polling concludes, the dataset undergoes thorough cleaning and processing to rectify missing values and discrepancies.
+Each entry corresponds to a specific poll, providing details such as the date of the poll, the methodology used, and the percentage of support for each candidate.
+A critical component is the Numeric Grade, which ranges from 0.5 ("Poor") to 3.0 ("Exceptional"), indicating poll quality and reliability in capturing true public sentiment.
+Based on FiveThirtyEight’s methodology [@ABCNews2024_Methodology], this grading system prioritizes data from more rigorous and representative polls, thereby enhancing forecast accuracy.
+
+To determine these Numeric Grades, FiveThirtyEight assesses factors like sampling methods, sample sizes, methodological transparency, and historical pollster accuracy.
+Polls with robust data collection practices and larger representative samples receive higher ratings, reflecting a greater likelihood of reliability.
+Conversely, lower ratings signify methodological concerns that could affect accuracy.
+
+In conclusion, this dataset transforms complex voter opinions into structured data, assigning reliability metrics through the Numeric Grade to facilitate systematic analysis of public sentiment trends.
+This methodology aids in making more precise predictions about election outcomes.
+However, challenges such as sampling bias, survey design limitations, and non-response bias must be acknowledged, as they can impact data accuracy and emphasize the importance of careful interpretation of polling results.
+
+## Outcome Variable
+### Distribution of Polls and State-Level Candidate Preference
+The chosen candidate—whether a poll favors Harris or Trump—serves as the outcome variable in this analysis.
+[@fig-candidate] shows the distribution of polls between the two candidates, Harris and Trump, revealing a nearly equal split in support, with 49.8% favoring Trump and 50.2% favoring Harris.
+This close distribution highlights a competitive polling landscape, suggesting that public opinion is almost evenly divided between the candidates.
+
+In [@fig-candbystate], the state-level distribution of support further emphasizes this balance, with only slight variations in candidate preference across states.
+For instance, states like Maryland, Montana, Nevada, Ohio, and Utah show a marginally higher level of support for Harris, representing the Democratic party.
+Conversely, in states such as Florida, Minnesota, and Pennsylvania, there is slightly greater support for Trump, representing the Republican party.
+These minor differences indicate regional trends but reinforce the overall theme of a closely contested race.
+
+```{r}
+#| label: fig-candidate
+#| tbl-cap: Distribution of Chosen Candidate
+#| echo: false
+#| warning: false
+#| fig-align: center
+ggplot(clean_president_polls, aes(x = factor(candidate_chosen), fill = factor(candidate_chosen))) +
+  geom_bar(alpha = 0.7) +
+  labs(title = "Distribution of Polls by Harris vs Trump", 
+       x = "Chosen Candidate", 
+       y = "Number of Polls",
+       fill = "Chosen Candidate") +  # This sets the legend title
+  scale_x_discrete(labels = c("0" = "Trump", "1" = "Harris")) +
+  theme_classic() +
+  scale_fill_manual(values = c("0" = "red", "1" = "blue"), labels = c("Trump", "Harris")) +
+  geom_text(stat = "count", aes(label = scales::percent(..count../sum(..count..), 
+                                                        accuracy = 0.2)), vjust = -0.5) 
+```
+
+```{r}
+#| label: fig-candbystate
+#| tbl-cap: Distribution of Candidate Preference by State
+#| echo: false
+#| warning: false
+#| fig-align: center
+ggplot(clean_president_polls, aes(x = state, fill = factor(candidate_chosen, labels = c("Trump", "Harris")))) +
+  geom_bar(position = "fill") +
+  labs(title = "Candidate Preference by State", x = "State", y = "Proportion", fill = "Chosen Candidate") +
+  scale_fill_manual(values = c("red", "blue")) +
+  theme_classic() +
+  theme(
+    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, margin = margin(t = 5), size = 8)
+  )
+```
+
+## Predictor Variables
+### Distribution of Sample Size
+[@fig-sample] shows the distribution of sample sizes across polls, with a right-skewed pattern indicating that most polls have smaller sample sizes and a few larger outliers.
+The majority of polls range between 300 and 1,500 respondents, peaking around 1,000 respondents, while only a few reach the maximum sample size of 6,526, as noted in [@tbl-summary].
+This highlights that larger sample sizes are relatively uncommon in the dataset.
+
+The prevalence of smaller sample sizes suggests greater variability in the precision of poll estimates, as smaller samples generally introduce more uncertainty.
+In contrast, larger sample sizes provide more stable and reliable reflections of public sentiment.
+This distribution emphasizes the need to account for sample size in the analysis to ensure that more reliable, larger polls are given appropriate weight, thus enhancing the accuracy of public opinion estimates.
+```{r}
+#| label: fig-sample
+#| tbl-cap: Distribution of Sample Size
+#| echo: false
+#| warning: false
+#| fig-align: center
+ggplot(clean_president_polls, aes(x = sample_size)) +
+  geom_histogram(binwidth = 200, fill = "darkred", color = "black", alpha = 0.7) +
+  labs(title = "Distribution of Sample Size", x = "Sample Size", y = "Frequency") +
+  theme_classic()
+```
+
+### Distribution of Percentage of Candidate Support by Chosen Candidate
+[@fig-percentage].
+The density plot shows the distribution of candidate support percentages for Donald Trump and Kamala Harris, with both peaking around the 50% mark, where Harris has a slightly higher concentration near this midpoint.
+The magenta overlap region highlights this balance, showing no significant advantage for either candidate.
+
+A few outliers in both distributions show support above 60% or below 40%, representing occasional strong deviations, though these instances are rare.
+This pattern reflects a balanced public sentiment with minor variation between the candidates and points to a highly competitive race.
+The symmetry and overlap of the distributions reinforce the close competition between Trump and Harris, indicating an almost evenly divided public opinion across the dataset.
+```{r}
+#| label: fig-percentage
+#| tbl-cap: Distribution of Candidate Support by Chosen Candidate
+#| echo: false
+#| warning: false
+#| fig-align: center
+ggplot(clean_president_polls, aes(x = pct, fill = candidate_name)) +
+  geom_density(alpha = 0.5) +
+  labs(title = "Candidate Support Distribution", x = "Percentage Support", y = "Density") +
+  scale_fill_manual(values = c("blue", "red"), name = "Candidate") +
+  theme_classic()
+```
+
+### Distribution of Pollster and Numeric Grades
+[@fig-numeric] displays the distribution of Numeric Grades assigned to pollsters, as per FiveThirtyEight’s grading system, with each grade indicating poll reliability.
+The x-axis shows the Numeric Grades, ranging from 0.5 to 3.0, while the y-axis represents the frequency of each grade.
+
+The chart reveals that most polls come from higher-rated pollsters, with the highest concentrations at grades 2.0 (298 polls) and 3.0 (386 polls).
+This distribution suggests that the dataset primarily comprises data from pollsters with moderate to high reliability, while lower-quality sources, such as those with grades 0.5 and 1.0, are less common.
+
+Complementing this, [@fig-pollsters] shows the distribution of polls by individual pollsters, with bar colors indicating each pollster's Numeric Grade.
+Darker shades of blue represent higher reliability scores.
+Notably, pollsters such as Siena/NYT and Marquette Law School have the highest Numeric Grade of 3.0, marking them as highly reliable sources.
+In contrast, pollsters like Trafalgar Group and Patriot Polling, with scores of 0.7 and 1.1, respectively, are represented as lower-quality sources.
+
+These insights are crucial, as they enable the model to assign weights based on pollster reliability, improving the accuracy of the analysis.
+By prioritizing data from higher-grade pollsters, the model can more effectively capture public sentiment, minimizing the impact of potential biases or inaccuracies from less reliable sources.
+```{r}
+#| label: fig-numeric
+#| tbl-cap: Pollsters by Count and Numeric Grade
+#| echo: false
+#| warning: false
+#| fig-align: center
+ggplot(clean_president_polls, aes(x = numeric_grade)) +
+  geom_bar(fill = "darkred", alpha = 0.7) +
+  geom_text(stat = "count", aes(label = ..count..), vjust = -0.5, size = 3) +
+  labs(title = "Distribution of Numeric Grades", x = "Numeric Grade", y = "Frequency") +
+  theme_classic()
+```
+
+```{r}
+#| label: fig-pollsters
+#| tbl-cap: Pollsters by Count and Numeric Grade
+#| echo: false
+#| warning: false
+#| fig-align: center
+# Calculate the number of polls for each pollster with more than 20 polls
+pollsters_over_20 <- clean_president_polls %>%
+  group_by(pollster) %>%
+  filter(n() > 20) %>%
+  summarize(
+    num_polls = n(),
+    avg_numeric_grade = mean(numeric_grade, na.rm = TRUE)
+  )
+
+# Plot 
+ggplot(pollsters_over_20, aes(x = reorder(pollster, -num_polls), y = num_polls)) +
+  geom_bar(stat = "identity", aes(fill = avg_numeric_grade), alpha = 0.7) +
+  scale_fill_gradient(low = "lightblue", high = "darkblue", name = "Numeric Grade") +
+  labs(title = "Distribution of Polls by Pollster", x = "Pollster", y = "Number of Polls") +
+  geom_text(aes(label = round(avg_numeric_grade, 2)), vjust = -0.5, size = 2.5) +
+  theme_classic() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8))
+```
+
+### Distribution of Polls and Candidate Support by State
+[@fig-numberstate] displays the number of polls conducted in each U.S.
+state.
+Pennsylvania has the highest poll count, followed by other key swing states such as Arizona, Georgia, Michigan, and Wisconsin.
+The concentration of polls in these states likely reflects their strategic importance as battlegrounds where shifts in public opinion can be crucial.
+In contrast, states like Vermont, Wyoming, and West Virginia have far fewer polls, suggesting less emphasis on states with more predictable or less competitive outcomes.
+
+[@fig-support] presents a map indicating the leading candidate by state, with blue for Kamala Harris, red for Donald Trump, and gray for states with insufficient polling data.
+The map reveals a geographic divide, with Trump leading in many central and southern states, while Harris is favored in parts of the Northeast and West.
+This distribution aligns with typical U.S.
+regional political trends, where certain areas have a historical leaning toward one party.
+The map provides a visual summary of these regional preferences, highlighting how candidate support varies across the country.
+```{r}
+#| label: fig-numberstate
+#| tbl-cap: Number of Pollsters by State
+#| echo: false
+#| warning: false
+#| fig-align: center
+ggplot(clean_president_polls, aes(x = reorder(state, -table(state)[state]))) +
+  geom_bar(fill = "darkblue") +
+  labs(title = "Number of Polls by State", x = "State", y = "Count") +
+  theme_classic() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+```
+
+```{r}
+#| label: fig-support
+#| tbl-cap: Candidate with Highest Support Percentage by State
+#| echo: false
+#| warning: false
+#| fig-align: center
+# Summarize data to find the candidate with the highest percentage in each state
+state_winners <- clean_president_polls %>%
+  filter(candidate_name %in% c("Kamala Harris", "Donald Trump")) %>%
+  group_by(state) %>%
+  summarize(winner = candidate_name[which.max(pct)])
+
+# Convert state names to lowercase to match map_data
+state_winners$state <- tolower(state_winners$state)
+
+# Load USA state map data
+usa_states <- map_data("state")
+
+# Merge the winners with the USA map data
+merged_data <- merge(usa_states, state_winners, by.x = "region", by.y = "state", sort = FALSE, all.x = TRUE)
+merged_data <- merged_data[order(merged_data$group, merged_data$order), ]
+
+# Plot the election map
+ggplot(data = merged_data, aes(x = long, y = lat, group = group, fill = winner)) +
+  geom_polygon(color = "black", size = 0.3) + 
+  labs(fill = "Winner") +
+  scale_fill_manual(values = c("Kamala Harris" = "blue", "Donald Trump" = "red")) +
+  coord_fixed(1.3) + 
+  labs(title = "Leading Candidate by State") +
+  theme_void() +
+  theme(
+    legend.position = "right",
+    plot.title = element_text(hjust = 0.5, size = 16, face = "bold")
+  )
+```
+
 # Model {#sec-mymodel}
+## Overview
+This section presents a Bayesian model designed to predict voter support for Donald Trump and Kamala Harris during the 2024 U.S.
+presidential election.
+The model assesses the level of support for each candidate across 43 states, including key swing states.
+It outputs the likelihood that a voter will favor either candidate, utilizing variables such as pollster, state, and percentage of candidate support.
+Additionally, weights were incorporated into the model to account for variations in pollster reliability and sample size.
 
-TODO: FINALIZE MODEL SECTION
-- VALIDATION
-- ALTERNATIVES
-- JUSTIFICATION
-- UPDATE MODEL BELOW
+To conduct this analysis, the R packages rstanarm [@citeArm] and MLmetrics [@MLmetrics] were employed.
+The rstanarm package streamlines the Bayesian modeling process by providing tools for fitting generalized linear models using a Bayesian framework, while MLmetrics supplies various metrics for assessing model performance, including accuracy, F1 score, and area under the curve (AUC).
 
+## Model Structure
+
+* Mathematical Representation
 $$
-\begin{aligned}
-\text{logit}(\mu_i) &= \beta_0 + \beta_1 \times \text{Pollster}_i + \beta_2 \times \text{State}_i + \beta_3 \times \text{Sample Size}_i + \beta_4 \times \text{Pct}_i
-\end{aligned}
+\begin{align*}
+\text{logit}(y_i) &= \beta_0 + \beta_1 \times \text{Pct}_i + u_{\text{pollster}} + u_{\text{state}} \\
+\end{align*}
 $$
-- $y_i$ is the dependent variable, representing the count of respondents who support Harris in a given poll, modeled as a binomial outcome.
-- $\beta_0$ is the intercept term, indicating the baseline log-odds of Harris support when all predictors are zero.
-- $\beta_1$, $\beta_2$, $\beta_3$, and $\beta_4$ are the coefficients for the predictor variables **Pollster**, **State**, **Sample Size**, and **Pct** (the percentage of support for Harris in a poll), respectively:
-  - $\beta_1$ represents the adjustment in log-odds based on the specific pollster conducting the survey.
-  - $\beta_2$ accounts for the impact of the state in which the poll is conducted.
-  - $\beta_3$ adjusts for the influence of the poll's sample size.
-  - $\beta_4$ represents the effect of the poll's percentage support for Harris on the log-odds.
-- The function $\text{logit}(\mu_i)$ transforms the linear combination of predictors into probabilities, providing the estimated probability of support for Harris.
+
+* Component Explanation
+$$
+- $y_i$: Binary outcome variable indicating whether a respondent supports Harris (1) or not (0).
+- $\beta_0$: Intercept term representing the baseline log-odds of support for Harris when the percentage support is zero.
+- $\beta_1$: Coefficient for the percentage support for Harris (Pct), reflecting how changes in support percentage affect the log-odds of selecting Harris.
+- $u_{\text{pollster}}$: Random effect for the specific pollster, accounting for variability in support across different pollsters.
+- $u_{\text{state}}$: Random effect for the state in which the poll is conducted, capturing regional variations in support.
+$$
+
+## Variable Selection 
+In constructing the Bayesian model, careful consideration was given to the selection of predictor variables to ensure that the analysis accurately reflects the complexities of voter support.
+
+* Percentage Support (Pct): This continuous variable is integral to the model as it captures the nuanced impact of varying levels of support for Kamala Harris.
+By utilizing percentage support as a continuous measure, the model allows for a more detailed examination of how incremental changes in support influence the likelihood of selecting Harris over Trump.
+This approach provides a richer understanding of voter behavior compared to categorizing support into discrete groups, which could obscure critical trends and insights.
+
+* Random Effects: The decision to treat pollster and state effects as random variables recognizes the potential for unobserved heterogeneity in the data.
+Different pollsters may employ varied methodologies, which can lead to fluctuations in reported support levels that are not adequately accounted for by fixed effects alone.
+Similarly, geographical differences can impact voter preferences, making it essential to model state-level variability.
+By incorporating random effects for both pollster and state, the model effectively captures this variability, enhancing its robustness and validity.
+This approach acknowledges that the influences on voter support can differ significantly based on the context in which polling is conducted.
+
+## Weights and Priors
+The weights in this model are calculated using the formula:
+$$
+\text{weights} = \text{Numeric Grade} \times \left(\frac{\text{Sample Size}}{\text{Mean(Sample Size)}}\right)
+$$
+This approach ensures that more reliable polls—indicated by a higher Numeric Grade—carry greater influence in the analysis.
+By incorporating the sample size relative to the average sample size, this weighting system acknowledges that larger, more representative samples tend to yield more accurate reflections of public sentiment.
+This method effectively balances the impact of various polls based on their reliability and representation, leading to a more robust estimation of voter support.
+
+Regarding the priors, we define them as follows:
+$$
+\text{priors} \sim \text{Normal}(0.5, 0.1)
+$$
+This prior is centered around 0.5 with a standard deviation of 0.1, reflecting the belief that the probabilities of support for both candidates should be approximately 50% in the absence of data.
+This choice of prior is particularly useful in situations where there is prior knowledge or belief about the expected voter support distribution.
+By using a normal prior, we allow for flexibility in the model, enabling the incorporation of prior information while avoiding overly restrictive assumptions.
+
+## Model Validation and Justification
+### Model Validation Methods
+The model validation process involved out-of-sample testing, where the dataset was split into training and testing sets in a 70-30 ratio.
+Both the Bayesian and logistic models utilized the same features, with the key difference being that the logistic model did not include random effects for pollster and state.
+As seen in [@tbl-comparison].
+The performance of each model was evaluated using three key metrics: F1 Score, Area Under the Curve (AUC), and Root Mean Square Error (RMSE).
+```{r}
+#| label: tbl-comparison
+#| tbl-cap: Comparison of Model Performance Metrics
+#| echo: false
+#| warning: false
+#| fig-align: center
+# Create a data frame with the comparison data
+comparison_data <- data.frame(
+  Model = c("Logistic Model", "Bayesian Model"),
+  `F1 Score` = c(0.4217, 0.5343),
+  AUC = c(0.4261, 0.5301),
+  RMSE = c(0.5551, 0.5096),
+  check.names = FALSE
+)
+
+# Use kable to format the table
+kable(comparison_data)
+```
+
+Although the Bayesian model still shows room for improvement—evidenced by F1 Score and AUC values not close to 1 and a RMSE that indicates some prediction error—it outperforms the logistic model across all metrics.
+This suggests that the Bayesian model, which incorporates random effects to account for variability from different pollsters and states, offers a more nuanced understanding of voter support.
+
+### Diagnostics
+To validate the Bayesian model, several diagnostic checks were conducted to assess its robustness and ensure that underlying assumptions were satisfied.
+Key among these were model convergence assessments and posterior predictive checks.
+The posterior predictive checks provided a visual representation of how well the model's predictions align with the observed data.
+
+The graph reveals a close overlap between the observed outcomes (\(y\)) and the replicated predictions (\(y_{\text{rep}}\)), demonstrating that the model effectively captures the distribution of voter support for Harris.
+This strong alignment suggests that the model not only fits the training data well but also generalizes effectively to new observations.
+The flexibility of the Bayesian framework further allows for the incorporation of prior information and consideration of variability due to pollster and state effects, enhancing the overall robustness of the model (Refer to the Appendix for a visual illustration).
+
+### Model Justification
+The choice of the Bayesian model is justified by its capacity to account for the random effects of pollsters and states, providing a richer understanding of public sentiment.
+This model allows for nuanced insights that a simpler logistic regression might miss, particularly in a context where polling data is subject to variation due to different survey methods and demographic factors.
+Ultimately, while both models exhibit imperfections, the Bayesian model’s superior performance metrics support its selection as the final model for predicting voter support in the 2024 U.S.
+presidential election.
 
@@
 
-TODO: FINISH RESULTS INCLUDING GRAPHS, TABLES, SUMMARY STATS
+```{r}
 
-Overall Percentage for Kamala Harris: 49.72895 %
-Overall Percentage for Donald Trump: 50.27105 %
-...
-  state_winner     n
-  <chr>        <int>
-1 Harris          19
-2 Trump           23
+# Calculate overall percentage for Harris and Trump
+overall_percentages <- data.frame(
+  Candidate = c("Harris", "Trump"),
+  Percentage = c(overall_percentage_harris, overall_percentage_trump)
+)
+
+# Plot national popular vote prediction
+ggplot(overall_percentages, aes(x = Candidate, y = Percentage, fill = Candidate)) +
+  geom_bar(stat = "identity") +
+  geom_text(aes(label = sprintf("%.2f%%", Percentage)), vjust = -0.5) +
+  labs(title = "National Popular Vote Prediction", y = "Percentage", x = "Candidate") +
+  scale_fill_manual(values = c("blue", "red")) +
+  theme_classic()
+```
+
+```{r}
+# Plot state-level predictions for Harris and Trump
+ggplot(state_predictions, aes(x = reorder(state, avg_predicted_prob_harris), y = avg_predicted_prob_harris, fill = state_winner)) +
+  geom_bar(stat = "identity") +
+  labs(title = "State-Level Predicted Support for Harris and Trump", x = "State", y = "Average Predicted Probability for Harris") +
+  scale_fill_manual(values = c("blue", "red")) +
+  theme_classic() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+```
+```{r}
+# Modify your state predictions dataframe to label swing states
+swing_states <- c("Michigan", "Pennsylvania", "Wisconsin", "Arizona", "Georgia", "North Carolina", "Nevada")
+state_predictions <- state_predictions %>%
+  mutate(swing_state = ifelse(state %in% swing_states, "Swing State", "Non-Swing State"))
+
+# Plot with enhancements
+ggplot(state_predictions, aes(x = reorder(state, avg_predicted_prob_harris), y = avg_predicted_prob_harris, fill = swing_state)) +
+  geom_bar(stat = "identity", color = "black", alpha = 0.8) +
+  scale_fill_manual(values = c("Swing State" = "darkblue", "Non-Swing State" = "lightgrey")) +
+  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red", size = 0.8) +
+  labs(
+    title = "Predicted Probability of Harris Winning by State",
+    x = "State",
+    y = "Predicted Probability for Harris",
+    fill = "State Type"
+  ) +
+  annotate("text", x = 1, y = 0.51, label = "0.5 Threshold (Neutral)", color = "red", hjust = 0) +
+  theme_classic() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  coord_flip()
+
+```
+```{r}
+# Calculate the standard error and margin of error for Harris's predicted probability in each state
+state_margins <- clean_president_polls %>%
+  group_by(state) %>%
+  summarize(
+    avg_predicted_prob_harris = mean(predicted_prob_harris),
+    se_predicted_prob_harris = sd(predicted_prob_harris) / sqrt(n()),
+    margin_of_error = 1.96 * se_predicted_prob_harris
+  ) %>%
+  mutate(
+    lower_bound = avg_predicted_prob_harris - margin_of_error,
+    upper_bound = avg_predicted_prob_harris + margin_of_error
+  )
+ggplot(state_margins, aes(x = reorder(state, avg_predicted_prob_harris), y = avg_predicted_prob_harris)) +
+  geom_point(color = "blue") +
+  geom_errorbar(aes(ymin = lower_bound, ymax = upper_bound), width = 0.2, color = "blue") +
+  labs(
+    title = "State-Level Predicted Probability for Harris with Margin of Error",
+    x = "State",
+    y = "Predicted Probability for Harris"
+  ) +
+  coord_flip() +
+  theme_classic() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red") +
+  labs(subtitle = "Dashed line represents a 50% threshold")
+
+```
 
@@
 
-# Appendix {#sec-appenx} {-}
+# Appendix {#sec-appenx}
 
 # Methodology Analysis of The Washington Post Polling  
+With an evaluation of sampling methodology, recruitment, handling non-response, and questionnaire design, this appendix offers an analysis of the polling methodology used by The Washington Post in collaboration with ABC News.
+The objective is to analyze these approaches' strengths and weaknesses and determine how they affect polling accuracy.
+
+## Population, Frame, and Sample  
+The Washington Post, in partnership with ABC News, employs a combination of text-to-web polls and random digit dialing (RDD) for landlines and mobile phones to reach a large and representative sample of American adults and registered voters [@WashPost2024_ABCMethodology].
+
 With an evaluation of sampling methodology, recruitment, handling non-response, and questionnaire design, this appendix analyzes the polling methodology used by The Washington Post in collaboration with ABC News.
@@
 ## Strengths and Weaknesses of the Methodology
+
 **Strength:** \
@@
 
-**Online Recruitment:**  
-- Target ads on Google, Facebook, and Twitter to engage younger voters and urban populations.  
+**Online Recruitment:**\
+- Target ads on Google, Facebook, and Twitter to engage younger voters and urban populations. \
 - Budget Allocation: $25,000
 
-**Random-Digit Dialing (RDD):**  
-- Phone surveys to engage older and rural voters with limited internet access. 
+**Random-Digit Dialing (RDD):**\
+- Phone surveys to engage older and rural voters with limited internet access. \
 - Budget Allocation: $30,000
 
-**Incentives:**  
+**Incentives:**\
 - $5 gift cards offered to participants to increase response rates. 
@@
 
-**Non-Response Handling:**  
+**Non-Response Handling:**\
 - Increase incentives for underrepresented groups and send multiple follow-up reminders.
@@
 
-**Survey Logic Check:**  
-- Identifies and flags contradictory responses (e.g., respondents under 18 claiming voter registration).   
-**Attention Check:**  
-- Includes questions to confirm active engagement (e.g., “Select ‘Confirm’ to proceed”) 
-**Post-Stratification Weighting: **  
+**Survey Logic Check:**\
+- Identifies and flags contradictory responses (e.g., respondents under 18 claiming voter registration). \
+**Attention Check:**\
+- Includes questions to confirm active engagement (e.g., “Select ‘Confirm’ to proceed”) \
+**Post-Stratification Weighting:**\
 - Adjusts the sample for over- and under-enumeration and weighting to reflect the demographic composition of the U.S. population.
 
-**Mode and Measurement Errors:**   
+**Mode and Measurement Errors:**\
 - Enumerators will be trained to reduce inconsistencies between online and phone responses, and survey questions will be straightforward to minimize misreporting.
@@
 A **poll-of-polls aggregation** method will be applied to reduce bias and balance fluctuation between individual polls. 
-
 **Weighting Criteria:**  
-**Sample Size:** larger samples receive higher weight for greater reliability.
-
-**Recency:** More recent polls are given higher weight to capture modern voter sentiment.
-
-**Pollster Rating:** Polls from highly rated pollsters receive higher weights to reduce the impact of bias.
+- **Sample Size:** larger samples receive higher weight for greater reliability.  
+- **Recency:** More recent polls are given higher weight to capture modern voter sentiment.
+- **Pollster Rating:** Polls from highly rated pollsters receive higher weights to reduce the impact of bias.
 
@@
 ## Budget Breakdown
-- **Online Recruitment** : $25,000
-- **RDD Recruitment**: $30,000
-- **Incentives for Participants**: $20,000
-- **Data Processing & Validation**: $15,000
-- **Miscellaneous Expenses**: $10,000
 
-**Total**: $100,000
+-   **Online Recruitment** : \$25,000
+-   **RDD Recruitment**: \$30,000
+-   **Incentives for Participants**: \$20,000
+-   **Data Processing & Validation**: \$15,000
+-   **Miscellaneous Expenses**: \$10,000
+
+**Total**: \$100,000
 
 ## Conclusion
+
 This survey methodology uses stratified sampling, multi-channel recruitment, and data validation procedures to ensure accurate forecasting of the 2024 US Presidential Election.

### 2024-11-04 paper/paper.qmd

@@
 library(maps) 
+library(knitr)
 ```
@@
 
-For visualization and analysis, the knitr (Xie 2014), kableExtra (Zhu 2024), and ggplot2 (Wickham 2016) packages were utilized to generate tables, graphs, and maps, presenting the data clearly and effectively.
-[[[UPDATE THIS LATER]]]
-
+For visualization and analysis, the knitr [@knitr], ggplot2 [@ggplot2_wickham_2016], kableExtra [@kableExtra], map [@maps] packages were utilized to generate tables, graphs, and maps, presenting the data clearly and effectively.
 ```{r}

### 2024-11-04 paper/paper.qmd
added from docs, fixing figures now

@@
 presidential winner using state-level poll averages for Kamala Harris and Donald Trump, weighted by poll quality and sample size.
-Applying a logistic regression model, we find a slight 0.8% lead for Harris over Trump, using polls taken since the reelection bid on July 21, 2024.
+Applying a Bayesian regression model, we find a slight 0.8% overall lead for Harris over Trump and Harris leading in all 7 swing states.
 We weigh the polls with aims to increase the influence of the most reliable data in shaping the forecast—limiting biased results and showing a clearer voter sentiment amidst a particularly competitive election.
-These results show who may be leading the elections via polls, giving campaigns the insights they need and reinforcing public trust in the election process, especially as the outcome likely hinges on key swing states."
+These results show who may be leading the elections via polls, giving campaigns the estimations they need and reinforcing public trust in the election process, especially as the outcome likely hinges on key swing states."
 toc: true
@@
 library(knitr)
+library(ggplot2)
+library(modelsummary)
 ```
@@
 Amid the chaotic, high-stakes 2024 U.S.
-presidential election, forecasting models offer crucial insights into voter sentiment as Americans face a tight race between top candidates Donald Trump and Kamala Harris.
-With the attempted assassination of Trump, Joe Biden's endorsement of Harris, and recent crises such as hurricanes disrupting voter access in swing states, public opinion remains volatile [@ABCNews2024_HurricaneImpact].
-Key events like Trump's policy proposals on social security and public safety, Harris's advocacy for reproductive rights, and other rising debates on gun control have further polarized the electorate, making forecasting models essential in offering clarity amid this competitive landscape [@CBSNews2024_CampaignPromises; @HRW2022_RoeVWade].
+presidential election, forecasting models offer useful evaluations into voter sentiment as Americans face a tight race between top candidates Donald Trump and Kamala Harris.
+With the attempted assassination of Trump, Joe Biden’s endorsement of Harris, and recent crises such as hurricanes disrupting voter access in swing states, public opinion remains volatile [ @ABCNews2024_HurricaneImpact].
+Key events like Trump’s policy proposals on social security and public safety, Harris’s advocacy for reproductive rights, and other rising debates on gun control have further polarized the electorate, making forecasting models essential in offering clarity amid this competitive landscape [@CBSNews2024_CampaignPromises; @HRW2022_RoeVWade].
 
-Using presidential polling data from FiveThirtyEight, our Bayesian model's estimand is the probability of support for Kamala Harris, weighted by pollster rating and sample size to improve the reliability of aggregated polling data.
-By refining the aggregation of polls through logistic regression, weighted by sample size and pollster rating, this study contributes to the growing body of forecast models---including media outlets' own polls-of-polls and various Bayesian methods.
-Our model estimates only a slim 0.8% lead for Harris over Trump in the overall popular vote---a close margin that highlights the high uncertainty in public sentiment.
+Using presidential polling data from FiveThirtyEight, our Bayesian model’s estimand is the probability of support for Kamala Harris, weighted by pollster rating and sample size to improve the reliability of aggregated polling data.
+By refining the aggregation of polls through Bayesian regression, filtered from the reelection day past July 21st, 2024 and weighted by sample size and pollster rating, this study contributes to the growing body of forecast models—including media outlets’ own polls-of-polls and various Bayesian methods.
+Our model estimates only a slim 0.8% lead for Harris over Trump in the overall popular vote—a close margin that highlights the high uncertainty in public sentiment.
 In line with other forecasts that struggle to declare a clear frontrunner, this narrow lead reflects the volatility of the race and the importance of updated, representative polling data [@NYTimes2024_ElectionPolls].
 
-This uncertainty emphasizes the significance of swing states in election forecasts---particularly states with nearly equal support for each party and a history of alternating between Democratic and Republican candidates in recent elections [@Bloomberg2024_KeyStates].
+This uncertainty emphasizes the significance of swing states in election forecasts—particularly states with nearly equal support for each party and a history of alternating between Democratic and Republican candidates in recent elections [@Bloomberg2024_KeyStates].
 Existing research identifies seven critical swing states, with Pennsylvania standing out as a tipping point in more than 1 in 5 simulations [@AP2024_ElectionForecasts].
 With 19 electoral votes, Pennsylvania could prove decisive in the event of a close race.
+Our estimates reveal Harris’ lead in each swing state, however they are slim leads.
 The competitive nature of these swing states shows the need for continued data updates to maintain forecast accuracy in a closely contested race.
 
-By prioritizing higher-quality polls and focusing on data collected after Harris's campaign announcement, the model seeks to mitigate biases and improve forecast reliability by using a polls-of-polls method over individual pollster data.
+By prioritizing higher-quality polls and focusing on data collected after Harris’s campaign announcement, the model seeks to mitigate biases and improve forecast reliability by using a polls-of-polls method over individual pollster data.
 This approach not only clarifies voter sentiment amid competing forecasts but also demonstrates the value of weighting variables such as pollster rating and sample size in enhancing prediction accuracy.
-While some states lack sufficient polling data---introducing potential limitations that could be improved with more data and model complexity---the overall findings provide strategic insights for decision-makers and reinforce public confidence in an election where no candidate holds a decisive lead.
+While some states lack sufficient polling data—introducing potential limitations that could be improved with more data and model complexity—the overall findings provide strategic forecasts for decision-makers and reinforce public confidence in an election where no candidate holds a decisive lead.
 Appendix A provides a detailed assessment on The Washington Post, one of the polling methodologies used by prominent sources, including sampling approaches, weighting, and non-response handling strategies to ensure robust data quality.
@@
 The rest of this paper is structured as follows:  @sec-mydatasection presents the data sources and methodology, @sec-mymodel presents the forecasting model, then  @sec-myresults followed by  @sec-mydisc, which discusses the results and their implications.
-@sec-mylimits concludes with a discussion of the limitations of the study and suggestions for future research, followed by an Appendix in @sec-appenx with The Washington Post's pollster methodology and a budgeted idealized methodology.
+@sec-mylimits concludes with a discussion of the limitations of the study and suggestions for future research, followed by @sec-appenx with The Washington Post’s pollster methodology and a budgeted idealized methodology.
 
@@
 
+Figure X shows the national predicted popular vote percentages for Harris (50.42%) and Trump (49.58%), calculated by averaging the mean predictions across all states.
+This aggregation reveals a very narrow 0.84% lead for Harris, highlighting a highly competitive race.
+This reveals a lack of dominance in candidacy and general uncertainty of pollster outcomes, suggesting that even this slight lead may not be substantial.
+
+Additionally, Figure Y presents a heat map of Harris's predicted probability by state, illustrating her lead over the majority of states but revealing that this advantage is generally weak, with most not peaking over a 1% advantage.
+Most states show only a slight preference for Harris over Trump, with the margin of error spanning both candidates in many cases.
+Specifically, out of the 43 states with available data, 25 states lean toward Harris, yet none of these states show lead of greater than XX over Trump.
+This pattern additionally shows the substantial closeness of the race, as evidenced by the margin of error, where overlapping intervals highlight the slim advantage and limited certainty in Harris's predicted support.
+
+Figure Z zooms in on the seven swing states, each of which shows a slight lean toward Harris.
+The graph includes a red dashed line at the 50% threshold, and although Harris is estimated to lead in these swing states, the margin is extremely tight, with most probabilities hovering just above 50%.
+Only Michigan and North Carolina show a slight lead of just over 1%, reinforcing the sensitivity of these predictions.
+This pattern suggests that even small shifts in voter sentiment could tip these critical states in either direction.
+
+Looking closer at the margin of error, in figure X we can see Harris’s predicted probabilities by state.
+States to the right of the dashed line (above 0.5) lean towards Harris, while those to the left lean towards Trump.
+In swing states like North Carolina, Michigan, and Pennsylvania, Harris’s minimal lead is coupled with error bars that overlap this threshold.
+This suggests a high level of uncertainty—small shifts could easily change the lead in these states.
+Harris has a slight edge, but in many states, this lead is within the margin of error, thus showing that the election remains too close to call.
+
+Figure W provides a summary of the model’s coefficients and fit statistics, capturing key model parameters like the intercepts, predictor estimates, and evaluation metrics.
+The negative intercept of -2.07 represents the baseline log odds for Harris, establishing the prediction's starting point in the absence of other predictors.
+With all intercepts being equal, there is initially a lower probability for Harris’s support without accounting for additional factors of percentage (pct) of support or pollster influence.
+The non-significant coefficient for sample size indicates that it does not directly affect prediction but serves as part of the poll weighting.
+In contrast, the pct coefficient of 0.04 suggests a slight positive effect on the odds for Harris, meaning an increase in her polling percentage slightly raises her predicted support odds.
+Low variance estimates for pollster and state effects imply limited inter-group variability, indicating that neither adds substantial random variation to the results.
+
+Regarding model evaluation, the WAIC of 7072.4 and log likelihood of -3518.4 do not offer much information alone however a higher likelihood and a lower WAIC is preferred.
+The RMSE of 0.5 points to moderate accuracy, which aligns with the close and uncertain nature of the election.
+Additionally, a low ICC value suggests limited random variance contributed by pollster and state, furthermore showing the model’s constrained precision in forecasting a definitive winner.
+
 ```{r}
+#| label: fig-summary-mod
+#| tbl-cap: "Distribution of Sample Size"
+#| echo: false
+#| warning: false
+#| fig-align: center
+
+# Display model summary
+modelsummary(bayesian_model, output = "kableExtra")
+```
+
+```{r}
+```
+
+```{r}
+#| label: fig-national-vote
+#| fig-cap: "National Popular Vote Prediction"
 
@@
   theme_classic()
+
 ```
@@
 
-TO DO:
+In a highly competitive election, accurately forecasting outcomes requires more than just raw polling data—it demands careful aggregation and weighting to capture true voter sentiment. 
+With this in mind, our model estimates a narrow 0.8% lead for Kamala Harris in the national popular vote, resulting in an aggregated support of approximately 50.4% across states.
 
-FINISH DISCUSSION, EDIT CURRENT BITS
+While this suggests a slight edge for Harris, the lead easily falls within a \pm 3\% margin that reflects substantial uncertainty [@PewResearch_MarginOfError].
+When a candidate’s lead is within this margin, the difference may not be statistically significant, meaning the true preference could feasibly favor either candidate.
+In this context, a 0.8% lead falls well within this range and could be due to sampling noise rather than a clear voter preference.
+In close races, even minor shifts in voter sentiment or turnout in key states could alter the outcome, showing the need for monitoring closely divided swing states and precise state-level analysis.
 
-## First discussion point {#sec-first-point}
+Our model, through poll aggregation, weighted by pollster reliability and sample size, refines polling data into a more realistic prediction, helping offset biases in individual polls.
+This is because the weighting higher-quality polls reduces the influence of less reliable data seen in unweighted or individual poll estimates, providing a clearer view of voter sentiment.
+This approach, especially in closely contested swing states, increased the model’s sensitivity to shifts in these pivotal regions, reflecting on their power to overturn the elections while having a historically uncertain favourable candidate.
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+In the 7 swing states– Arizona, Wisconsin, North Carolina, Pennsylvania, Georgia, Michigan and Nevada, our model shows Harris leading in each, with only 2 swing states showing a lead greater than 1%.
+Furthermore, the small margins seen in FUGURENARGIN show swing state leads ithin their respective margin of errors, pushing the uncertainty of the estimated lead.
+Other forecasts like FiveThirtyEight’s and The NYTimes’ offer similar leads for Harris [@FiveThirtyEight; nytimes2024_estimate].
+When independent models using varied methodologies predict a similar outcome, it can suggest a trend that might hold under different sampling assumptions or adjustments.
+This convergence across forecasts further emphasizes the tightness of the race however, they can enhance confidence in a potential Harris lead, although it remains small.
 
-## Second discussion point
+The results are can additionally be hard to anticipate due to various unaccounted biases and inaccurate model adjustments.
+Trump may simply overturn his polls despite Harris’ slight forecasted leads due to shy voters, where people don’t publicly identify as conservative due to the negative social stigmas around it.
+What is more likely is an overturn due to the non-response bias.
+Republicans tend to be less responsive when it comes to polls which can lead to an unrepresentative dataset despite reweighed models altered to account for poor levels of distributed surveys.
+As there is no guarantee that the altered models work, it is not a confident show of prediction.
+Harris may also experience a large overturn due to similar inaccurately adjusted estimates in attempts to mitigate seemingly nonrepresentative data to an accurate prediction.
+However there the election may face incorrect election results due to small surveying or sampling errors, which can be devastating due to the tightness of the race [@NYT_ElectionPollsOpinion].
 
-## Third discussion point
-
-## Weaknesses and next steps
-
-What are some weaknesses of what was done? + implications
-
-Environmental and Societal Factors in Surveys, contributing to the -- bias
-
-Recent envoronmental disasters have disruoted life in major swing states
-
-Didnt account for partisian pollsters and the potential bias ,
-
-Weaknesses in forecasting itself, How trump could lose despite pollster estimates
-
-Lack of data, could introduce bias, not entirely representative, uncertainty due to declaration for Kamala, This can cause
-
-# Limitations {#sec-mylimits}
-
-What is left to learn or how should we proceed in the future?
-
-Future models could benefit from more complex models which account for other potentially influential factors such as poll methodology (i.e.
-online vs in-person surveying) and partisanship or respondents' political affiliations [@cambridge_polling_factors_2024].
-Models using polls-of-polls approaches which use multi-level regression and post stratification tend to do particularly well in forecasting elections (REF tab 2), and increasing complexity to account for more influential factors like those mentioned above.
-They historically perform /....
-Better estimation .
-and accounts for uncertainty propagation, which is particularly relevant in this quarter's election.
-(Reference in tab) Furthermore, using training data based on historical polls regarding democratic and republican election results and polling data would allow for a more robust model.
-This is especially because of Biden's late declaration for Harris, thus substituting fot prior state poll behaviour will help account for a lack of polling data.
-Although many forecasting models aim to estimate the national popular vote, Our model similarly could benefit from estimations for electoral vote
+# Limitations and Next Steps {#sec-mylimits}
+This analysis identifies several limitations that may impact the reliability of the predictions.
+First, the logistic model shows high levels of multicollinearity between state and pollster variables.
+This multicollinearity may inflate standard errors, complicating the interpretation of coefficients and impacting the model's accuracy.
+Addressing this issue in future analyses could enhance interpretability and offer a clearer comparison to the Bayesian and other models.
+Additionally, environmental disruptions, such as recent hurricanes in key swing states, may introduce biases.
+These events affect both public sentiment and access to polling, potentially skewing results.
+While the state-level margin of error partially reflects this uncertainty, it also shows the model’s sensitivity to external societal influences that are challenging to control.
+Maine and Nebraska, which allocate some electoral votes by congressional district, are simplified in this analysis by grouping each state into a single unit.
+While this approach captures broader state trends, it could underrepresent district-specific voting patterns, and future studies could address district-level data.
+The model is fairly simple in comparison to other forecasts and does not currently adjust for other traditional variables or bias (i.e.
+partisianship) as polling sources, which could lead to an unbalanced or less accurate view of candidate support [@cambridge_polling_factors_2024].
+Incorporating adjustments for pollster partisanship would improve accuracy by reflecting a more comprehensive cross-section of public opinion.
+Another simplification in this analysis is the treatment of Maine and Nebraska, where data were aggregated at the state level instead of by congressional district.
+While this approach provides an overall view of candidate support in these states, it overlooks district-specific dynamics that could impact electoral vote allocation.
+This simplification streamlines the model but may underrepresent nuanced trends within these states.
+Our method of weighting the polls does not have any constraints and could lead to the overweighting of certain data groups.
+If one sample size was much bigger than the majority, its weight would be extreme as well, potentially overcompensating for its affects.
+Finally, data gaps remain a significant factor, as polling data for several states were unavailable.
+Although these states are not swing states, the lack of data introduces uncertainty and potentially limits the model’s representativeness.
+This sparsity in effect with Harris’ recent campaign declaration, contributes to an overall margin of error.
+Future Directions
+Future models could benefit from greater complexity to address these limitations.
+Incorporating additional influential factors, such as poll methodology (e.g., online versus in-person surveys), and pollster partisanship, would add depth to the analysis.
+Multi-level regression and post-stratification approaches, commonly used in polls-of-polls models, have shown to be effective in forecasting elections by capturing uncertainty more comprehensively [@citeTS].
+Increasing model complexity with these factors could improve forecast reliability, particularly in an election as competitive as this one.
+Integrating historical state-level polling data from previous elections could help fill data gaps and provide stronger context for recent trends.
+This is especially relevant given Biden’s late endorsement of Harris, where past polling data could help address the deficit of recent, state-level Democratic polling information.
+While this model focuses on the national popular vote, future work could benefit from projections of electoral votes, as these hold greater relevance in determining the election outcome.
+In practice, reducing the influence of large polls is essential to avoid disproportionately weighting them.
+Implementing a cap on sample sizes, along with winsorizing extreme values, would prevent larger surveys from skewing results.
+Similarly, adjusting the weights of pollsters that release multiple polls in a short time frame would balance their impact, ensuring no single pollster overwhelms the dataset’s representation [@FiveThirtyEight_Poll_Avg ].
 

### 2024-11-04 paper/paper.qmd
did some changes

--- 2024-11-04 paper/paper.qmd
@@
 
-In conclusion, this dataset transforms complex voter opinions into structured data, assigning reliability metrics through the Numeric Grade to facilitate systematic analysis of public sentiment trends.
+The dataset transforms complex voter opinions into structured data, assigning reliability metrics through the Numeric Grade to facilitate systematic analysis of public sentiment trends.
 This methodology aids in making more precise predictions about election outcomes.
@@
 These minor differences indicate regional trends but reinforce the overall theme of a closely contested race.
-
 ```{r}
 #| label: fig-candidate
-#| tbl-cap: Distribution of Chosen Candidate
+#| fig-cap: Distribution of Polls by Harris vs Trump
 #| echo: false
@@
   geom_bar(alpha = 0.7) +
-  labs(title = "Distribution of Polls by Harris vs Trump", 
-       x = "Chosen Candidate", 
+  labs(x = "Chosen Candidate", 
        y = "Number of Polls",
@@
   geom_bar(position = "fill") +
-  labs(title = "Candidate Preference by State", x = "State", y = "Proportion", fill = "Chosen Candidate") +
+  labs(x = "State", y = "Proportion", fill = "Chosen Candidate") +
   scale_fill_manual(values = c("red", "blue")) +
@@
 #| label: fig-sample
-#| tbl-cap: Distribution of Sample Size
+#| fig-cap: Distribution of Sample Size
 #| echo: false
@@
   geom_histogram(binwidth = 200, fill = "darkred", color = "black", alpha = 0.7) +
-  labs(title = "Distribution of Sample Size", x = "Sample Size", y = "Frequency") +
+  labs(x = "Sample Size", y = "Frequency") +
   theme_classic()
@@
   geom_density(alpha = 0.5) +
-  labs(title = "Candidate Support Distribution", x = "Percentage Support", y = "Density") +
+  labs(x = "Percentage Support", y = "Density") +
   scale_fill_manual(values = c("blue", "red"), name = "Candidate") +
@@
 
-Complementing this, [@fig-pollsters] shows the distribution of polls by individual pollsters, with bar colors indicating each pollster's Numeric Grade.
-Darker shades of blue represent higher reliability scores.
+Complementing this, [@fig-pollsters] shows the distribution of polls by individual pollsters, including only those with more than 20 counts.
+In this graph, bar colors indicate each pollster's Numeric Grade, with darker shades of blue representing higher reliability scores.
 Notably, pollsters such as Siena/NYT and Marquette Law School have the highest Numeric Grade of 3.0, marking them as highly reliable sources.
@@
 #| label: fig-numeric
-#| tbl-cap: Pollsters by Count and Numeric Grade
+#| fig-cap: Distribution of Numeric Grade
 #| echo: false
@@
   geom_text(stat = "count", aes(label = ..count..), vjust = -0.5, size = 3) +
-  labs(title = "Distribution of Numeric Grades", x = "Numeric Grade", y = "Frequency") +
+  labs(x = "Numeric Grade", y = "Frequency") +
   theme_classic()
@@
 #| label: fig-pollsters
-#| tbl-cap: Pollsters by Count and Numeric Grade
+#| fig-cap: Distribution of Pollsters by Numeric Grade 
 #| echo: false
@@
   scale_fill_gradient(low = "lightblue", high = "darkblue", name = "Numeric Grade") +
-  labs(title = "Distribution of Polls by Pollster", x = "Pollster", y = "Number of Polls") +
+  labs(x = "Pollster", y = "Number of Polls") +
   geom_text(aes(label = round(avg_numeric_grade, 2)), vjust = -0.5, size = 2.5) +
@@
 #| label: fig-numberstate
-#| tbl-cap: Number of Pollsters by State
+#| fig-cap: Number of Pollsters by State
 #| echo: false
@@
   geom_bar(fill = "darkblue") +
-  labs(title = "Number of Polls by State", x = "State", y = "Count") +
+  labs(x = "State", y = "Count") +
   theme_classic() +
@@
 
-* Percentage Support (Pct): This continuous variable is integral to the model as it captures the nuanced impact of varying levels of support for Kamala Harris.
-By utilizing percentage support as a continuous measure, the model allows for a more detailed examination of how incremental changes in support influence the likelihood of selecting Harris over Trump.
-This approach provides a richer understanding of voter behavior compared to categorizing support into discrete groups, which could obscure critical trends and insights.
+* Percentage Support (Pct): This variable is integral to the model as it captures the nuanced impact of varying levels of support for the candidates.
+By utilizing percentage support as a continuous measure, the model facilitates a more detailed examination of how incremental changes in support influence the likelihood of selecting Harris over Trump.
+This approach enhances the understanding of voter behavior compared to categorizing support into discrete groups, which could obscure critical trends and insights.
 
-* Random Effects: The decision to treat pollster and state effects as random variables recognizes the potential for unobserved heterogeneity in the data.
-Different pollsters may employ varied methodologies, which can lead to fluctuations in reported support levels that are not adequately accounted for by fixed effects alone.
-Similarly, geographical differences can impact voter preferences, making it essential to model state-level variability.
-By incorporating random effects for both pollster and state, the model effectively captures this variability, enhancing its robustness and validity.
-This approach acknowledges that the influences on voter support can differ significantly based on the context in which polling is conducted.
+* Random Effects: Treating pollster and state effects as random variables accounts for unobserved heterogeneity in the data.
+Different pollsters may use varied methodologies, resulting in fluctuations in reported support levels that fixed effects alone cannot capture.
+Additionally, geographical differences can influence voter preferences, highlighting the need to model state-level variability.
+By incorporating random effects for both pollster and state, the model effectively addresses this variability, enhancing its robustness and validity.
+This approach recognizes that the factors influencing voter support can differ significantly depending on the polling context.
 
@@
 $$
-This prior is centered around 0.5 with a standard deviation of 0.1, reflecting the belief that the probabilities of support for both candidates should be approximately 50% in the absence of data.
-This choice of prior is particularly useful in situations where there is prior knowledge or belief about the expected voter support distribution.
-By using a normal prior, we allow for flexibility in the model, enabling the incorporation of prior information while avoiding overly restrictive assumptions.
+This prior is centered at 0.5 with a standard deviation of 0.1, reflecting the belief that, in the absence of strong data, the probabilities of support for both candidates are expected to be balanced.
+This choice of prior is particularly advantageous in situations where data is limited or when there is uncertainty regarding the expected distribution of voter support.
+By employing a normal prior, we establish a baseline assumption that allows the model to remain flexible and effectively integrate new data as it becomes available.
 
@@
 The model validation process involved out-of-sample testing, where the dataset was split into training and testing sets in a 70-30 ratio.
-Both the Bayesian and logistic models utilized the same features, with the key difference being that the logistic model did not include random effects for pollster and state.
+Both the Bayesian and Logistic models utilized the same features, with the key difference being that the logistic model did not include random effects for pollster and state.
 As seen in [@tbl-comparison].
@@
 #| warning: false
-#| fig-align: center
+#| tbl-align: center
 # Create a data frame with the comparison data
@@
 
-Although the Bayesian model still shows room for improvement—evidenced by F1 Score and AUC values not close to 1 and a RMSE that indicates some prediction error—it outperforms the logistic model across all metrics.
-This suggests that the Bayesian model, which incorporates random effects to account for variability from different pollsters and states, offers a more nuanced understanding of voter support.
+Although the Bayesian model still shows room for improvement—evidenced by F1 Score and AUC values not close to 1 and a RMSE indicating some prediction error—it outperforms the logistic model across all metrics.
+This suggests that the approach, which incorporates random effects to account for variability from different pollsters and states, offers a more nuanced understanding of voter support.
 
 ### Diagnostics
-To validate the Bayesian model, several diagnostic checks were conducted to assess its robustness and ensure that underlying assumptions were satisfied.
-Key among these were model convergence assessments and posterior predictive checks.
-The posterior predictive checks provided a visual representation of how well the model's predictions align with the observed data.
-
-The graph reveals a close overlap between the observed outcomes (\(y\)) and the replicated predictions (\(y_{\text{rep}}\)), demonstrating that the model effectively captures the distribution of voter support for Harris.
-This strong alignment suggests that the model not only fits the training data well but also generalizes effectively to new observations.
-The flexibility of the Bayesian framework further allows for the incorporation of prior information and consideration of variability due to pollster and state effects, enhancing the overall robustness of the model (Refer to the Appendix for a visual illustration).
+To validate the Bayesian framework, several diagnostic checks were conducted to assess robustness and ensure that underlying assumptions were satisfied.
+Key among these were convergence assessments and posterior predictive checks.
+These checks provided a visual representation of how well the predictions align with the observed data.
+The graph reveals a close overlap between the observed outcomes (\(y\)) and the replicated predictions (\(y_{\text{rep}}\)), demonstrating effective capture of the distribution of voter support for Harris.
+This strong alignment suggests that the approach not only fits the training data well but also generalizes effectively to new observations.
+The flexibility of this framework allows for the incorporation of prior information and consideration of variability due to pollster and state effects, enhancing overall robustness (Refer to the Appendix C, [@tbl-pp] for a visual illustration).
 
@@
 # Results {#sec-myresults}
-
 ```{r}
@@
 ```
+
 ```{r}
@@
 
-# Methodology Analysis of The Washington Post Polling  
+# Appendix A: Methodology Analysis of The Washington Post Polling  
 With an evaluation of sampling methodology, recruitment, handling non-response, and questionnaire design, this appendix offers an analysis of the polling methodology used by The Washington Post in collaboration with ABC News.
@@
 
-# Idealized Survey & Methodology - $100K Budget
+# Appendix B: Idealized Survey & Methodology - $100K Budget
 
@@
 
+# Apendix C
+```{r}
+#| label: tbl-pp
+#| tbl-cap: Posterior Predictive Check: Model Fit Assessment
+#| echo: false
+#| warning: false
+#| tbl-align: center
+# Posterior predictive check
+pp_check(model_validation_train)
+```
+
 \newpage

### 2024-11-04 paper/paper.qmd
figs added- NOT RENDERING ON MINE

--- 2024-11-04 paper/paper.qmd
@@
 
-Figure X shows the national predicted popular vote percentages for Harris (50.42%) and Trump (49.58%), calculated by averaging the mean predictions across all states.
+[@fig-national-vote] shows the national predicted popular vote percentages for Harris (50.42%) and Trump (49.58%), calculated by averaging the mean predictions across all states.
 This aggregation reveals a very narrow 0.84% lead for Harris, highlighting a highly competitive race.
@@
 
-Additionally, Figure Y presents a heat map of Harris's predicted probability by state, illustrating her lead over the majority of states but revealing that this advantage is generally weak, with most not peaking over a 1% advantage.
+Additionally, [@fig-heatmap] presents a heat map of Harris's predicted probability by state, illustrating her lead over the majority of states but revealing that this advantage is generally weak, with most not peaking over a 1% advantage.
 Most states show only a slight preference for Harris over Trump, with the margin of error spanning both candidates in many cases.
-Specifically, out of the 43 states with available data, 25 states lean toward Harris, yet none of these states show lead of greater than XX over Trump.
+Specifically, out of the 43 states with available data, 25 states lean toward Harris, yet none of these states shows a substantial lead over more than a few percent over Trump.
 This pattern additionally shows the substantial closeness of the race, as evidenced by the margin of error, where overlapping intervals highlight the slim advantage and limited certainty in Harris's predicted support.
 
-Figure Z zooms in on the seven swing states, each of which shows a slight lean toward Harris.
+[@fig-swing-states] zooms in on the seven swing states, each of which shows a slight lean toward Harris.
 The graph includes a red dashed line at the 50% threshold, and although Harris is estimated to lead in these swing states, the margin is extremely tight, with most probabilities hovering just above 50%.
@@
 
-Looking closer at the margin of error, in figure X we can see Harris’s predicted probabilities by state.
+Looking closer at the margin of error, in [@fig-margin] we can see Harris’s predicted probabilities by state.
 States to the right of the dashed line (above 0.5) lean towards Harris, while those to the left lean towards Trump.
@@
 
-Figure W provides a summary of the model’s coefficients and fit statistics, capturing key model parameters like the intercepts, predictor estimates, and evaluation metrics.
+[@fig-summary-mod] provides a summary of the model’s coefficients and fit statistics, capturing key model parameters like the intercepts, predictor estimates, and evaluation metrics.
 The negative intercept of -2.07 represents the baseline log odds for Harris, establishing the prediction's starting point in the absence of other predictors.
@@
 ```{r}
-```
-
-```{r}
 #| label: fig-national-vote
 #| fig-cap: "National Popular Vote Prediction"
+#| echo: false
+#| warning: false
+#| fig-align: center
 
@@
 ```{r}
-# Plot state-level predictions for Harris and Trump
-ggplot(state_predictions, aes(x = reorder(state, avg_predicted_prob_harris), y = avg_predicted_prob_harris, fill = state_winner)) +
-  geom_bar(stat = "identity") +
-  labs(title = "State-Level Predicted Support for Harris and Trump", x = "State", y = "Average Predicted Probability for Harris") +
-  scale_fill_manual(values = c("blue", "red")) +
-  theme_classic() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-```
-```{r}
-# Modify your state predictions dataframe to label swing states
-swing_states <- c("Michigan", "Pennsylvania", "Wisconsin", "Arizona", "Georgia", "North Carolina", "Nevada")
-state_predictions <- state_predictions %>%
-  mutate(swing_state = ifelse(state %in% swing_states, "Swing State", "Non-Swing State"))
+#| label: fig-margin
+#| fig-cap: "State-Level Predicted Probability for Harris with Margin of Error"
+#| echo: false
+#| warning: false
+#| fig-align: center
 
-# Plot with enhancements
-ggplot(state_predictions, aes(x = reorder(state, avg_predicted_prob_harris), y = avg_predicted_prob_harris, fill = swing_state)) +
-  geom_bar(stat = "identity", color = "black", alpha = 0.8) +
-  scale_fill_manual(values = c("Swing State" = "darkblue", "Non-Swing State" = "lightgrey")) +
-  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red", size = 0.8) +
-  labs(
-    title = "Predicted Probability of Harris Winning by State",
-    x = "State",
-    y = "Predicted Probability for Harris",
-    fill = "State Type"
-  ) +
-  annotate("text", x = 1, y = 0.51, label = "0.5 Threshold (Neutral)", color = "red", hjust = 0) +
-  theme_classic() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-  coord_flip()
-
-```
-```{r}
-# Calculate the standard error and margin of error for Harris's predicted probability in each state
+# Calculate margin of error for each state
 state_margins <- clean_president_polls %>%
@@
   )
+
+# Plot with error bars
 ggplot(state_margins, aes(x = reorder(state, avg_predicted_prob_harris), y = avg_predicted_prob_harris)) +
@@
   labs(subtitle = "Dashed line represents a 50% threshold")
+```
+
+```{r}
+#| label: fig-swing-states
+#| fig-cap: "Predicted Probability of Harris Winning by State with Swing State Highlight and 50% Threshold"
+#| echo: false
+#| warning: false
+#| fig-align: center
+
+# Define swing states and highlight in state_predictions
+swing_states <- c("Michigan", "Pennsylvania", "Wisconsin", "Arizona", "Georgia", "North Carolina", "Nevada")
+state_predictions <- state_predictions %>%
+  mutate(swing_state = ifelse(state %in% swing_states, "Swing State", "Non-Swing State"))
+
+# Plot with threshold line at 0.5
+ggplot(state_predictions, aes(x = reorder(state, avg_predicted_prob_harris), y = avg_predicted_prob_harris, fill = swing_state)) +
+  geom_bar(stat = "identity", color = "black", alpha = 0.8) +
+  scale_fill_manual(values = c("Swing State" = "darkblue", "Non-Swing State" = "lightgrey")) +
+  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red", size = 0.8) +
+  labs(
+    title = "Predicted Probability of Harris Winning by Swing State",
+    x = "State",
+    y = "Predicted Probability for Harris",
+    fill = "State Type"
+  ) +
+  annotate("text", x = 1, y = 0.51, label = "0.5 Threshold (Neutral)", color = "red", hjust = 0) +
+  theme_classic() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  coord_flip()
 
@@
 
+```{r}
+#| label: fig-heatmap
+#| fig-cap: "Heatmap of Predicted Support for Harris by State"
+#| echo: false
+#| warning: false
+#| fig-align: center
+
+# Ensure `state_predictions` has the necessary data
+state_predictions <- clean_president_polls %>%
+  group_by(state) %>%
+  summarize(avg_predicted_prob_harris = mean(predicted_prob_harris)) %>%
+  mutate(state_winner = ifelse(avg_predicted_prob_harris > 0.5, "Harris", "Trump"))
+
+# Convert state names to lowercase for matching with `maps` package data
+state_predictions$state <- tolower(state_predictions$state)
+
+# Load USA state map data and merge with predictions
+usa_states <- map_data("state")
+merged_data <- merge(usa_states, state_predictions, by.x = "region", by.y = "state", sort = FALSE, all.x = TRUE)
+merged_data <- merged_data[order(merged_data$group, merged_data$order), ]
+
+# Plot heatmap for Harris's support by state, with NA values as dark grey
+ggplot(merged_data, aes(x = long, y = lat, group = group, fill = avg_predicted_prob_harris)) +
+  geom_polygon(color = "black", size = 0.3) + 
+  scale_fill_gradient(low = "lightblue", high = "darkblue", na.value = "darkgrey", name = "Predicted Support") +
+  coord_fixed(1.3) + 
+  labs(
+    title = "Predicted Support for Harris by State",
+    x = NULL,
+    y = NULL
+  ) +
+  theme_void() +
+  theme(
+    plot.title = element_text(hjust = 0.5, size = 16, face = "bold")
+  )
+```
 # Discussion {#sec-mydisc}
@@
 In the 7 swing states– Arizona, Wisconsin, North Carolina, Pennsylvania, Georgia, Michigan and Nevada, our model shows Harris leading in each, with only 2 swing states showing a lead greater than 1%.
-Furthermore, the small margins seen in FUGURENARGIN show swing state leads ithin their respective margin of errors, pushing the uncertainty of the estimated lead.
+Furthermore, the small margins seen in @figure-margin show swing state leads ithin their respective margin of errors, pushing the uncertainty of the estimated lead.
 Other forecasts like FiveThirtyEight’s and The NYTimes’ offer similar leads for Harris [@FiveThirtyEight; nytimes2024_estimate].
@@
 Non-response can introduce non-response bias, where the views of non-respondents may differ from respondents, potentially skewing results.
-The Washington Post addresses this issue using response weighting based on age, race, and education to align the final sample with the overall population distribution [@WashPost2023_Standard].
+The Washington Post addresses this issue using response weighting based on age, race, and education to align the final sample with the overall population distribution [@WashPost2023_Standards].
 

### 2024-11-04 paper/paper.qmd

--- 2024-11-04 paper/paper.qmd
@@
 # Results {#sec-myresults}
-
 Figure X shows the national predicted popular vote percentages for Harris (50.42%) and Trump (49.58%), calculated by averaging the mean predictions across all states.
@@
 ```{r}
+#| label: tbl-summary-mod
+#| tbl-cap: "Summary of Bayesian Model"
+#| echo: false
+#| warning: false
+#| tbl-align: center
+
+# Display model summary with customized formatting
+modelsummary(bayesian_model, 
+             output = "kableExtra")
+
+# Create a custom data frame with the results
+custom_summary <- data.frame(
+  Parameter = c("Intercept", 
+                "Percentage Support", 
+                "Pollster Variability", 
+                "State Variability", 
+                "Number of Observations", 
+                "R²", 
+                "Marginal R²", 
+                "Intraclass Correlation (ICC)", 
+                "Log-Likelihood", 
+                "WAIC", 
+                "RMSE"),
+  Estimate = c(-3.779, 
+               0.081, 
+               0.001, 
+               0.001, 
+               2367, 
+               0.027, 
+               0.026, 
+               0.0, 
+               -3849.195, 
+               7728.2, 
+               0.50)
+)
+
+kable(custom_summary)
+```
+
+```{r}
 #| label: fig-summary-mod
-#| tbl-cap: "Distribution of Sample Size"
+#| tbl-cap: "Summary of Bayesian Model"
 #| echo: false
@@
 #| fig-align: center
-
-# Display model summary
-modelsummary(bayesian_model, output = "kableExtra")
-```
-
-```{r}
-```
-
-```{r}
-#| label: fig-national-vote
-#| fig-cap: "National Popular Vote Prediction"
-
 # Calculate overall percentage for Harris and Trump
@@
 Addressing this issue in future analyses could enhance interpretability and offer a clearer comparison to the Bayesian and other models.
+
 Additionally, environmental disruptions, such as recent hurricanes in key swing states, may introduce biases.
@@
 While this approach captures broader state trends, it could underrepresent district-specific voting patterns, and future studies could address district-level data.
+
 The model is fairly simple in comparison to other forecasts and does not currently adjust for other traditional variables or bias (i.e.
@@
 Incorporating adjustments for pollster partisanship would improve accuracy by reflecting a more comprehensive cross-section of public opinion.
+
 Another simplification in this analysis is the treatment of Maine and Nebraska, where data were aggregated at the state level instead of by congressional district.
@@
 This simplification streamlines the model but may underrepresent nuanced trends within these states.
+
 Our method of weighting the polls does not have any constraints and could lead to the overweighting of certain data groups.

### 2024-11-04 paper/paper.qmd

@@
 library(modelsummary)
+library(rstanarm)
 ```
@@
 bayesian_model <- readRDS("../models/first_model.rds")
-
 ```
@@
 $$
-\begin{align*}
+\begin{align}
 \text{logit}(y_i) &= \beta_0 + \beta_1 \times \text{Pct}_i + u_{\text{pollster}} + u_{\text{state}} \\
-\end{align*}
+\end{align}
 $$
@@
 * Component Explanation
-$$
-- $y_i$: Binary outcome variable indicating whether a respondent supports Harris (1) or not (0).
-- $\beta_0$: Intercept term representing the baseline log-odds of support for Harris when the percentage support is zero.
-- $\beta_1$: Coefficient for the percentage support for Harris (Pct), reflecting how changes in support percentage affect the log-odds of selecting Harris.
-- $u_{\text{pollster}}$: Random effect for the specific pollster, accounting for variability in support across different pollsters.
+- $y_i$: Binary outcome variable indicating whether a respondent supports Harris (1) or not (0). \\
+- $\beta_0$:Intercept term representing the baseline log-odds of support for Harris when the percentage support is zero. \\
+- $\beta_1$: Coefficient for the percentage support for Harris (Pct), reflecting how changes in support percentage affect the log-odds of selecting Harris. \\
+- $u_{\text{pollster}}$:Random effect for the specific pollster, accounting for variability in support across different pollsters. \\
 - $u_{\text{state}}$: Random effect for the state in which the poll is conducted, capturing regional variations in support.
-$$
 
@@
 $$
+\begin{equation}
 \text{weights} = \text{Numeric Grade} \times \left(\frac{\text{Sample Size}}{\text{Mean(Sample Size)}}\right)
+\end{equation}
 $$
@@
 $$
+\begin{equation}
 \text{priors} \sim \text{Normal}(0.5, 0.1)
+\end{equation}
 $$
@@
 This strong alignment suggests that the approach not only fits the training data well but also generalizes effectively to new observations.
-The flexibility of this framework allows for the incorporation of prior information and consideration of variability due to pollster and state effects, enhancing overall robustness (Refer to the Appendix C, [@tbl-pp] for a visual illustration).
+The flexibility of this framework allows for the incorporation of prior information and consideration of variability due to pollster and state effects, enhancing overall robustness (Refer to the @sec-appc, [@tbl-pp] for a visual illustration).
 
@@
 # Results {#sec-myresults}
-Figure X shows the national predicted popular vote percentages for Harris (50.42%) and Trump (49.58%), calculated by averaging the mean predictions across all states.
-This aggregation reveals a very narrow 0.84% lead for Harris, highlighting a highly competitive race.
+```{r}
+#| include: false
+#| warning: false
+#| message: false
+# Add weights to the dataset
+clean_president_polls <- clean_president_polls %>%
+  mutate(weight = (numeric_grade * (sample_size / mean(sample_size))),
+         pollster = as.factor(pollster),
+         state = as.factor(state)
+         )
+
+#### Bayesian Model ####
+# Define priors
+priors <- normal(0.5, 0.1, autoscale = TRUE)
+
+# Fit model with complete dataset
+set.seed(123)
+bayesian_model <- stan_glmer(
+  formula = candidate_chosen ~ pct + (1 | pollster) + (1 | state),
+  data = clean_president_polls,
+  family = binomial(link = "logit"),
+  prior = priors,
+  prior_intercept = priors,
+  seed = 123,
+  cores = 4,
+  weights = weight,
+  adapt_delta = 0.95
+)
+
+# Predict probabilities for Harris winning 
+clean_president_polls <- clean_president_polls |> 
+  mutate(
+    predicted_prob_harris = posterior_predict(bayesian_model, newdata = clean_president_polls, type = "response") |> colMeans(),
+    winner_harris = ifelse(predicted_prob_harris > 0.5 & candidate_chosen == 1, 1, 0)
+  )
+# Calculate the unweighted average predicted probability for Kamala Harris
+overall_predicted_prob_harris <- mean(clean_president_polls$predicted_prob_harris)
+
+# Convert to percentage
+overall_percentage_harris <- overall_predicted_prob_harris * 100
+overall_percentage_trump <- (1 - overall_predicted_prob_harris) * 100
+
+# Print the overall results
+cat("Overall Percentage for Kamala Harris:", overall_percentage_harris, "%\n")
+cat("Overall Percentage for Donald Trump:", overall_percentage_trump, "%\n")
+
+# Calculate the average predicted probability by state for Harris
+state_predictions <- clean_president_polls %>%
+  group_by(state) %>%
+  summarize(
+    avg_predicted_prob_harris = mean(predicted_prob_harris)
+  )
+
+# Determine the winner for each state based on the average probability
+state_predictions <- state_predictions %>%
+  mutate(
+    state_winner = ifelse(avg_predicted_prob_harris > 0.5, "Harris", "Trump")
+  )
+
+# Count the number of states won by each candidate
+overall_winner_summary <- state_predictions %>%
+  count(state_winner)
+
+# View the state-level predictions
+print(state_predictions, n = Inf)
+
+# View the summary of states won by each candidate
+print(overall_winner_summary)
+```
+
+[@fig-national-vote] shows the national predicted popular vote percentages for Harris (50.44%) and Trump (49.56%), calculated by averaging the mean predictions across all states.
+This aggregation reveals a very narrow 0.88% lead for Harris, highlighting a highly competitive race.
 This reveals a lack of dominance in candidacy and general uncertainty of pollster outcomes, suggesting that even this slight lead may not be substantial.
-
-Additionally, Figure Y presents a heat map of Harris's predicted probability by state, illustrating her lead over the majority of states but revealing that this advantage is generally weak, with most not peaking over a 1% advantage.
+```{r}
+#| label: fig-national-vote
+#| fig-cap: "National Popular Vote Prediction"
+#| echo: false
+#| warning: false
+#| fig-align: center
+
+# Calculate overall percentage for Harris and Trump
+overall_percentages <- data.frame(
+  Candidate = c("Harris", "Trump"),
+  Percentage = c(overall_percentage_harris, overall_percentage_trump)
+)
+
+# Plot national popular vote prediction
+ggplot(overall_percentages, aes(x = Candidate, y = Percentage, fill = Candidate)) +
+  geom_bar(stat = "identity") +
+  geom_text(aes(label = sprintf("%.2f%%", Percentage)), vjust = -0.5) +
+  labs(y = "Percentage", x = "Candidate") +
+  scale_fill_manual(values = c("blue", "red")) +
+  theme_classic()
+```
+
+Additionally, [@fig-heatmap] presents a heat map of Harris's predicted probability by state, illustrating her lead over the majority of states but revealing that this advantage is generally weak, with most not peaking over a 1% advantage.
 Most states show only a slight preference for Harris over Trump, with the margin of error spanning both candidates in many cases.
-Specifically, out of the 43 states with available data, 25 states lean toward Harris, yet none of these states show lead of greater than XX over Trump.
+Specifically, out of the 43 states with available data, 25 states lean toward Harris, yet none of these states shows a substantial lead over more than a few percent over Trump.
 This pattern additionally shows the substantial closeness of the race, as evidenced by the margin of error, where overlapping intervals highlight the slim advantage and limited certainty in Harris's predicted support.
-
-Figure Z zooms in on the seven swing states, each of which shows a slight lean toward Harris.
+```{r}
+#| label: fig-heatmap
+#| fig-cap: "Heatmap of Predicted Support for Harris by State"
+#| echo: false
+#| warning: false
+#| fig-align: center
+
+# Ensure `state_predictions` has the necessary data
+state_predictions <- clean_president_polls %>%
+  group_by(state) %>%
+  summarize(avg_predicted_prob_harris = mean(predicted_prob_harris)) %>%
+  mutate(state_winner = ifelse(avg_predicted_prob_harris > 0.5, "Harris", "Trump"))
+
+# Convert state names to lowercase for matching with `maps` package data
+state_predictions$state <- tolower(state_predictions$state)
+
+# Load USA state map data and merge with predictions
+usa_states <- map_data("state")
+merged_data <- merge(usa_states, state_predictions, by.x = "region", by.y = "state", sort = FALSE, all.x = TRUE)
+merged_data <- merged_data[order(merged_data$group, merged_data$order), ]
+
+# Plot heatmap for Harris's support by state, with NA values as dark grey
+ggplot(merged_data, aes(x = long, y = lat, group = group, fill = avg_predicted_prob_harris)) +
+  geom_polygon(color = "black", size = 0.3) + 
+  scale_fill_gradient(low = "lightblue", high = "blue", na.value = "darkgrey", name = "Predicted Support") +
+  coord_fixed(1.3) + 
+  labs(
+    x = NULL,
+    y = NULL
+  ) +
+  theme_void() +
+  theme(
+    plot.title = element_text(hjust = 0.5, size = 16, face = "bold")
+  )
+```
+
+[@fig-swing-states] zooms in on the seven swing states, each of which shows a slight lean toward Harris.
 The graph includes a red dashed line at the 50% threshold, and although Harris is estimated to lead in these swing states, the margin is extremely tight, with most probabilities hovering just above 50%.
@@
 This pattern suggests that even small shifts in voter sentiment could tip these critical states in either direction.
-
-Looking closer at the margin of error, in figure X we can see Harris’s predicted probabilities by state.
+```{r}
+#| label: fig-swing-states
+#| fig-cap: "Predicted Probability of Harris Winning by State with Swing State Highlight and 50% Threshold"
+#| echo: false
+#| warning: false
+#| fig-align: center
+
+# Modify your state predictions dataframe to label swing states
+swing_states <- c("michigan", "pennsylvania", "wisconsin", "arizona", "georgia", "north carolina", "nevada")
+state_predictions <- state_predictions %>%
+  mutate(swing_state = ifelse(state %in% swing_states, "Swing State", "Non-Swing State"))
+
+# Plot with enhancements
+ggplot(state_predictions, aes(x = reorder(state, avg_predicted_prob_harris), y = avg_predicted_prob_harris, fill = swing_state)) +
+  geom_bar(stat = "identity", color = "black", alpha = 0.8) +
+  scale_fill_manual(values = c("Swing State" = "darkblue", "Non-Swing State" = "lightgrey")) +
+  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red", size = 0.8) +
+  labs(
+    title = "Predicted Probability of Harris Winning by State",
+    x = "State",
+    y = "Predicted Probability for Harris",
+    fill = "State Type"
+  ) +
+  annotate("text", x = 1, y = 0.51, label = "0.5 Threshold (Neutral)", color = "red", hjust = 0) +
+  theme_classic() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  coord_flip()
+```
+
+Looking closer at the margin of error, in [@fig-margin] we can see Harris’s predicted probabilities by state.
 States to the right of the dashed line (above 0.5) lean towards Harris, while those to the left lean towards Trump.
@@
 Harris has a slight edge, but in many states, this lead is within the margin of error, thus showing that the election remains too close to call.
-
-[@fig-summary-mod] provides a summary of the model’s coefficients and fit statistics, capturing key model parameters like the intercepts, predictor estimates, and evaluation metrics.
+```{r}
+#| label: fig-margin
+#| fig-cap: "State-Level Predicted Probability for Harris with Margin of Error"
+#| echo: false
+#| warning: false
+#| fig-align: center
+
+# Calculate margin of error for each state
+state_margins <- clean_president_polls %>%
+  group_by(state) %>%
+  summarize(
+    avg_predicted_prob_harris = mean(predicted_prob_harris),
+    se_predicted_prob_harris = sd(predicted_prob_harris) / sqrt(n()),
+    margin_of_error = 1.96 * se_predicted_prob_harris
+  ) %>%
+  mutate(
+    lower_bound = avg_predicted_prob_harris - margin_of_error,
+    upper_bound = avg_predicted_prob_harris + margin_of_error
+  )
+
+# Plot with error bars
+ggplot(state_margins, aes(x = reorder(state, avg_predicted_prob_harris), y = avg_predicted_prob_harris)) +
+  geom_point(color = "blue") +
+  geom_errorbar(aes(ymin = lower_bound, ymax = upper_bound), width = 0.2, color = "blue") +
+  labs(
+    x = "State",
+    y = "Predicted Probability for Harris"
+  ) +
+  coord_flip() +
+  theme_classic() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red") +
+  labs(subtitle = "Dashed line represents a 50% threshold")
+```
+
+[@tbl-summary-mod] provides a summary of the model’s coefficients and fit statistics, capturing key model parameters like the intercepts, predictor estimates, and evaluation metrics.
 The negative intercept of -2.07 represents the baseline log odds for Harris, establishing the prediction's starting point in the absence of other predictors.
@@
 
-```{r}
-#| label: fig-summary-mod
-#| tbl-cap: "Summary of Bayesian Model"
-#| echo: false
-#| warning: false
-#| fig-align: center
-# Calculate overall percentage for Harris and Trump
-overall_percentages <- data.frame(
-  Candidate = c("Harris", "Trump"),
-  Percentage = c(overall_percentage_harris, overall_percentage_trump)
-)
-
-# Plot national popular vote prediction
-ggplot(overall_percentages, aes(x = Candidate, y = Percentage, fill = Candidate)) +
-  geom_bar(stat = "identity") +
-  geom_text(aes(label = sprintf("%.2f%%", Percentage)), vjust = -0.5) +
-  labs(title = "National Popular Vote Prediction", y = "Percentage", x = "Candidate") +
-  scale_fill_manual(values = c("blue", "red")) +
-  theme_classic()
-
-```
-
-```{r}
-# Plot state-level predictions for Harris and Trump
-ggplot(state_predictions, aes(x = reorder(state, avg_predicted_prob_harris), y = avg_predicted_prob_harris, fill = state_winner)) +
-  geom_bar(stat = "identity") +
-  labs(title = "State-Level Predicted Support for Harris and Trump", x = "State", y = "Average Predicted Probability for Harris") +
-  scale_fill_manual(values = c("blue", "red")) +
-  theme_classic() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-```
-
-```{r}
-# Modify your state predictions dataframe to label swing states
-swing_states <- c("Michigan", "Pennsylvania", "Wisconsin", "Arizona", "Georgia", "North Carolina", "Nevada")
-state_predictions <- state_predictions %>%
-  mutate(swing_state = ifelse(state %in% swing_states, "Swing State", "Non-Swing State"))
-
-# Plot with enhancements
-ggplot(state_predictions, aes(x = reorder(state, avg_predicted_prob_harris), y = avg_predicted_prob_harris, fill = swing_state)) +
-  geom_bar(stat = "identity", color = "black", alpha = 0.8) +
-  scale_fill_manual(values = c("Swing State" = "darkblue", "Non-Swing State" = "lightgrey")) +
-  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red", size = 0.8) +
-  labs(
-    title = "Predicted Probability of Harris Winning by State",
-    x = "State",
-    y = "Predicted Probability for Harris",
-    fill = "State Type"
-  ) +
-  annotate("text", x = 1, y = 0.51, label = "0.5 Threshold (Neutral)", color = "red", hjust = 0) +
-  theme_classic() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-  coord_flip()
-
-```
-
-```{r}
-# Calculate the standard error and margin of error for Harris's predicted probability in each state
-state_margins <- clean_president_polls %>%
-  group_by(state) %>%
-  summarize(
-    avg_predicted_prob_harris = mean(predicted_prob_harris),
-    se_predicted_prob_harris = sd(predicted_prob_harris) / sqrt(n()),
-    margin_of_error = 1.96 * se_predicted_prob_harris
-  ) %>%
-  mutate(
-    lower_bound = avg_predicted_prob_harris - margin_of_error,
-    upper_bound = avg_predicted_prob_harris + margin_of_error
-  )
-ggplot(state_margins, aes(x = reorder(state, avg_predicted_prob_harris), y = avg_predicted_prob_harris)) +
-  geom_point(color = "blue") +
-  geom_errorbar(aes(ymin = lower_bound, ymax = upper_bound), width = 0.2, color = "blue") +
-  labs(
-    title = "State-Level Predicted Probability for Harris with Margin of Error",
-    x = "State",
-    y = "Predicted Probability for Harris"
-  ) +
-  coord_flip() +
-  theme_classic() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red") +
-  labs(subtitle = "Dashed line represents a 50% threshold")
-
-```
-
 # Discussion {#sec-mydisc}
-
 In a highly competitive election, accurately forecasting outcomes requires more than just raw polling data—it demands careful aggregation and weighting to capture true voter sentiment. 
@@
 
-While this suggests a slight edge for Harris, the lead easily falls within a \pm 3\% margin that reflects substantial uncertainty [@PewResearch_MarginOfError].
+While this suggests a slight edge for Harris, the lead easily falls within a $\pm 3\%$ margin that reflects substantial uncertainty [@PewResearch_MarginOfError].
 When a candidate’s lead is within this margin, the difference may not be statistically significant, meaning the true preference could feasibly favor either candidate.
@@
 In the 7 swing states– Arizona, Wisconsin, North Carolina, Pennsylvania, Georgia, Michigan and Nevada, our model shows Harris leading in each, with only 2 swing states showing a lead greater than 1%.
-Furthermore, the small margins seen in FUGURENARGIN show swing state leads ithin their respective margin of errors, pushing the uncertainty of the estimated lead.
-Other forecasts like FiveThirtyEight’s and The NYTimes’ offer similar leads for Harris [@FiveThirtyEight; nytimes2024_estimate].
+Furthermore, the small margins seen in [@fig-margin] show swing state leads ithin their respective margin of errors, pushing the uncertainty of the estimated lead.
+Other forecasts like FiveThirtyEight’s and The NYTimes’ offer similar leads for Harris [@FiveThirtyEight; @nytimes2024_estimate].
 When independent models using varied methodologies predict a similar outcome, it can suggest a trend that might hold under different sampling assumptions or adjustments.
@@
 
-# Apendix C
+# Appendix C {#sec-appc}
+```{r}
+#| include: false
+#| warning: false
+#| message: false
+### Model Validation/Checking ###
+# Set seed and create train/test split
+set.seed(123)
+train_indices <- sample(seq_len(nrow(clean_president_polls)), size = 0.7 * nrow(clean_president_polls))
+training_data <- clean_president_polls[train_indices, ]
+training_data <- training_data %>%
+  mutate(pct_scaled = scale(pct))
+testing_data <- clean_president_polls[-train_indices, ]
+
+# Ensure test data has compatible pollster levels with training data
+testing_data <- testing_data |>
+  filter(pollster %in% unique(training_data$pollster))
+
+# Fit `Bayesian model` with training dataset
+model_validation_train <- stan_glmer(
+  formula = candidate_chosen ~ pct + (1 | pollster) + (1 | state),
+  data = training_data,
+  family = binomial(link = "logit"),
+  prior = priors,
+  prior_intercept = priors,
+  weights = weight,
+  cores = 4,
+  adapt_delta = 0.99,
+  seed = 123
+)
+```
+
 ```{r}
 #| label: tbl-pp
-#| tbl-cap: Posterior Predictive Check: Model Fit Assessment
+#| tbl-cap: Posterior Predictive Check for Model Fit Assessment
 #| echo: false

### 2024-11-04 paper/paper.qmd
finale

@@
 * Mathematical Representation
-$$
+
 \begin{align}
@@
 \end{align}
-$$
 
@@
 The weights in this model are calculated using the formula:
-$$
+
 \begin{equation}
@@
 \end{equation}
-$$
+
 This approach ensures that more reliable polls—indicated by a higher Numeric Grade—carry greater influence in the analysis.
@@
 Regarding the priors, we define them as follows:
-$$
+
 \begin{equation}
@@
 \end{equation}
-$$
+
 This prior is centered at 0.5 with a standard deviation of 0.1, reflecting the belief that, in the absence of strong data, the probabilities of support for both candidates are expected to be balanced.
@@
 # Limitations and Next Steps {#sec-mylimits}
+## Limitations
 This analysis identifies several limitations that may impact the reliability of the predictions.
@@
 If one sample size was much bigger than the majority, its weight would be extreme as well, potentially overcompensating for its affects.
+
 Finally, data gaps remain a significant factor, as polling data for several states were unavailable.
@@
 This sparsity in effect with Harris’ recent campaign declaration, contributes to an overall margin of error.
-Future Directions
+
+## Next Steps
 Future models could benefit from greater complexity to address these limitations.

### 2024-11-05 paper/paper.qmd
finish

@@
 The rest of this paper is structured as follows:  @sec-mydatasection presents the data sources and methodology, @sec-mymodel presents the forecasting model, then  @sec-myresults followed by  @sec-mydisc, which discusses the results and their implications.
-@sec-mylimits concludes with a discussion of the limitations of the study and suggestions for future research, followed by @sec-appenx with The Washington Post’s pollster methodology and a budgeted idealized methodology.
+@sec-mylimits concludes with a discussion of the limitations of the study and suggestions for future research, followed by @sec-appenxA with The Washington Post’s pollster methodology, and @sec-appenxB a budgeted idealized methodology.
 
@@
 ## Outcome Variable
-### Distribution of Polls and State-Level Candidate Preference
 The chosen candidate—whether a poll favors Harris or Trump—serves as the outcome variable in this analysis.
@@
 #| fig-align: center
+#| fig-width: 5
+#| fig-height: 3
 ggplot(clean_president_polls, aes(x = factor(candidate_chosen), fill = factor(candidate_chosen))) +
@@
   geom_text(stat = "count", aes(label = scales::percent(..count../sum(..count..), 
-                                                        accuracy = 0.2)), vjust = -0.5) 
+                                                        accuracy = 0.2)), vjust = -0.5, size = 3)  +
+  theme(
+    axis.title.x = element_text(size = 8),
+    axis.title.y = element_text(size = 8),
+    legend.text = element_text(size = 6), 
+    legend.title = element_text(size = 6)
+  )
 ```
@@
 #| label: fig-candbystate
-#| tbl-cap: Distribution of Candidate Preference by State
-#| echo: false
-#| warning: false
-#| fig-align: center
+#| fig-cap: Distribution of Candidate Preference by State
+#| echo: false
+#| warning: false
+#| fig-align: center
+#| fig-width: 5
+#| fig-height: 3
 ggplot(clean_president_polls, aes(x = state, fill = factor(candidate_chosen, labels = c("Trump", "Harris")))) +
@@
   theme(
-    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, margin = margin(t = 5), size = 8)
-  )
-```
+    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, margin = margin(t = 5), size = 6),
+    axis.title.x = element_text(size = 8),
+    axis.title.y = element_text(size = 8),
+    legend.text = element_text(size = 6), 
+    legend.title = element_text(size = 6)
+  )
+```
+\newpage
 
@@
 #| fig-align: center
+#| fig-width: 5
+#| fig-height: 3
 ggplot(clean_president_polls, aes(x = sample_size)) +
@@
   labs(x = "Sample Size", y = "Frequency") +
-  theme_classic()
+  theme_classic() +
+  theme(
+    axis.title.x = element_text(size = 8),
+    axis.title.y = element_text(size = 8)
+  )
 ```
@@
 #| label: fig-percentage
-#| tbl-cap: Distribution of Candidate Support by Chosen Candidate
-#| echo: false
-#| warning: false
-#| fig-align: center
+#| fig-cap: Distribution of Candidate Support by Chosen Candidate
+#| echo: false
+#| warning: false
+#| fig-align: center
+#| fig-width: 5
+#| fig-height: 3
 ggplot(clean_president_polls, aes(x = pct, fill = candidate_name)) +
@@
   scale_fill_manual(values = c("blue", "red"), name = "Candidate") +
-  theme_classic()
+  theme_classic() +
+  theme(
+    axis.text.x = element_text(hjust = 1, size = 4), 
+    axis.title.x = element_text(size = 8),
+    axis.title.y = element_text(size = 8),
+    legend.text = element_text(size = 6), 
+    legend.title = element_text(size = 6)
+  )
 ```
@@
 #| fig-align: center
+#| fig-width: 5
+#| fig-height: 3
 ggplot(clean_president_polls, aes(x = numeric_grade)) +
   geom_bar(fill = "darkred", alpha = 0.7) +
-  geom_text(stat = "count", aes(label = ..count..), vjust = -0.5, size = 3) +
+  geom_text(stat = "count", aes(label = ..count..), vjust = -0.5, size = 2) +
   labs(x = "Numeric Grade", y = "Frequency") +
-  theme_classic()
+  theme_classic() +
+  theme(
+    axis.title.x = element_text(size = 8),
+    axis.title.y = element_text(size = 8)
+  )
 ```
@@
 #| fig-align: center
+#| fig-width: 5
+#| fig-height: 3
 # Calculate the number of polls for each pollster with more than 20 polls
@@
   group_by(pollster) %>%
-  filter(n() > 20) %>%
+  filter(n() > 30) %>%
   summarize(
@@
   theme_classic() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8))
-```
+  theme(
+    axis.text.x = element_text(angle = 45, hjust = 1, size = 4), 
+    axis.title.x = element_text(size = 8),
+    axis.title.y = element_text(size = 8),
+    legend.text = element_text(size = 6), 
+    legend.title = element_text(size = 6)
+  )
+```
+\newpage
 
@@
 #| fig-align: center
+#| fig-width: 5
+#| fig-height: 3
 ggplot(clean_president_polls, aes(x = reorder(state, -table(state)[state]))) +
-  geom_bar(fill = "darkblue") +
+  geom_bar(fill = "darkred") +
   labs(x = "State", y = "Count") +
   theme_classic() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 4),
+        axis.text.y = element_text(hjust = 1, size = 6),
+    axis.title.x = element_text(size = 8),
+    axis.title.y = element_text(size = 8)
+  )
 ```
@@
 #| label: fig-support
-#| tbl-cap: Candidate with Highest Support Percentage by State
-#| echo: false
-#| warning: false
-#| fig-align: center
+#| fig-cap: Candidate with Highest Support Percentage by State
+#| echo: false
+#| warning: false
+#| fig-align: center
+#| fig-width: 5
+#| fig-height: 3
 # Summarize data to find the candidate with the highest percentage in each state
@@
   coord_fixed(1.3) +
-  labs(title = "Leading Candidate by State") +
   theme_void() +
@@
     legend.position = "right",
-    plot.title = element_text(hjust = 0.5, size = 16, face = "bold")
-  )
-```
+    legend.text = element_text(size = 6), 
+    legend.title = element_text(size = 6)
+  )
+```
+\newpage
 
@@
 * Component Explanation
-- $y_i$: Binary outcome variable indicating whether a respondent supports Harris (1) or not (0). \\
-- $\beta_0$:Intercept term representing the baseline log-odds of support for Harris when the percentage support is zero. \\
-- $\beta_1$: Coefficient for the percentage support for Harris (Pct), reflecting how changes in support percentage affect the log-odds of selecting Harris. \\
-- $u_{\text{pollster}}$:Random effect for the specific pollster, accounting for variability in support across different pollsters. \\
+- $y_i$: Binary outcome variable indicating whether a respondent supports Harris (1) or not (0). 
+- $\beta_0$:Intercept term representing the baseline log-odds of support for Harris when the percentage support is zero. 
+- $\beta_1$: Coefficient for the percentage support for Harris (Pct), reflecting how changes in support percentage affect the log-odds of selecting Harris. 
+- $u_{\text{pollster}}$:Random effect for the specific pollster, accounting for variability in support across different pollsters. 
 - $u_{\text{state}}$: Random effect for the state in which the poll is conducted, capturing regional variations in support.
@@
 These checks provided a visual representation of how well the predictions align with the observed data.
-The graph reveals a close overlap between the observed outcomes (\(y\)) and the replicated predictions (\(y_{\text{rep}}\)), demonstrating effective capture of the distribution of voter support for Harris.
+The graph reveals a close overlap between the observed outcomes \(y\) and the replicated predictions ($y_{\text{rep}}$), demonstrating effective capture of the distribution of voter support for Harris.
 This strong alignment suggests that the approach not only fits the training data well but also generalizes effectively to new observations.
@@
 
-[@fig-national-vote] shows the national predicted popular vote percentages for Harris (50.44%) and Trump (49.56%), calculated by averaging the mean predictions across all states.
-This aggregation reveals a very narrow 0.88% lead for Harris, highlighting a highly competitive race.
+The analysis of the national predicted popular vote percentages indicates that Kamala Harris is projected to receive 50.44% of the vote, while Donald Trump is expected to garner 49.56%.
+These figures, illustrated in [@fig-national-vote] and calculated by averaging the mean predictions across all states, reveal a narrow lead of just 0.88% for Harris, highlighting a highly competitive race.
 This reveals a lack of dominance in candidacy and general uncertainty of pollster outcomes, suggesting that even this slight lead may not be substantial.
@@
 #| fig-align: center
-
+#| fig-width: 5
+#| fig-height: 3
 # Calculate overall percentage for Harris and Trump
@@
   geom_bar(stat = "identity") +
-  geom_text(aes(label = sprintf("%.2f%%", Percentage)), vjust = -0.5) +
+  geom_text(aes(label = sprintf("%.2f%%", Percentage)), vjust = -0.5, size = 3) +
   labs(y = "Percentage", x = "Candidate") +
   scale_fill_manual(values = c("blue", "red")) +
-  theme_classic()
+  theme_classic() +
+  theme(
+    axis.title.x = element_text(size = 8),
+    axis.title.y = element_text(size = 8),
+    legend.text = element_text(size = 6), 
+    legend.title = element_text(size = 6)
+  )
 ```
@@
 #| fig-align: center
+#| fig-width: 5
+#| fig-height: 3
 
@@
   theme(
-    plot.title = element_text(hjust = 0.5, size = 16, face = "bold")
+    legend.text = element_text(size = 6), 
+    legend.title = element_text(size = 6)
   )
@@
 #| fig-align: center
+#| fig-width: 6
+#| fig-height: 4
 
@@
 
-# Plot with enhancements
+# Plot 
 ggplot(state_predictions, aes(x = reorder(state, avg_predicted_prob_harris), y = avg_predicted_prob_harris, fill = swing_state)) +
@@
   labs(
-    title = "Predicted Probability of Harris Winning by State",
     x = "State",
@@
   ) +
-  annotate("text", x = 1, y = 0.51, label = "0.5 Threshold (Neutral)", color = "red", hjust = 0) +
+  annotate("text", x = 1, y = 0.51, label = "0.5", color = "red", hjust = 0, size = 4) +
   theme_classic() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  theme(
+    axis.text.y = element_text(size = 5),
+    axis.title.x = element_text(size = 8),
+    axis.title.y = element_text(size = 8),
+    legend.text = element_text(size = 6), 
+    legend.title = element_text(size = 6)
+  ) +
   coord_flip()
 ```
-
+\newpage
 Looking closer at the margin of error, in [@fig-margin] we can see Harris’s predicted probabilities by state.
@@
 #| fig-align: center
-
+#| fig-width: 6
+#| fig-height: 4
 # Calculate margin of error for each state
@@
   theme_classic() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
   geom_hline(yintercept = 0.5, linetype = "dashed", color = "red") +
-  labs(subtitle = "Dashed line represents a 50% threshold")
+  labs(subtitle = "Dashed line represents a 50% threshold") +
+  theme(plot.subtitle = element_text(hjust = 0.5),
+    axis.text.y = element_text(size = 4),
+    axis.title.x = element_text(size = 8),
+    axis.title.y = element_text(size = 8)
+    ) 
 ```
@@
 #| tbl-align: center
-
-# Display model summary with customized formatting
-modelsummary(bayesian_model, 
-             output = "kableExtra")
+#| fig-width: 6
+#| fig-height: 4
 
@@
 
-\appendix
-
-# Appendix {#sec-appenx}
-
-# Appendix A: Methodology Analysis of The Washington Post Polling  
+# Appendix A: Methodology Analysis of The Washington Post Polling  {#sec-appenxA}
 With an evaluation of sampling methodology, recruitment, handling non-response, and questionnaire design, this appendix offers an analysis of the polling methodology used by The Washington Post in collaboration with ABC News.
@@
 
-# Appendix B: Idealized Survey & Methodology - $100K Budget
+# Appendix B: Idealized Survey & Methodology - $100K Budget {#sec-appenxB}
 
@@
 The design aims to balance accuracy, inclusivity, and efficiency within a $100K budget, offering meaningful insights into voter behavior and sentiment.
+\newpage
 
 # Appendix C {#sec-appc}
+* [@tbl-pp]
 ```{r}

### 2024-11-13 paper/paper.qmd
appendix formatting - NOT rendered as pdf

@@
 The rest of this paper is structured as follows:  @sec-mydatasection presents the data sources and methodology, @sec-mymodel presents the forecasting model, then  @sec-myresults followed by  @sec-mydisc, which discusses the results and their implications.
-@sec-mylimits concludes with a discussion of the limitations of the study and suggestions for future research, followed by @sec-appenxA with The Washington Post’s pollster methodology, and @sec-appenxB a budgeted idealized methodology.
+@sec-mylimits concludes with a discussion of the limitations of the study and suggestions for future research, followed by Appendix @sec-meth-analysis and @sec-ideal-surv with The Washington Post’s pollster methodology and a budgeted idealized methodology.
 
@@
 This strong alignment suggests that the approach not only fits the training data well but also generalizes effectively to new observations.
-The flexibility of this framework allows for the incorporation of prior information and consideration of variability due to pollster and state effects, enhancing overall robustness (Refer to the @sec-appc, [@tbl-pp] for a visual illustration).
+The flexibility of this framework allows for the incorporation of prior information and consideration of variability due to pollster and state effects, enhancing overall robustness (Refer to the @sec-app-pp, [@tbl-pp] for a visual illustration).
 
@@
 
-# Appendix A: Methodology Analysis of The Washington Post Polling  {#sec-appenxA}
+# Appendix {-}
+
+# Methodology Analysis of The Washington Post Polling {#sec-meth-analysis}
 With an evaluation of sampling methodology, recruitment, handling non-response, and questionnaire design, this appendix offers an analysis of the polling methodology used by The Washington Post in collaboration with ABC News.
@@
 
-# Appendix B: Idealized Survey & Methodology - $100K Budget {#sec-appenxB}
+# Idealized Survey & Methodology - $100K Budget {#sec-ideal-surv}
 
@@
 
-# Appendix C {#sec-appc}
+# Posterior Predictive Check {#sec-app-pp}
 * [@tbl-pp]
